### 2015-05-06 version1.0/application.tex
adding ms

%% APPLICATION
%%

We have presented a HBM to infer the probability distribution of a star's
surface gravity and bulk density using flicker. These models are calibrated by a 
comparison of flicker to asteroseismic values. In this section, we provide a
comparison of the predicted $\rho_{\star}$ values from our model versus an
independent metric, in order to validate our predictictions.

A suitable comparison comes from stars with transiting planets, where the shape 
of the transit light curve may also be used to derive $\rho_{\star}$. Several
effects can change

### 2015-05-06 version1.0/introduction.tex
adding ms

%% INTRODUCTION
%%

Stellar characterization techniaues overview, traditionally frequentist

use of HBMs, particularly for exoplanets lately

Here we introduce an application of HBMs to fundamental stellar chaarcetrization
specifically for flicker

### 2015-05-06 version1.0/manuscript.tex
adding ms

%\documentclass[12pt,preprint]{aastex}
\documentclass[apjl]{emulateapj}
\usepackage{comment}
\usepackage{ifthen}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{wasysym}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage[%
	colorlinks=true,	% false: boxed links; true: colored links
	urlcolor=blue,		% color of external links
	pdfpagelabels=true,
	hypertexnames=true,
	plainpages=false,
	naturalnames=true,
%	bookmarks=true,         % show bookmarks bar?
%	unicode=false,          % non-Latin characters in Acrobat’s bookmarks
%	pdftoolbar=true,        % show Acrobat’s toolbar?
%	pdfmenubar=true,        % show Acrobat’s menu?
%	pdffitwindow=false,     % window fit to page when opened
%	pdfstartview={FitH},    % fits the width of the page to the window
	pdftitle={Priors for Orbital Eccentricity},    % title
	pdfauthor={David M. Kipping},     % author
%	pdfsubject={Subject},   % subject of the document
%	pdfcreator={Creator},   % creator of the document
%	pdfproducer={Producer}, % producer of the document
%	pdfkeywords={keyword1} {key2} {key3}, % list of keywords
%	pdfnewwindow=true,      % links in new window
	linkcolor=red,          % color of internal links (change box color with linkbordercolor)
	citecolor=green,        % color of links to bibliography
%	filecolor=magenta,      % color of file links
        ]{hyperref}
%

% #############################################################################
% Specific latex definitions for this paper

\input{shortcuts.tex}

%% #############################################################################

%% VARIABLE DEFINITIONS
%% 
\newboolean{emulateapj}
\setboolean{emulateapj}{true}

\newboolean{astroph}
\setboolean{astroph}{true}

\newboolean{png}
\setboolean{png}{false}

\newboolean{color}
\setboolean{color}{true}

%% ############################################################################

\shortauthors{Angus \& Kipping}
\shorttitle{Discovery of a Transiting Planet Near the Snow-Line}
\ifthenelse{\boolean{emulateapj}}{
    \newcommand{\titledag}{$\dagger$}
}{
    \newcommand{\titledag}{\dagger}
}
\def\mathbi#1{\textbf{\em #1}}

\begin{document}

%% Titlepage
\title {Hierarchical Bayesian Inference of Fundamental Stellar Parameters:\\
Calibration of Flicker to Surface Gravity and Mean Density
\altaffilmark{\titledag}}

%% Authors
\author{
	{\bf	R.~Angus\altaffilmark{1,2},
		D.~M.~Kipping\altaffilmark{2,3},
	}
}

\altaffiltext{1}{Dept. of Physics, University of Oxford, UK}

\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics,
		Cambridge, MA 02138, USA; email: dkipping@cfa.harvard.edu}

\altaffiltext{3}{Department of Astronomy, Columbia University, 550 W 120th St., 
                 New York, NY 10027, US}

\altaffiltext{$\dagger$}{
%%
Based on archival data of the \emph{Kepler} telescope. 
}

%% EOF authors

% #####################################################################
%% abstract
\begin{abstract}

TBD

\end{abstract}

% #####################################################################
%% keywords
\keywords{
%%
	stars: fundamental parameters --- techniques: photometric --- 
        methods: statistical
%%
}

%% EOF keywords
%% EOF titlepage

% #####################################################################
%% Introduction
\section{INTRODUCTION}
\label{sec:intro}

\input{introduction.tex}

\section{HIERARCHICAL BAYESIAN MODEL}
\label{sec:HBM}

\input{HBM.tex}

\section{APPLICATION}
\label{sec:appication}

\input{application.tex}

\section{DISCUSSION}
\label{sec:discussion}

\input{discussion.tex}

% #####################################################################
%% Acknowledgements
\acknowledgements
\section*{Acknowledgements}

RA thanks...
DMK thanks...

%% EOF Acknowledgements

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Abe et al.}{2000}]{abe:2000} 
Abe, Y., Ohtani, E., Okuchi, T., Righter, K. \& Drake, M., 2000, Origin of the
Earth and Moon, ed. R. M. Canup \& K. Righter (Tucson: Univ. Arizona
Press), 1, 413
\end{thebibliography}

\end{document}

%% EOF document

### 2015-05-07 version1.0/HBM.tex
hack day cont

%% HBM
%%

TBD

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
TBD
} 
\label{fig:rhostar}
\end{center}
\end{figure}

### 2015-05-07 version1.0/introduction.tex
hack day cont

@@
 
-Stellar characterization techniaues overview, traditionally frequentist
+% Stellar characterization techniaues overview, traditionally frequentist
 
-use of HBMs, particularly for exoplanets lately
+% who cares about characterization?
 
-Here we introduce an application of HBMs to fundamental stellar chaarcetrization
-specifically for flicker
+Accurate and precise stellar characterization forms the bedrock of a number
+fields within astronomy.
+For example, stellar populations, galactic archaeology, binary stars,
+asteroseismology and exoplanet studies all rely on inferences of fundamental
+stellar parameters.
+These stellar parameters are inferred from a variety of observable properties
+measured using spectroscopy, photometry, interferometry, etc.
+Many stellar parameters are inferred via theoretical stellar models, e.g.
+fitting model isochrones to spectroscopic measurements, and
+some are inferred using empirically calibrated relations.
+Regardless of the approach, be it theoretical or empirical, the methods used
+for the inference of stellar parameters are traditionally frequentist.
+However, a shift towards probabilistic methods in other areas of astronomy,
+and particularly within the exoplanet community, prompts the need to treat
+stars in the same probabilistic framework.
+
+The demand for probilistic stellar parameters not only motivated by the fact
+that probability distributions are far more representative of our `beliefs'
+about astrophysical parameters, but also
+
+%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
+%(HBM) to model the dispersion, or intrinsic scatter in relations between
+%astrophysical properties.
+%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
+%relation for sub-Neptunes and infer the level of astrophysical dispersion
+
+% We use the example of flicker; the brightness fluctuations of stars on
+% timescales less than 8 hours, which has been shown to correlate both with
+% surface gravity and stellar density, to demonstrate our probilistic approach.
+
+% Previous studies calibrating the relations between flicker, surface
+% gravity and stellar density have assumed that the relations between these
+% properties are deterministic, i.e. if you know the flicker of a star with
+% infinite precision, then you can exactly predict the dependant property.
+% We here provide a probabilistic recalibration of the relations between surface
+% gravity, stellar density and flicker, allowing for the astrophysical
+% dispersion in the relation.
+% This probabilitic recalibration will produce predictive {\it distributions}
+% for stellar density and surface gravity, given a measurement of flicker.
+
+% use of HBMs, particularly for exoplanets lately
+
+% Here we introduce an application of HBMs to fundamental stellar
+% characterization specifically for flicker.

### 2015-05-07 version1.0/manuscript.tex
hack day cont

@@
 \documentclass[apjl]{emulateapj}
+
+\pdfoutput=1
+
 \usepackage{comment}
@@
 %% Titlepage
-\title {Hierarchical Bayesian Inference of Fundamental Stellar Parameters:\\
-Calibration of Flicker to Surface Gravity and Mean Density
+\title {Probabilistic Inference of Stellar Parameters:\\
+Application to Flickering Stars %to Surface Gravity and Mean Density
 \altaffilmark{\titledag}}
@@
 
+One of the more recent tools developed to characterize stars is known as 
+``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an 
+8-hour timescale (dubbed $F_8$) in a broad visible bandpass time series 
+photometric light curve, such as that from \textit{Kepler} or the upcoming TESS 
+mission. A more detailed account of the proceedure to calculate flicker is 
+described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker 
+displays a remarkable correlation to the asteroseismically determined parent 
+star surface gravities (\logg). Turning this round, the observation implies 
+that flicker can be used to blindly infer surface gravities at the level of 
+$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light 
+curves available through the array of exoplanet transit missions flying and 
+scheduled to launch.
+
+\citet{cranmer:2014} demonstrated that stellar surface granulation is expected
+to reproduce a flicker effect in close agreement with that observed by 
+\citet{bastien:2013}, providing a physically-plausible explanation. Since
+surface gravity is highly correlated to mean stellar density (\rhostar) on 
+evolutionary tracks, \citet{kipping:2014} showed that flicker can be also 
+be used to infer \rhostar, which is more useful for exoplanet transit 
+analysis.
+
+Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
+of the problem which are attractive for our purposes of a simple demonstration
+of probabilistic inference of stellar parameters. Firstly, in log-log space the
+relationship is very simple, appearing to be linear. Secondly, there is a
+sufficiently number of points in the sample (439 stars) to constrain a 
+population-based model. Thirdly, there is significant excess scatter around the
+best-fitting relation implying that a deterministic model is inadequate. For
+these reasons, we identify the calibration of flicker to \logg\ and \rhostar\
+as a well-posed problem to first demonstrate probabilistic inference in the 
+arena of stellar characterization.
+
 \section{HIERARCHICAL BAYESIAN MODEL}
@@
 \begin{thebibliography}{99}
-\bibitem[\protect\citeauthoryear{Abe et al.}{2000}]{abe:2000} 
-Abe, Y., Ohtani, E., Okuchi, T., Righter, K. \& Drake, M., 2000, Origin of the
-Earth and Moon, ed. R. M. Canup \& K. Righter (Tucson: Univ. Arizona
-Press), 1, 413
+\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien, 
+F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
+\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014} 
+Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781, 
+124
+\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014} 
+Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
+Buchhave, L.~A., 2014, ApJ, 785, 32
+\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015} 
+Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted 
+(astro-ph:1504.07557)
 \end{thebibliography}

### 2015-05-07 code/plots.py

@@ -28,9 +28,9 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
 
     if fname == "rho":
         plt.ylim(.5, 4.)
-        plt.ylabel("$\log_{10}(\\rho_{star}[\mathrm{kgm}^{-3}])$")
+        plt.ylabel("$\log_{10}(\\rho_{star}[\mathrm{kg~m}^{-3}])$")
         col = cols.blue
-        plt.text(1.7, 3.5, "$\log_{10}(\\rho_{star}) = %.3f \\times F8 + %.3f$"
+        plt.text(1.7, 3.5, "$\log_{10}(\\rho_{\star}) = %.3f \\times F8 + %.3f$"
                  % (alpha, beta))
         plt.text(1.2, 1.5, "$\\tau = %.3f,~\sigma = %.3f$"
                  % (tau, np.sqrt(tau)))
@@ -41,7 +41,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
         plt.ylabel("$\log(g)$")
         plt.text(1.7, 4.5, "$\log(g) = %.3f \\times F8 + %.3f$"
                  % (alpha, beta))
-        plt.text(1.2, 3.5, "$\\tau = %.3f,~\sigma = %.3f$" %
+        plt.text(1.2, 3.5, "\sigma = %.3f$" %
                  (tau, np.sqrt(tau)))
 
     plt.fill_between(ys, ((ys-alpha)/beta)-sigma, ((ys-alpha)/beta)+sigma,
@@ -53,8 +53,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     plt.plot(ys, (ys-alpha)/beta, ".2", linewidth=1)
 
     plt.xlim(1, 2.4)
-    plt.xlabel("$\log_{10}\mathrm{(8-hour~flicker~[ppm])}$")
-    plt.savefig("/Users/angusr/Dropbox/Flickerhackday/figs/%s_vs_flicker.eps" % fname)
+    plt.xlabel("$\log_{10}\mathrm{(F}_8~\mathrm{[ppm]})$")
+    plt.savefig("/Users/angusr/Dropbox/Flickerhackday/figs/%s_vs_flicker.pdf" % fname)
     plt.savefig("flicker_inv_%s" % fname)
 
 def make_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
[…]

### 2015-05-07 version1.0/introduction.tex
leaving with Dave

@@
 that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, but also
+about astrophysical parameters, but also... (ruth will continue this sentence)
 
@@
 %relation for sub-Neptunes and infer the level of astrophysical dispersion
-
-% We use the example of flicker; the brightness fluctuations of stars on
-% timescales less than 8 hours, which has been shown to correlate both with
-% surface gravity and stellar density, to demonstrate our probilistic approach.
-
-% Previous studies calibrating the relations between flicker, surface
-% gravity and stellar density have assumed that the relations between these
-% properties are deterministic, i.e. if you know the flicker of a star with
-% infinite precision, then you can exactly predict the dependant property.
-% We here provide a probabilistic recalibration of the relations between surface
-% gravity, stellar density and flicker, allowing for the astrophysical
-% dispersion in the relation.
-% This probabilitic recalibration will produce predictive {\it distributions}
-% for stellar density and surface gravity, given a measurement of flicker.
-
-% use of HBMs, particularly for exoplanets lately
-
-% Here we introduce an application of HBMs to fundamental stellar
-% characterization specifically for flicker.

### 2015-05-12 code/main.py
adding main

@@ -1,51 +1,64 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from noisy_plane import generate_samples, lnlike
+from noisy_plane import generate_samples, lnlike, lnlikeH
 from model import model1, model
 import emcee
 import triangle
+import sys
 
 def lnprior(pars):
     return 0.
 
 def lnprob(pars, samples, obs, u):
-    return lnlike(pars, samples, obs, u) + lnprior(pars)
+    return lnlikeH(pars, samples, obs, u) + lnprior(pars)
 
-if __name__ == "__main__":
+def MCMC(whichx, nsamp):
+
+    # set initial params
+    rho_pars = [-.5, 3.6, .065]
+    logg_pars = [-1.850, 5.413, .065]
+    pars_init = logg_pars
+    if whichx == "rho":
+        pars_init = rho_pars
 
     # load data
-    # x = f, y = rho, z = teff
-    x, xerr, y, yerr, z, zerr = np.genfromtxt("data/log.dat").T
+    fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
+    fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
+    x, xerr, y, yerr = fl[:10], flerr[:10], l[:10], lerr[:10]
+    if whichx == "rho":
+        x, xerr, y, yerr = fr[:10], frerr[:10], r[:10], rerr[:10]
 
+    # format data and generate samples
     obs = np.vstack((x, y))
     u = np.vstack((xerr, yerr))
-    nsamp = 3
     s = generate_samples(obs, u, nsamp)
 
-    pars_init = [-1.850, 5.413]
-
+    # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
     pos = [pars_init + 1e-4*np.random.randn(ndim) for i in range(nwalkers)]
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
                                     args=(s, obs, u))
-
     print "burning in..."
     pos, _, _, = sampler.run_mcmc(pos, 500)
     sampler.reset()
     print "production run..."
     sampler.run_mcmc(pos, 1000)
     samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
-    m, c = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
[…]

### 2015-05-12 code/main.py
making corrections

@@ -15,8 +15,8 @@ def lnprob(pars, samples, obs, u):
 def MCMC(whichx, nsamp):
 
     # set initial params
-    rho_pars = [-.5, 3.6, .065]
-    logg_pars = [-1.850, 5.413, .065]
+    rho_pars = [-2., 6., .0065]
+    logg_pars = [-1.850, 7., .0065]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -24,9 +24,9 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    x, xerr, y, yerr = fl[:10], flerr[:10], l[:10], lerr[:10]
+    x, xerr, y, yerr = fl, flerr, l, lerr
     if whichx == "rho":
-        x, xerr, y, yerr = fr[:10], frerr[:10], r[:10], rerr[:10]
+        x, xerr, y, yerr = fr, frerr, r, rerr
 
     # format data and generate samples
     obs = np.vstack((x, y))
@@ -51,6 +51,7 @@ def MCMC(whichx, nsamp):
     # make plots
     plt.clf()
     plt.plot(s[0, :, :], s[1, :, :], "r.", markersize=2)
+    print np.shape(s[0, :, :])
     plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0)
     print pars_init, pars
     plt.plot(x, model1(pars_init, x))
@@ -61,4 +62,4 @@ def MCMC(whichx, nsamp):
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    MCMC(whichx, 10)
+    MCMC(whichx, 2)

### 2015-05-12 code/pgm.py

@@ -8,12 +8,12 @@ import daft
 pgm = daft.PGM([3.2, 4], origin=[0.3, 0.3])
 
 # Data.
-pgm.add_node(daft.Node("rho_obs", r"$\hat{\rho_n}$", 1, 1, observed=True))
-pgm.add_node(daft.Node("f_obs", r"$\hat{F8_n}$", 2, 1, observed=True))
+pgm.add_node(daft.Node("rho_obs", r"$\hat{\rho_n}$", 2, 1, observed=True))
+pgm.add_node(daft.Node("f_obs", r"$\hat{F_{8, n}}$", 1, 1, observed=True))
 
 # Latent variables.
-pgm.add_node(daft.Node("rho", r"$\rho_n$", 1, 1.8))
-pgm.add_node(daft.Node("f", r"$F8_n$", 2, 1.8))
+pgm.add_node(daft.Node("rho", r"$\rho_n$", 2, 1.8))
+pgm.add_node(daft.Node("f", r"$F_{8, n}$", 1, 1.8))
 pgm.add_node(daft.Node("mean", r"$\mu_n$", 1.5, 2.5))
 
 # add the plate.
@@ -23,17 +23,17 @@ pgm.add_plate(daft.Plate([0.5, 0.5, 2, 2.5], label=r"$n = 1, \cdots, N$",
 # Model parameters
 pgm.add_node(daft.Node("alpha", r"$\alpha$", 1, 3.5))
 pgm.add_node(daft.Node("beta", r"$\beta$", 2, 3.5))
-pgm.add_node(daft.Node("tau", r"$\tau$", 3, 1.8))
+pgm.add_node(daft.Node("sigma", r"$\sigma$", 3, 1.8))
 
 # # Add in the edges.
 pgm.add_edge("rho_obs", "rho")
-pgm.add_edge("rho", "mean")
-pgm.add_edge("mean", "f")
-pgm.add_edge("f", "f_obs")
+pgm.add_edge("mean", "rho")
+pgm.add_edge("f", "mean")
+pgm.add_edge("f_obs", "f")
 pgm.add_edge("alpha", "mean")
 pgm.add_edge("beta", "mean")
-pgm.add_edge("tau", "f")
+pgm.add_edge("sigma", "rho")
 
 # Render and save.
 pgm.render()
-pgm.figure.savefig("pgm.png", dpi=150)
+pgm.figure.savefig("pgm.pdf", dpi=150)

### 2015-05-12 code/plots.py

@@ -1,8 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import triangle
-from params import plot_params
-reb = plot_params()
 from colours import plot_colours
 cols = plot_colours()
 import scipy.interpolate as spi
@@ -27,30 +25,42 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     plt.clf()
 
     if fname == "rho":
-        plt.ylim(.5, 4.)
-        plt.ylabel("$\log_{10}(\\rho_{star}[\mathrm{kg~m}^{-3}])$")
+        plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = cols.blue
-        plt.text(1.7, 3.5, "$\log_{10}(\\rho_{\star}) = %.3f \\times F8 + %.3f$"
-                 % (alpha, beta))
-        plt.text(1.2, 1.5, "$\\tau = %.3f,~\sigma = %.3f$"
-                 % (tau, np.sqrt(tau)))
+        plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} (\\alpha + \\beta F_8, \sigma)$")
+        plt.text(1.7, .2, "$\\alpha = %.3f$" % (alpha-2))
+        plt.text(1.7, .05, "$\\beta = %.3f$" % beta)
+        plt.text(1.7, -.1, "$\sigma = %.3f$" % tau**.5)
+        plt.fill_between(ys, ((ys-alpha)/beta)-sigma-3,
+                         ((ys-alpha)/beta)+sigma-3,
+                         color=col, alpha=.3, edgecolor="w")
+        plt.fill_between(ys, ((ys-alpha)/beta)-2*sigma-3,
+                         ((ys-alpha)/beta)+2*sigma-3,
+                         color=col, alpha=.2, edgecolor="w")
+        dx = x/xerr
+        plt.errorbar(y, x-3, xerr=yerr, yerr=xerr, fmt="k.", capsize=0,
+                     alpha=.5, ecolor=".5", mec=".2")
+        plt.plot(ys, (ys-alpha)/beta-3, ".2", linewidth=1)
+        plt.ylim(-2, 1)
+        plt.subplots_adjust(bottom=.1)
 
     elif fname == "logg":
         plt.ylim(3, 5)
         col = cols.pink
[…]

### 2015-05-12 code/rho2d.py

@@ -16,10 +16,10 @@ def lnprob(pars, samples, obs, u, is2d=False):
 if __name__ == "__main__":
 
     # load data. x = f, y = rho, z = teff
-    x, xerr, y, yerr = np.genfromtxt("data/flickers.dat").T
+    x, xerr, y, yerr = np.genfromtxt("../data/flickers.dat").T
 
     # generate nsamp importance samples unless 2d is False
-    is2d = False
+    is2d = True
     obs = np.vstack((x, y))
     u = np.vstack((xerr, yerr))
     nsamp = 1

### 2015-05-12 version1.0/introduction.tex
adding words

--- 2015-05-06 version1.0/introduction.tex
@@
 
-Stellar characterization techniaues overview, traditionally frequentist
+% Stellar characterization techniaues overview, traditionally frequentist
 
-use of HBMs, particularly for exoplanets lately
+% who cares about characterization?
 
-Here we introduce an application of HBMs to fundamental stellar chaarcetrization
-specifically for flicker
+Accurate stellar characterization plays a vital role for many active research 
+fields within astronomy. For example, stellar populations, galactic archaeology, 
+the study of binary stars, asteroseismology and exoplanet studies all rely on 
+inferences of basic stellar parameters to varying degrees. Empirically-derived
+and reliable estimates are of particular value, increasing our confidence in the
+end-product results built upon these inputs.
+
+Basic stellar parameters, such as effective temperature and surface gravity, 
+can be inferred using one (or more) of several types of observations, e.g. 
+spectroscopy, photometry, interferometry, etc. This inference can be done by
+invoking theoretical models or by building an empirical calibration library.
+As an example, an observed stellar spectrum could be matched against a library of 
+theoretical spectra generated using stellar atmosphere models, or, against a 
+library of observed spectra of ``standard stars'', serving as calibrators.
+%Fundamental stellar parameters, such as mass and radius, which cannot be 
+%directly inferred from the available observations can often be inferred by 
+%invoking theoretical stellar models, e.g. fitting model isochrones to 
+%spectroscopic measurements.
+Regardless of the approach, be it theoretical or empirical, the methods used
+for the inference of stellar parameters are traditionally ``deterministic''.
+In this context, a deterministic model can be loosely described as one where
+a particular observational input always returns a single-valued output for a 
+parameter of interest. In this framework, nature itself has no variance and
+our underlying model is considered to be a perfect description of reality.
+
+An alternative approach for inferring model parameters is the 
+``probabilistic'' one. Unlike the deterministic case, a single observational 
+input is intepretted to be caused by a range of possible model parameters
+described by a probability distribution. This statement is true even in the case
+of a perfect observation of infinite signal-to-noise, since our underlying 
+model itself comes with uncertainty. In recent years, there has been a shift 
+towards such methods in several areas of astronomy, particularly within the 
+exoplanet community. For example, \citet{wolfgang:2015} considered that
+the mass-radius relationship of exoplanets is probabilistic, since a particular
+sized planet could be due to a range of planet masses, thanks to variances in
+compositions, environment and other circumstances. These recent demonstrations
+in exoplanetary science have prompted us to consider the need for treating the 
+parent stars in the same probabilistic framework, with potential applications 
+spanning many fields of astronomy.
+
+The demand for probilistic stellar parameters is not only motivated by the fact
+that probability distributions are far more representative of our `beliefs'
+about astrophysical parameters, but also... (ruth will continue this sentence)
+
+%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
+%(HBM) to model the dispersion, or intrinsic scatter in relations between
+%astrophysical properties.
+%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
+%relation for sub-Neptunes and infer the level of astrophysical dispersion

### 2015-05-12 version1.0/manuscript.tex
adding words

@@
 \shortauthors{Angus \& Kipping}
-\shorttitle{Discovery of a Transiting Planet Near the Snow-Line}
+\shorttitle{Probabilistic Inference of Stellar Parameters}
 \ifthenelse{\boolean{emulateapj}}{
@@
 %% Titlepage
-\title {Probabilistic Inference of Stellar Parameters:\\
+\title {Probabilistic Inference of Basic Stellar Parameters:\\
 Application to Flickering Stars %to Surface Gravity and Mean Density

### 2015-05-12 code/main.py
debugging multi-modal

@@ -5,6 +5,7 @@ from model import model1, model
 import emcee
 import triangle
 import sys
+import h5py
 
 def lnprior(pars):
     return 0.
@@ -24,14 +25,23 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    x, xerr, y, yerr = fl, flerr, l, lerr
+    nd = 20
+    x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
-        x, xerr, y, yerr = fr, frerr, r, rerr
+        x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
 
     # format data and generate samples
     obs = np.vstack((x, y))
     u = np.vstack((xerr, yerr))
     s = generate_samples(obs, u, nsamp)
+
+#     plt.clf()
+#     x, y = obs[0, :], obs[1, :]
+#     xerr, yerr = u[0, :], u[1, :]
+#     plt.plot(s[0, :, :], s[1, :, :], "r.", markersize=2, alpha=.3, zorder=0)
+#     plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="ko", capsize=0, zorder=1)
+#     plt.show()
+#     assert 0
 
     # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
@@ -48,18 +58,46 @@ def MCMC(whichx, nsamp):
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))
     pars = [m[0], c[0], sig[0]]
 
-    # make plots
+    # save samples
+    print np.shape(samp)
+    f = h5py.File("%s_samples.h5" % whichx, "w")
+    data = f.create_dataset("samples", np.shape(samp))
+    data[:, 0] = samp[:, 0]
+    data[:, 1] = samp[:, 1]
+    data[:, 2] = samp[:, 2]
+    f.close()
+    return s
+
+def make_plots(whichx, s):
+
+    # load data
+    fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
+    fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
+    nd = 20
[…]

### 2015-05-12 code/main.py
now with mixture model

@@ -1,6 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from noisy_plane import generate_samples, lnlike, lnlikeH
+from noisy_plane import generate_samples, lnlike, lnlikeH, lnlikeHM
 from model import model1, model
 import emcee
 import triangle
@@ -8,16 +8,19 @@ import sys
 import h5py
 
 def lnprior(pars):
-    return 0.
+    m, c, sig, Y, V, P = pars
+    if 0. < P < 1. and 0. < V:
+        return 0.
+    return -np.inf
 
 def lnprob(pars, samples, obs, u):
-    return lnlikeH(pars, samples, obs, u) + lnprior(pars)
+    return lnlikeHM(pars, samples, obs, u) + lnprior(pars)
 
 def MCMC(whichx, nsamp):
 
-    # set initial params
-    rho_pars = [-2., 6., .0065]
-    logg_pars = [-1.850, 7., .0065]
+    # set initial params. slope, intercept, sigma, Y, V, P
+    rho_pars = [-2., 6., .0065, -.5, .5, .1]
+    logg_pars = [-1.850, 7., .0065, 4., 1., .1]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -35,14 +38,6 @@ def MCMC(whichx, nsamp):
     u = np.vstack((xerr, yerr))
     s = generate_samples(obs, u, nsamp)
 
-#     plt.clf()
-#     x, y = obs[0, :], obs[1, :]
-#     xerr, yerr = u[0, :], u[1, :]
-#     plt.plot(s[0, :, :], s[1, :, :], "r.", markersize=2, alpha=.3, zorder=0)
-#     plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="ko", capsize=0, zorder=1)
-#     plt.show()
-#     assert 0
-
     # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
     pos = [pars_init + 1e-4*np.random.randn(ndim) for i in range(nwalkers)]
@@ -54,7 +49,7 @@ def MCMC(whichx, nsamp):
     print "production run..."
     sampler.run_mcmc(pos, 1000)
     samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
[…]

### 2015-05-12 code/main.py
testing logarithmic errorbars

@@ -25,10 +25,42 @@ def MCMC(whichx, nsamp):
     if whichx == "rho":
         pars_init = rho_pars
 
-    # load data
-    fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
-    fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    nd = 20
+#     # load data
+#     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
+#     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
+#     nd = 20
+
+    data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
+    r, rerrp, rerrm = data[7:10]
+    f, ferr = data[20:22]
+    logg, loggerrp, loggerrm = data[10:13]
+
+    plt.clf()
+    plt.subplot(2, 1, 1)
+    plt.errorbar(r, f, xerr=(rerrm, rerrp), yerr=ferr, fmt="k.", capsize=0)
+
+    rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
+    r = np.log10(1000*r)
+    ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
+    f = np.log10(1000*f)
+
+    plt.subplot(2, 1, 2)
+    plt.errorbar(r, f, xerr=(rerrm, rerrp), yerr=(ferrm, ferrp), fmt="k.",
+                 capsize=0)
+    plt.savefig("test")
+    raw_input('enter')
+
+    plt.clf()
+    plt.subplot(2, 1, 1)
+    plt.errorbar(logg, f, xerr=(loggerrm, loggerrp), yerr=ferr, fmt="k.", capsize=0)
+
+    ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
+    f = np.log10(1000*f)
+    plt.subplot(2, 1, 2)
+    plt.errorbar(logg, f, xerr=(loggerrm, loggerrp), yerr=(ferrm, ferrp), fmt="k.", capsize=0)
+    plt.savefig("test2")
+    raw_input('enter')
+
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
[…]

### 2015-05-12 code/main.py
back to working, trying to fix errorbars

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from noisy_plane import generate_samples, lnlike, lnlikeH, lnlikeHM
+from noisy_plane import generate_samples, generate_samples_log, \
+        lnlike, lnlikeH, lnlikeHM
 from model import model1, model
 import emcee
 import triangle
@@ -25,42 +26,30 @@ def MCMC(whichx, nsamp):
     if whichx == "rho":
         pars_init = rho_pars
 
-#     # load data
-#     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
-#     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-#     nd = 20
+#     data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
+#     r, rerrp, rerrm = data[7:10]
+#     f, ferr = data[20:22]
+#     logg, loggerrp, loggerrm = data[10:13]
+#     rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
+#     r = np.log10(1000*r)
+#     ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
+#     f = np.log10(1000*f)
+#
+#     nd = len(f)
+#     x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
+#     y, yerrp, yerrm = logg[:nd], loggerrp[:nd], loggerrm[:nd]
+#     xerr = .5*(xerrp + xerrm)
+#     yerr = .5*(yerrp + yerrm)
+#     if whichx == "rho":
+#         x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
+#         y, yerrp, yerrm = r[:nd], rerrp[:nd], rerrm[:nd]
+#         xerr = .5*(xerrp + xerrm)
+#         yerr = .5*(yerrp + yerrm)
 
-    data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
-    r, rerrp, rerrm = data[7:10]
-    f, ferr = data[20:22]
-    logg, loggerrp, loggerrm = data[10:13]
-
-    plt.clf()
-    plt.subplot(2, 1, 1)
-    plt.errorbar(r, f, xerr=(rerrm, rerrp), yerr=ferr, fmt="k.", capsize=0)
-
[…]

### 2015-05-12 code/plots.py
saving data

@@ -47,7 +47,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     elif fname == "logg":
         plt.ylim(3, 5)
         col = cols.pink
-        plt.ylabel("$\log_{10}(g) [\mathrm{cm~s}^{-2}]$")
+        plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
         plt.text(1.7, 4.7, "$\log(g) \sim \mathcal{N} (\\alpha + \\beta F_8, \sigma)$")
         plt.text(1.7, 4.5, "$\\alpha = %.3f$" % alpha)
         plt.text(1.7, 4.4, "$\\beta = %.3f$" % beta)
@@ -114,6 +114,7 @@ def make_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
 
     resids = y - (alpha+beta*x)
     normed_resids = resids / np.sqrt(yerr**2 + sigma**2)
+    np.savetxt("normed_resids_%s.txt", np.transpose(normed_resids))
     plt.clf()
     plt.hist(normed_resids, 20, histtype="stepfilled", color="w")
     plt.xlabel("Normalised residuals")

### 2015-05-12 version1.0/HBM.tex
adding words
testing sampling

@@
 %%
+
+\subsection{Calibration Data}
+
+For our calibration of \rhostar-$F_8$, we used a sample of Kepler stars with
+both asteroseismic and flicker measurements available. \citet{chaplin:2014}
+report asteroseismic \rhostar\ estimates (and the associated uncertainties) for 
+518 Kepler stars. The authors report three different sets of results, depending 
+on the choice of \Teff\ and \FeH, and in this work we elected to use values 
+reported in their Table 6 over Table 5, and Table 5 over Table 4. We 
+additionally used the 71 additional planet hosting stars with asteroseismology 
+reported in \citet{huber:2013} but not reported in \citet{chaplin:2014}. Values 
+for flicker and range were taken from \citet{kipping:2014}, based upon the 
+methods described in \citet{bastien:2013}. Following \citet{kipping:2014} and 
+for reasons described there-in, we only include targets in our calibration for 
+which:
+
+\begin{itemize}
+\item[{\tiny$\blacksquare$}] Range (defined in \citealt{bastien:2013}) 
+$<1$\,ppt
+\item[{\tiny$\blacksquare$}] $4500<T_{\mathrm{eff}}<6500$\,K
+\item[{\tiny$\blacksquare$}] $K_P<14$
+\item[{\tiny$\blacksquare$}] $1.2 < \log_{10}$($F_8$\,[ppm])$< 2.2$
+\end{itemize}
+
+We use the same sample for our calibration of \logg, except that we exclude the
+\citet{huber:2013} data, since these authors do not provide estimates of
+\logg\footnote{Whilst we could compute \logg\ ourselves from the reported
+masses and radii, this could only be done under the incorrect assumption of
+zero covariance between $M_{\star}$ and $R_{\star}$.}
+
+\subsection{Hierarchical Bayesian Model}
 
@@
 \end{figure}
+
+%%% logg plot
+\begin{figure}
+\begin{center}
+\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
+\caption{
+TBD
+} 
+\label{fig:logg}
+\end{center}
+\end{figure}

### 2015-05-12 version1.0/manuscript.tex
testing sampling

@@
 population-based model. Thirdly, there is significant excess scatter around the
-best-fitting relation implying that a deterministic model is inadequate. For
+best-fitting relation implying that a deterministic model is inadequate. This is
+not surprising given that granulation is a complex and messy process for which
+one should not expect any parametric model to provide a perfect description. For
 these reasons, we identify the calibration of flicker to \logg\ and \rhostar\
@@
 
-\section{HIERARCHICAL BAYESIAN MODEL}
+\section{PROBABILISTIC CALIBRATION}
 \label{sec:HBM}
@@
 F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
+\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014} 
+Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
 \bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014} 
@@
 124
+\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013} 
+Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
+127
 \bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014} 

### 2015-05-12 code/main.py

@@ -57,11 +57,10 @@ def MCMC(whichx, nsamp):
     # format data and generate samples
     obs = np.vstack((x, y))
     u = np.vstack((xerr, yerr))
-#     up = np.vstack((xerrp, yerrm))
-#     um = np.vstack((xerrp, yerrm))
-#     u = .5*(up + um)
-#     s = generate_samples_log(obs, up, um, nsamp)
-    s = generate_samples(obs, u, nsamp)
+    up = np.vstack((xerr, yerr))
+    um = np.vstack((xerr*.5, yerr*.5))
+    s = generate_samples_log(obs, up, um, nsamp)
+#     s = generate_samples(obs, u, nsamp)
 
     # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
@@ -140,5 +139,5 @@ def make_plots(whichx):
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    MCMC(whichx, 10)
+    MCMC(whichx, 1000)
     make_plots(whichx)

### 2015-05-12 code/main.py
trying to fix sampling

@@ -49,7 +49,7 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    nd = 20
+    nd = 5
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
@@ -139,5 +139,5 @@ def make_plots(whichx):
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    MCMC(whichx, 1000)
+    MCMC(whichx, 500)
     make_plots(whichx)

### 2015-06-01 code/pgm.py
adding pgm.py

@@ -8,12 +8,12 @@ import daft
 pgm = daft.PGM([3.2, 4], origin=[0.3, 0.3])
 
 # Data.
-pgm.add_node(daft.Node("rho_obs", r"$\hat{\rho_n}$", 2, 1, observed=True))
-pgm.add_node(daft.Node("f_obs", r"$\hat{F_{8, n}}$", 1, 1, observed=True))
+pgm.add_node(daft.Node("rho_obs", r"$\hat{\rho_n}$", 1, 1, observed=True))
+pgm.add_node(daft.Node("f_obs", r"$\hat{F8_n}$", 2, 1, observed=True))
 
 # Latent variables.
-pgm.add_node(daft.Node("rho", r"$\rho_n$", 2, 1.8))
-pgm.add_node(daft.Node("f", r"$F_{8, n}$", 1, 1.8))
+pgm.add_node(daft.Node("rho", r"$\rho_n$", 1, 1.8))
+pgm.add_node(daft.Node("f", r"$F8_n$", 2, 1.8))
 pgm.add_node(daft.Node("mean", r"$\mu_n$", 1.5, 2.5))
 
 # add the plate.
@@ -23,17 +23,17 @@ pgm.add_plate(daft.Plate([0.5, 0.5, 2, 2.5], label=r"$n = 1, \cdots, N$",
 # Model parameters
 pgm.add_node(daft.Node("alpha", r"$\alpha$", 1, 3.5))
 pgm.add_node(daft.Node("beta", r"$\beta$", 2, 3.5))
-pgm.add_node(daft.Node("sigma", r"$\sigma$", 3, 1.8))
+pgm.add_node(daft.Node("tau", r"$\sigma$", 3, 1.8))
 
 # # Add in the edges.
-pgm.add_edge("rho_obs", "rho")
-pgm.add_edge("mean", "rho")
-pgm.add_edge("f", "mean")
-pgm.add_edge("f_obs", "f")
+pgm.add_edge("rho", "rho_obs")
+pgm.add_edge("rho", "mean")
+pgm.add_edge("mean", "f")
+pgm.add_edge("f", "f_obs")
 pgm.add_edge("alpha", "mean")
 pgm.add_edge("beta", "mean")
-pgm.add_edge("sigma", "rho")
+pgm.add_edge("tau", "f")
 
 # Render and save.
 pgm.render()
-pgm.figure.savefig("pgm.pdf", dpi=150)
+pgm.figure.savefig("../figs/pgm.pdf", dpi=150)

### 2015-06-02 code/main.py
fixing log sampling

@@ -8,20 +8,16 @@ import triangle
 import sys
 import h5py
 
-def lnprior(pars):
-    m, c, sig, Y, V, P = pars
-    if 0. < P < 1. and 0. < V:
-        return 0.
-    return -np.inf
+def lnprior(pars, mm=False):
+    return 0.
 
-def lnprob(pars, samples, obs, u):
-    return lnlikeHM(pars, samples, obs, u) + lnprior(pars)
+def lnprob(pars, samples, obs, u, mm=False):
+    return lnlikeH(pars, samples, obs, u) + lnprior(pars)
 
 def MCMC(whichx, nsamp):
 
-    # set initial params. slope, intercept, sigma, Y, V, P
-    rho_pars = [-2., 6., .0065, -.5, .5, .1]
-    logg_pars = [-1.850, 7., .0065, 4., 1., .1]
+    rho_pars = [-2., 6., .0065]
+    logg_pars = [-1.850, 7., .0065]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -34,7 +30,7 @@ def MCMC(whichx, nsamp):
 #     r = np.log10(1000*r)
 #     ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
 #     f = np.log10(1000*f)
-#
+
 #     nd = len(f)
 #     x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
 #     y, yerrp, yerrm = logg[:nd], loggerrp[:nd], loggerrm[:nd]
@@ -49,7 +45,7 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    nd = 5
+    nd = 20
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
@@ -61,6 +57,7 @@ def MCMC(whichx, nsamp):
     um = np.vstack((xerr*.5, yerr*.5))
     s = generate_samples_log(obs, up, um, nsamp)
 #     s = generate_samples(obs, u, nsamp)
+#     print np.shape(s)
 
[…]

### 2015-06-02 code/main.py
using big data set

@@ -22,33 +22,36 @@ def MCMC(whichx, nsamp):
     if whichx == "rho":
         pars_init = rho_pars
 
-#     data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
-#     r, rerrp, rerrm = data[7:10]
-#     f, ferr = data[20:22]
-#     logg, loggerrp, loggerrm = data[10:13]
-#     rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
-#     r = np.log10(1000*r)
-#     ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
-#     f = np.log10(1000*f)
-
-#     nd = len(f)
-#     x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
-#     y, yerrp, yerrm = logg[:nd], loggerrp[:nd], loggerrm[:nd]
-#     xerr = .5*(xerrp + xerrm)
-#     yerr = .5*(yerrp + yerrm)
-#     if whichx == "rho":
-#         x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
-#         y, yerrp, yerrm = r[:nd], rerrp[:nd], rerrm[:nd]
-#         xerr = .5*(xerrp + xerrm)
-#         yerr = .5*(yerrp + yerrm)
-
     # load data
-    fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
-    fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
+    data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
+    r, rerrp, rerrm = data[7:10]
+    f, ferr = data[20:22]
+    logg, loggerrp, loggerrm = data[10:13]
+    rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
+    r = np.log10(1000*r)
+    ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
+    f = np.log10(1000*f)
+
+    # format data
+#     nd = len(f)
     nd = 20
-    x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
+    x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
+    y, yerrp, yerrm = logg[:nd], loggerrp[:nd], loggerrm[:nd]
[…]

### 2015-06-02 version1.0/HBM.tex
ruth adds words

@@
 
-TBD
+We model the stochastic relationship between flicker, log($g$) and stellar
+density, accounting for the fact that there exists some intrinsic scatter in
+the dependent variable, and including the heteroskedastic uncertainties on both
+the dependent and independent variables.
+There are two excellent reasons for modelling the relation stochastically;
+firstly, if the intrinsic scatter is ignored and the relation between
+variables is assumed to be deterministic, those data points with smaller
+measurement uncertainties may have an unrepresentative greater weighting
+during the fitting process.
+Secondly, we are interested in producing probability distributions over stellar
+densities and surface gravities, as opposed to point estimates, and propagating
+these probability distributions through to subsequent analyses.
+Several recent studies have required posterior Probability Distribution
+Function (PDF) samples, in order to conduct their (hierarchical)
+analyses \citep[e.g.][]{Rogers2014, Foreman-Mackey2014, Angus2015}.
+Including observational uncertainties on the independent variable within our
+analysis is also important, because ordinary least squares regression methods
+performed on data with two-dimensional uncertainties will result in a slope
+that is biased towards zero
+(e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
+or a demonstration of the effects of neglecting intrinsic scatter and
+two-dimensional uncertainties, see Kelly (2007).
+
+The two models describing the relationship between flicker, logg and stellar
+density can be written as
+\begin{equation}
+	\log(g) \sim \mathcal{N}(\mu = \alpha + \beta F_8, \sigma^2 = \
+	\sigma^2 + \gamma F_8)
+\end{equation}
+\label{eq:logg}
+and
+\begin{equation}
+	\rho_* \sim \mathcal{N}(\mu = \delta + \epsilon F_8, \sigma^2 = \
+	\sigma^2 + \zeta F_8),
+\end{equation}
+\label{eq:rho}
+where $F_8$ is flicker and $\rho_*$ is stellar density.
+The free parameters of the two models are $\alpha$, $\beta$, $\sigma_G$,
+$\gamma$, $\delta$, $\epsilon$ and $\zeta$.
+The two relationships can be described as Gaussian distributions with means
+given by the equation of a straight line, and standard deviations given by
+observational uncertainties, plus some dispersion which is a function of the
+dependent variable.
 

### 2015-06-02 code/main.py
going back to original data file

@@ -1,7 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from noisy_plane import generate_samples, generate_samples_log, \
-        lnlike, lnlikeH, lnlikeHM
+        lnlike, lnlikeH
 from model import model1, model
 import emcee
 import triangle
@@ -22,43 +22,45 @@ def MCMC(whichx, nsamp):
     if whichx == "rho":
         pars_init = rho_pars
 
-    # load data
-    data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
-    r, rerrp, rerrm = data[7:10]
-    f, ferr = data[20:22]
-    logg, loggerrp, loggerrm = data[10:13]
-    rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
-    r = np.log10(1000*r)
-    ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
-    f = np.log10(1000*f)
-
-    # format data
-#     nd = len(f)
-    nd = 20
-    x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
-    y, yerrp, yerrm = logg[:nd], loggerrp[:nd], loggerrm[:nd]
-    xerr = .5*(xerrp + xerrm)
-    yerr = .5*(yerrp + yerrm)
-    if whichx == "rho":
-        x, xerrp, xerrm = f[:nd], ferrp[:nd], ferrm[:nd]
-        y, yerrp, yerrm = r[:nd], rerrp[:nd], rerrm[:nd]
-        xerr = .5*(xerrp + xerrm)
-        yerr = .5*(yerrp + yerrm)
-
 #     # load data
-#     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
-#     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
+#     data = np.genfromtxt("../data/BIGDATA.filtered.dat").T
+#     r, rerrp, rerrm = data[7:10]
+#     f, ferr = data[20:22]
+#     logg, loggerrp, loggerrm = data[10:13]
+#     rerrp, rerrm = rerrp/r/np.log(10), rerrm/r/np.log(10)
+#     r = np.log10(1000*r)
+#     ferrp, ferrm = ferr/f/np.log(10), ferr/f/np.log(10)
[…]

### 2015-06-02 code/main.py
logarithmic multiplicative factor

@@ -1,7 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from noisy_plane import generate_samples, generate_samples_log, \
-        lnlike, lnlikeH
+        lnlike, lnlikeH, lnlikeHF
 from model import model1, model
 import emcee
 import triangle
@@ -9,15 +9,18 @@ import sys
 import h5py
 
 def lnprior(pars, mm=False):
-    return 0.
+    if np.exp(pars[2]) > 0:
+        return 0.
+    return -np.inf
 
 def lnprob(pars, samples, obs, u, mm=False):
-    return lnlikeH(pars, samples, obs, u) + lnprior(pars)
+    return lnlikeHF(pars, samples, obs, u) + lnprior(pars)
 
 def MCMC(whichx, nsamp):
 
     rho_pars = [-2., 6., .0065]
-    logg_pars = [-1.850, 7., .0065]
+#     logg_pars = [-1.850, 7., .0065]
+    logg_pars = [-1.850, 7., np.log(.0065)]  # multiplicative
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -50,7 +53,8 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    nd = len(fr)
+#     nd = len(fr)
+    nd = 20
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]

### 2015-06-03 code/main.py
sampling in log

@@ -9,12 +9,14 @@ import sys
 import h5py
 
 def lnprior(pars, mm=False):
-    if np.exp(pars[2]) > 0:
-        return 0.
-    return -np.inf
+    return 0.
+
+def lnpriorHF(pars):
+    return 0
 
 def lnprob(pars, samples, obs, u, mm=False):
-    return lnlikeHF(pars, samples, obs, u) + lnprior(pars)
+    pars[2] = np.exp(pars[2])
+    return lnlikeHF(pars, samples, obs, u) + lnpriorHF(pars)
 
 def MCMC(whichx, nsamp):
 
@@ -128,6 +130,7 @@ def make_plots(whichx):
     m, c, sig = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))
     pars = [m[0], c[0], sig[0]]
+    print pars
 
     plt.clf()
     plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0, ecolor=".7")

### 2015-06-03 code/plots.py
changing greek letters

@@ -27,10 +27,11 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     if fname == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = cols.blue
-        plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} (\\alpha + \\beta F_8, \sigma)$")
+        plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
+                 (\\alpha + \\beta F_8, \sigma)$")
         plt.text(1.7, .2, "$\\alpha = %.3f$" % (alpha-2))
         plt.text(1.7, .05, "$\\beta = %.3f$" % beta)
-        plt.text(1.7, -.1, "$\sigma = %.3f$" % tau**.5)
+        plt.text(1.7, -.1, "$\\gamma = %.3f$" % tau**.5)
         plt.fill_between(ys, ((ys-alpha)/beta)-sigma-3,
                          ((ys-alpha)/beta)+sigma-3,
                          color=col, alpha=.3, edgecolor="w")
@@ -48,15 +49,17 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
         plt.ylim(3, 5)
         col = cols.pink
         plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
-        plt.text(1.7, 4.7, "$\log(g) \sim \mathcal{N} (\\alpha + \\beta F_8, \sigma)$")
-        plt.text(1.7, 4.5, "$\\alpha = %.3f$" % alpha)
-        plt.text(1.7, 4.4, "$\\beta = %.3f$" % beta)
-        plt.text(1.7, 4.3, "$\sigma = %.3f$" % tau**.5)
+        plt.text(1.7, 4.7, "$\log(g) \sim \mathcal{N} \
+                 (\\delta + \\epsilon F_8, \zeta)$")
+        plt.text(1.7, 4.5, "$\\delta = %.3f$" % alpha)
+        plt.text(1.7, 4.4, "$\\epsilon = %.3f$" % beta)
+        plt.text(1.7, 4.3, "$\\zeta = %.3f$" % tau**.5)
 
         plt.fill_between(ys, ((ys-alpha)/beta)-sigma, ((ys-alpha)/beta)+sigma,
[…]

### 2015-06-03 version1.0/HBM.tex
adding words and references

@@
 Function (PDF) samples, in order to conduct their (hierarchical)
-analyses \citep[e.g.][]{Rogers2014, Foreman-Mackey2014, Angus2015}.
+analyses \citep[e.g.][]{rogers:2014, foreman-Mackey:2014, angus:2015}.
 Including observational uncertainties on the independent variable within our
@@
 \begin{equation}
-	\log(g) \sim \mathcal{N}(\mu = \alpha + \beta F_8, \sigma^2 = \
-	\sigma^2 + \gamma F_8)
+	\rho_* \sim \mathcal{N}(\mu = \alpha + \beta F_8, \sigma^2 = \
+	\sigma^2 + \gamma F_8),
+\end{equation}
+\label{eq:rho}
+and
+\begin{equation}
+	\log_{10}(g) \sim \mathcal{N}(\mu = \delta + \epsilon F_8, \sigma^2 = \
+	\sigma^2 + \zeta F_8)
 \end{equation}
 \label{eq:logg}
-and
-\begin{equation}
-	\rho_* \sim \mathcal{N}(\mu = \delta + \epsilon F_8, \sigma^2 = \
-	\sigma^2 + \zeta F_8),
-\end{equation}
-\label{eq:rho}
 where $F_8$ is flicker and $\rho_*$ is stellar density.
@@
 dependent variable.
+
+In order to account for the two-dimensional observational uncertainties in
+this data set, we marginalize over the latent, `true' values of flicker,
+stellar density and $\log(g)$
+\citep[see, e.g.][]{hogg:2010, foreman-mackey:2014, rogers:2015, angus:2015}.
+The observed values of the variables, flicker, stellar density and $\log(g)$
+can be thought of as a single `draw' from an underlying (assumed Gaussian)
+probability distribution, with mean equal to the `true' value and standard
+deviation equal to the quoted uncertainty.
+We compute the likelihood of the data given the model, marginalized over the
+`true' values of the variables, using importance sampling.
+We sample from the PDF of the `truth', given the observed value and compute
+the likelihood of each of those samples.
+This PDF is actually a {\it posterior} PDF, produced by the astronomers who
+provide the catalog.
+For example, the $log(g)$ values used here are point estimates of the
+posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
+{\it Kepler} light curves and stellar models.
+We would use the posterior samples generated in previous model fitting process
+if they were available, however generating new samples is an acceptable
+approximation provided the posteriors are Gaussian (or log-Normal) and the
+priors used by the previous fitters were uninformative.
+Such an approximation {\it will not be necessary} for those studies using
+the stellar densities or $\log(g)$s calculated using newly-calibrated relations
+presented here, as long as the posterior PDF samples are published!
+After producing new posterior PDF samples we then add up the individual log-likelihoods of each sample to compute the total marginalized likelihood for a star.
+The marginalized likelihood of the whole dataset is then computed as the sum
+of the log-likelihoods of each star.
 

### 2015-06-03 version1.0/manuscript.tex
ruth adds words
adding words and references

@@
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
+Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
+450, 1787
 \bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
@@
 124
+\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
+{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
+ApJ, 795, 64
+\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
+Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
+2166
 \bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
@@
 Buchhave, L.~A., 2014, ApJ, 785, 32
+\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
+Rogers, L.~A., 2015, ApJ, 801, 41
 \bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}

### 2015-06-05 code/main.py
going back to simple sigma

@@ -12,17 +12,20 @@ def lnprior(pars, mm=False):
     return 0.
 
 def lnpriorHF(pars):
+#     m = np.exp(pars[2])
+#     m = pars[2]
+    if pars[2] < 0:
+        return -np.inf
     return 0
 
 def lnprob(pars, samples, obs, u, mm=False):
-    pars[2] = np.exp(pars[2])
-    return lnlikeHF(pars, samples, obs, u) + lnpriorHF(pars)
+    return lnlikeH(pars, samples, obs, u) + lnpriorHF(pars)
 
 def MCMC(whichx, nsamp):
 
     rho_pars = [-2., 6., .0065]
-#     logg_pars = [-1.850, 7., .0065]
-    logg_pars = [-1.850, 7., np.log(.0065)]  # multiplicative
+    logg_pars = [-1.850, 7., .0065, .01]
+#     logg_pars = [-1.850, 7., np.log(.0065)]  # multiplicative
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -56,7 +59,7 @@ def MCMC(whichx, nsamp):
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
 #     nd = len(fr)
-    nd = 20
+    nd = 50
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
@@ -66,7 +69,7 @@ def MCMC(whichx, nsamp):
     u = np.vstack((xerr, yerr))
     up = np.vstack((xerr, yerr))
     um = np.vstack((xerr*.5, yerr*.5))
-    s = generate_samples_log(obs, up, um, 500)
+    s = generate_samples_log(obs, up, um, nsamp)
 
     # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
@@ -79,6 +82,9 @@ def MCMC(whichx, nsamp):
     print "production run..."
     sampler.run_mcmc(pos, 1000)
     samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
+#     m, c, sig, b = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
[…]

### 2015-06-05 code/pgm.py
save as png file

@@ -36,4 +36,4 @@ pgm.add_edge("tau", "f")
 
 # Render and save.
 pgm.render()
-pgm.figure.savefig("../figs/pgm.pdf", dpi=150)
+pgm.figure.savefig("../figs/pgm", dpi=150)

### 2015-06-05 version1.0/HBM.tex
adding words

@@
 \citep[see, e.g.][]{hogg:2010, foreman-mackey:2014, rogers:2015, angus:2015}.
-The observed values of the variables, flicker, stellar density and $\log(g)$
+The observed values of the variables; flicker, stellar density and $\log(g)$
 can be thought of as a single `draw' from an underlying (assumed Gaussian)
 probability distribution, with mean equal to the `true' value and standard
-deviation equal to the quoted uncertainty.
+deviation equal to the observational uncertainty.
 We compute the likelihood of the data given the model, marginalized over the
 `true' values of the variables, using importance sampling.
-We sample from the PDF of the `truth', given the observed value and compute
-the likelihood of each of those samples.
-This PDF is actually a {\it posterior} PDF, produced by the astronomers who
-provide the catalog.
-For example, the $log(g)$ values used here are point estimates of the
-posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
-{\it Kepler} light curves and stellar models.
+We sample from the posterior PDFs of the `true', latent variable values,
+conditioned on the observed values and compute the likelihood of each of those
+samples.
+% These posterior PDFs could be produced by the astronomers who provide the
+% catalog.
+% If posterior samples are made available, they can be used in this stage of
+% the inference.
+% In the event that they are not provided however, they can be generated.
+% For example, the $log(g)$ values used here are point estimates of the
+% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
+% {\it Kepler} light curves and stellar models.
 We would use the posterior samples generated in previous model fitting process
@@
 \caption{
-TBD
+Stellar density vs. flicker. This figure shows the model, conditioned on the
+data. The solid black line shows the median model values, the dark blue
+region shows the width of the intrinsic scatter in the relation; the $\sigma$
+paramete. The faint blue region shows twice the width (the 2$\sigma$ region).
 }
@@
 \caption{
-TBD
+$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
+the model, conditioned on the data. The solid black line shows the median model
+values, the dark pink region shows the width of the intrinsic scatter in the
+relation; the $\sigma$ parameter. The faint pink region shows twice the width
+(the 2$\sigma$ region).
 }

### 2015-06-05 version1.0/discussion.tex
adding words

%% DISCUSSION
%%

We have recalibrated the relation between short timescale brightness
fluctuations in the {\it Kepler} light curves of stars (flicker) with both
stellar density and surface gravity.
This is a simple `fitting a line to data' exercise, however it continues the
discussion of probabilistic modelling that is an active topic within the
fields of exoplanet and stellar astronomy.
We use Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
scatter in the relationship between flicker, surface gravity and density.
We also include the effects of the non-negligable two-dimensional observational
uncertainties by marginalizing over latent variables using importance sampling.
Although none of these methods, nor the data used here are new taken alone,
we use the example of flicker to demonstrate the ideas and the importance of
probabilistic, hierarchical inference.
Relationships between astronomical parameters are almost always
non-deterministic; an element of stochasticity effects the physical parameters
of stars so one can never perfectly predict $y$ given an observation of $x$.
We advocate a probabilistic approach in both the `fitting the model to data'
step, {\it and} when using an empirically calibrated model to predict
parameter values.
The fitting stage benefits because if the relationships between parameters are
falsely assumed to be deterministic, they will be skewed by data points with
unrepresentative uncertainties.
The prediction stage benefits from the stochastic treatment both because a
probability distribution is in many ways more representative of an observation
than a point estimate, and because posterior PDF samples can be used in
subsequent studies (provided the prior used during the fitting process is
described).

We provide posterior PDF samples and the code used in this project at
\url{https://github.com/RuthAngus/flicker}.
Whenever a prediction for the surface gravity or density of a star is required,
for a given estimate of flicker, we recommend using these posterior samples
within the calculation of $\rho_{*}$ or $\log(g)$ and its (Monte Carlo)
uncertainty.
These posterior samples will naturally fold-in the covariances between
parameters.
Simple analytical uncertainty propagation is only valid when uncertainties are
Gaussian and uncorrelated which is rarely true and certainly not the case when
the model is a straight line (the slope and intercept are correlated by
definition).
A flicker value with uncertainties (or even better: posterior PDF
samples), input into our model will result in a probability distribution over
stellar densities or surface gravities which reflects both the uncertainties
on the flicker measurement, the uncertainties on the model parameters {\it and}
the intrinsic scatter in the flicker-$\rho_{*}$-$\log(g)$ relations.

### 2015-06-05 version1.0/introduction.tex
adding words

@@
 Accurate stellar characterization plays a vital role for many active research
-fields within astronomy. For example, stellar populations, galactic archaeology, 
-the study of binary stars, asteroseismology and exoplanet studies all rely on 
-inferences of basic stellar parameters to varying degrees. Empirically-derived
-and reliable estimates are of particular value, increasing our confidence in the
-end-product results built upon these inputs.
+fields within astronomy.
+For example, stellar populations, galactic archaeology, the study of binary
+stars, asteroseismology and exoplanet studies all rely on inferences of basic
+stellar parameters to varying degrees.
+Empirically-derived and reliable estimates are of particular value for
+increasing our confidence in the end-product results built upon these inputs.
 
-Basic stellar parameters, such as effective temperature and surface gravity, 
+Basic stellar parameters such as effective temperature and surface gravity
 can be inferred using one (or more) of several types of observations, e.g.
-spectroscopy, photometry, interferometry, etc. This inference can be done by
-invoking theoretical models or by building an empirical calibration library.
-As an example, an observed stellar spectrum could be matched against a library of 
-theoretical spectra generated using stellar atmosphere models, or, against a 
+spectroscopy, photometry, interferometry, etc.
+This inference can be performed using theoretical models or by building an
+empirical calibration library.
+For example, an observed stellar spectrum could be matched against a library of
+theoretical spectra generated using stellar atmosphere models, or against a
 library of observed spectra of ``standard stars'', serving as calibrators.
@@
 for the inference of stellar parameters are traditionally ``deterministic''.
-In this context, a deterministic model can be loosely described as one where
-a particular observational input always returns a single-valued output for a 
-parameter of interest. In this framework, nature itself has no variance and
-our underlying model is considered to be a perfect description of reality.
+A deterministic model can be loosely described as one where a particular
+observational input always returns a single-valued output for a parameter of
+interest.
+In this framework, nature itself has no variance and our underlying model is
+considered to be a perfect description of reality.
 
-An alternative approach for inferring model parameters is the 
-``probabilistic'' one. Unlike the deterministic case, a single observational 
-input is intepretted to be caused by a range of possible model parameters
-described by a probability distribution. This statement is true even in the case
-of a perfect observation of infinite signal-to-noise, since our underlying 
-model itself comes with uncertainty. In recent years, there has been a shift 
-towards such methods in several areas of astronomy, particularly within the 
-exoplanet community. For example, \citet{wolfgang:2015} considered that
-the mass-radius relationship of exoplanets is probabilistic, since a particular
-sized planet could be due to a range of planet masses, thanks to variances in
-compositions, environment and other circumstances. These recent demonstrations
-in exoplanetary science have prompted us to consider the need for treating the 
-parent stars in the same probabilistic framework, with potential applications 
-spanning many fields of astronomy.
+An alternative approach for inferring model parameters is a ``probabilistic''
+one.
+Unlike the deterministic case, a single observational input is intepreted to be
+caused by a range of possible model parameters, described by a probability
+distribution.
+This statement is true even in the case of a perfect observation of infinite
+signal-to-noise, since our underlying model itself comes with uncertainty.
+In recent years there has been a shift towards such methods in several areas of
+ astronomy and particularly within the exoplanet community.
+For example, \citet{wolfgang:2015} considere the mass-radius relationship
+of exoplanets to be probabilistic, since a planet of a particular radius could
+have a range of masses thanks to variances in compositions, environment and
+other circumstances.
+These recent demonstrations in exoplanetary science prompt us to consider the
+need for treating the parent stars in the same probabilistic framework, with
+potential applications spanning many fields of astronomy.
 
@@
 that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, but also... (ruth will continue this sentence)
+about astrophysical parameters, it also has a practical purpose.
+When using data published in the astronomical literature to, for example, infer
+relationships between observed parameters, that inference can be performed as
+the final stage in a hierarchical treatment.
+This is useful if one wants to account for multi-dimensional uncertainties by
+marginalizing over the `true' parameter values upon which the noisy
+observations are dependent.
+This point is discussed further in \textsection\ref{sec:HBM}.
 

### 2015-06-05 version1.0/manuscript.tex
adding words

@@
 
-\altaffiltext{1}{Dept. of Physics, University of Oxford, UK}
+\altaffiltext{1}{Dept. of Physics, University of Oxford, UK; email:
+		 ruth.angus@astro.ox.ac.uk}
 
@@
 
-TBD
+% aims
+The relations between observable stellar parameters are usually assumed to be
+deterministic, that is, given an infinitely precise measurement of independant
+variable, `$x$', and some model, the value of dependant variable, `y' can be
+known exactly.
+In practise this assumption is rarely valid and intrinsic stochasticity means
+that two stars with exactly the same `$x$', will have slightly different
+`$y$`s.
+% methods
+The relation between short-timescale brightness fluctuations (flicker) of stars
+and both stellar density and surface gravity \citep{bastien:2013} are two such
+stochastic relations.
+We recalibrate these relations in a probabilistic framework, using
+Hierarchical Bayesian Modelling (HBM) and constrain the instrinsic scatter in
+the relations.
+Our treatment also fully accounts for the observational uncertainties on all
+variables.
+% results
+These new relations provide probability distributions over values of $\log(g)$
+or $\rho_*$ for given measurements of flicker.
+Such probability distributions are both more representative of what we know
+about stars and will be useful for subsequent studies using probabilistic
+inference.
 

### 2015-06-05 code/plots.py
switching to draws

@@ -6,6 +6,7 @@ cols = plot_colours()
 import scipy.interpolate as spi
 import h5py
 import sys
+from noisy_plane import model1
 
 def interp(xold, yold, xnew, s):
     tck = spi.splrep(xold, yold, s=s)
@@ -15,8 +16,10 @@ def interp(xold, yold, xnew, s):
 def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     assert np.shape(samples)[0] < np.shape(samples)[1], \
             "samples is wrong shape"
-    alpha, beta, tau = np.median(samples, axis=1)
-    sigma = np.sqrt(tau)
+    beta, alpha, tau = np.median(samples, axis=1)
+    sigma = abs(tau)
+    pars = [beta, alpha, sigma]
+
     print alpha, beta, tau, sigma
     ys = alpha + beta * x
     sig_tot = sigma + yerr
@@ -26,28 +29,28 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
 
     if fname == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
-        col = cols.blue
+        col = cols.pink
         plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
                  (\\alpha + \\beta F_8, \sigma)$")
         plt.text(1.7, .2, "$\\alpha = %.3f$" % (alpha-2))
         plt.text(1.7, .05, "$\\beta = %.3f$" % beta)
         plt.text(1.7, -.1, "$\\gamma = %.3f$" % tau**.5)
-        plt.fill_between(ys, ((ys-alpha)/beta)-sigma-3,
-                         ((ys-alpha)/beta)+sigma-3,
-                         color=col, alpha=.3, edgecolor="w")
-        plt.fill_between(ys, ((ys-alpha)/beta)-2*sigma-3,
-                         ((ys-alpha)/beta)+2*sigma-3,
-                         color=col, alpha=.2, edgecolor="w")
-        dx = x/xerr
-        plt.errorbar(y, x-3, xerr=yerr, yerr=xerr, fmt="k.", capsize=0,
-                     alpha=.5, ecolor=".5", mec=".2")
-        plt.plot(ys, (ys-alpha)/beta-3, ".2", linewidth=1)
-        plt.ylim(-2, 1)
[…]

### 2015-06-05 code/plots.py
remember to subtract 3

@@ -24,39 +24,32 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     ys = alpha + beta * x
     sig_tot = sigma + yerr
     ys = np.linspace(min(y)-5, max(y)+5, 1000)
-    xs = np.linspace(min(x), max(x), 1000)
+    xs = np.linspace(1., 2.4, 100)
     plt.clf()
 
     if fname == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = cols.pink
         plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
-                 (\\alpha + \\beta F_8, \sigma)$")
+                 (\\alpha + \\beta F_8, \sigma_{\rho})$")
         plt.text(1.7, .2, "$\\alpha = %.3f$" % (alpha-2))
         plt.text(1.7, .05, "$\\beta = %.3f$" % beta)
-        plt.text(1.7, -.1, "$\\gamma = %.3f$" % tau**.5)
+        plt.text(1.7, -.1, "$\\sigma_{\rho} = %.3f$" % tau**.5)
 #         plt.ylim(-2, 1)
-
-#         plt.fill_between(xs, ((xs-alpha)/beta)-sigma-3,
-#                          ((xs-alpha)/beta)+sigma-3,
-#                          color=col, alpha=.3, edgecolor="w")
-#         plt.fill_between(xs, ((xs-alpha)/beta)-2*sigma-3,
-#                          ((xs-alpha)/beta)+2*sigma-3,
-#                          color=col, alpha=.2, edgecolor="w")
-#         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
-#                      alpha=.5, ecolor=".5", mec=".2")
-#         plt.plot(ys, (ys-alpha)/beta-3, ".2", linewidth=1)
-#         plt.subplots_adjust(bottom=.1)
+        plt.errorbar(y, x-3, xerr=yerr, yerr=xerr, fmt="k.", capsize=0,
+                             alpha=.5, ecolor=".5", mec=".2")
 
     elif fname == "logg":
         plt.ylim(3, 5)
         col = cols.blue
         plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
[…]

### 2015-06-05 code/main.py
getting ready to run on cluster

@@ -24,7 +24,8 @@ def lnprob(pars, samples, obs, u, mm=False):
 def MCMC(whichx, nsamp):
 
     rho_pars = [-2., 6., .0065]
-    logg_pars = [-1.850, 7., .0065, .01]
+#     logg_pars = [-1.850, 7., .0065, .01]
+    logg_pars = [-1.850, 7., .0065]
 #     logg_pars = [-1.850, 7., np.log(.0065)]  # multiplicative
     pars_init = logg_pars
     if whichx == "rho":
@@ -58,8 +59,8 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-#     nd = len(fr)
-    nd = 50
+    nd = len(fr)
+#     nd = 50
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
@@ -155,11 +156,12 @@ def make_plots(whichx):
 #         plt.plot(x, (y + p2s[i] + p3s[i] * y), "k", alpha=.1)
         plt.plot(x, (y + p2s[i]), "k", alpha=.1)
     plt.savefig("mcmc_%s" % whichx)
-    labels = ["$m$", "$c$", "$\sigma$", "$b$"]
+#     labels = ["$m$", "$c$", "$\sigma$", "$b$"]
+    labels = ["$m$", "$c$", "$\sigma$"]
     fig = triangle.corner(samp, labels=labels)
     fig.savefig("triangle_%s" % whichx)
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    MCMC(whichx, 5)
+    MCMC(whichx, 500)
     make_plots(whichx)

### 2015-06-05 code/plots.py
plotting posterior samples and adjusting labels

@@ -27,36 +27,40 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     xs = np.linspace(1., 2.4, 100)
     plt.clf()
 
+    ndraws = 3000
+    b_samp = np.random.choice(samples[0, :], ndraws)
+    a_samp = np.random.choice(samples[1, :], ndraws)
+    s_samp = np.random.choice(samples[2, :], ndraws) * np.random.randn(ndraws)
+
     if fname == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = cols.pink
-        plt.text(1.7, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
-                 (\\alpha + \\beta F_8, \sigma_{\rho})$")
-        plt.text(1.7, .2, "$\\alpha = %.3f$" % (alpha-2))
-        plt.text(1.7, .05, "$\\beta = %.3f$" % beta)
-        plt.text(1.7, -.1, "$\\sigma_{\rho} = %.3f$" % tau**.5)
-#         plt.ylim(-2, 1)
-        plt.errorbar(y, x-3, xerr=yerr, yerr=xerr, fmt="k.", capsize=0,
+        plt.text(1.55, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
+                 (\\alpha + \\beta \log_{10}(F_8), \sigma_{\\rho})$")
+        plt.text(1.95, .22, "$\\alpha = %.3f$" % (alpha-3))
+        plt.text(1.95, .07, "$\\beta = %.3f$" % beta)
+        plt.text(1.95, -.08, "$\\sigma_{\\rho} = %.3f$" % tau**.5)
+        plt.ylim(-2, 1)
+        for i in range(ndraws):
+            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)-3, col, alpha=.01)
+        plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
+        plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
 
     elif fname == "logg":
         plt.ylim(3, 5)
         col = cols.blue
         plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
[…]

### 2015-06-05 code/noisy_plane.py
adding lhf and model

@@ -1,0 +1,130 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from model import model, model1
+
+def generate_samples(obs, u, N):
+    '''
+    obs is a 2darray of observations. shape = (ndims, nobs)
+    u in an ndarray of the uncertainties associated with the observations
+    returns a 3d array of samples. shape = (nobs, ndims, nsamples)
+    '''
+    ndims, nobs = np.shape(obs)
+    samples = np.zeros((ndims, nobs, N))
+    for i in range(ndims):
+        samples[i, :, :] = np.vstack([x0+xe*np.random.randn(N) for x0, xe in \
+                zip(obs[i, :], u[i, :])])
+    return samples
+
+def lnlike(pars, samples, obs, u):
+    '''
+    Generic likelihood function for importance sampling with any number of
+    dimensions.
+    In samples, obs and u, the dependent variable should be 1st
+    '''
+    ndims, nobs, nsamp = samples.shape
+    zpred = model(pars, samples)
+    zobs = obs[1, :]
+    zerr = u[1, :]
+    ll = np.zeros((nobs, nsamp*nobs))
+    for i in range(nobs):
+        ll[i, :] = -.5*((zobs[i] - zpred)/zerr[i])**2
+    loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
+    return loglike
+#     return np.logaddexp.reduce(loglike, axis=0)
+
+# n-D, hierarchical log-likelihood
+def lnlikeH(pars, samples, obs, u):
+    '''
+    Generic likelihood function for importance sampling with any number of
+    dimensions.
+    Now with added jitter parameter (hierarchical)
+    obs should be a 2d array of observations. shape = (ndims, nobs)
+    u should be a 2d array of uncertainties. shape = (ndims, nobs)
+    samples is a 3d array of samples. shape = (ndims, nobs, nsamp)
+    '''
+    ndims, nobs, nsamp = samples.shape
+    zpred = model(pars, samples)
+    zobs = obs[1, :]
+    zerr = u[1, :]
+    ll = np.zeros((nobs, nsamp*nobs))
+    for i in range(nobs):
+        inv_sigma2 = 1.0/(zerr[i]**2 + pars[2]**2)
+        ll[i, :] = -.5*((zobs[i] - zpred)**2*inv_sigma2) + np.log(inv_sigma2)
[…]

### 2015-06-05 version1.0/HBM.tex
making some corrections to the text

--- 2015-06-02 version1.0/HBM.tex
@@
 
-We model the stochastic relationship between flicker, log($g$) and stellar
-density, accounting for the fact that there exists some intrinsic scatter in
+We model the stochastic relationship between $F_8$, \logg\ and \rhostar,
+accounting for the fact that there exists some intrinsic scatter in
 the dependent variable, and including the heteroskedastic uncertainties on both
@@
 Function (PDF) samples, in order to conduct their (hierarchical)
-analyses \citep[e.g.][]{Rogers2014, Foreman-Mackey2014, Angus2015}.
+analyses \citep[e.g.][]{rogers:2015, foreman-mackey:2014, angus:2015}.
+
 Including observational uncertainties on the independent variable within our
@@
 (e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
-or a demonstration of the effects of neglecting intrinsic scatter and
+For a demonstration of the effects of neglecting intrinsic scatter and
 two-dimensional uncertainties, see Kelly (2007).
 
-The two models describing the relationship between flicker, logg and stellar
+The two models describing the relationship between $F_8$, \logg and \rhostar
 density can be written as
 \begin{equation}
-	\log(g) \sim \mathcal{N}(\mu = \alpha + \beta F_8, \sigma^2 = \
-	\sigma^2 + \gamma F_8)
+	\log_{10}(\rho_\star) \sim \mathcal{N}(\mu = \alpha + \beta
+	\log_{10}(F_8), \sigma = \sigma_{\rho}),
+\end{equation}
+\label{eq:rho}
+and
+\begin{equation}
+	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
+	\sigma = \sigma_g)
 \end{equation}
 \label{eq:logg}
-and
-\begin{equation}
-	\rho_* \sim \mathcal{N}(\mu = \delta + \epsilon F_8, \sigma^2 = \
-	\sigma^2 + \zeta F_8),
-\end{equation}
-\label{eq:rho}
-where $F_8$ is flicker and $\rho_*$ is stellar density.
-The free parameters of the two models are $\alpha$, $\beta$, $\sigma_G$,
-$\gamma$, $\delta$, $\epsilon$ and $\zeta$.
-The two relationships can be described as Gaussian distributions with means
-given by the equation of a straight line, and standard deviations given by
-observational uncertainties, plus some dispersion which is a function of the
-dependent variable.
+The free parameters of the two models are $\alpha$, $\beta$, $\sigma_{\rho}$,
+$\gamma$, $\delta$, and $\sigma_g$.
+These relationships can be described as Gaussian distributions with means
+given by the equation of a straight line, and standard deviations which
+describe the intrinsic scatter about the mean.
+
+In order to account for the two-dimensional observational uncertainties in
+this data set, we marginalize over the latent, `true' values of $F_8$,
+\rhostar and \logg using the importance sampling method of
+\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
+\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
+likelhood, we expect any bias introduced by the importance sampling method to
+be negligable.}.
+The observed values of the variables; $F_8$, \rhostar and \logg can be thought
+of as a single `draw' from an underlying (assumed Gaussian) probability
+distribution, with mean equal to the `true' value\footnote{We use `true' to
+mean the value that would have been observed, given infinitely high
+signal-to-noise.} and standard deviation equal to the observational
+uncertainty.
+We compute the likelihood of the data given the model, marginalized over the
+true values, using importance sampling.
+We sample from the posterior PDFs of the true values of the variables,
+conditioned on the observed values, and compute the likelihood of each of those
+samples.
+% These posterior PDFs could be produced by the astronomers who provide the
+% catalog.
+% If posterior samples are made available, they can be used in this stage of
+% the inference.
+% In the event that they are not provided however, they can be generated.
+% For example, the $log(g)$ values used here are point estimates of the
+% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
+% {\it Kepler} light curves and stellar models.
+We would use the posterior samples generated in previous model fitting process
+if they were available, however generating new samples is an acceptable
+approximation provided the posteriors are Gaussian (or log-Normal) and the
+priors used by the previous fitters were uninformative.
+Such an approximation {\it will not be necessary} for those studies using
+\rhostar or \logg calculated using the newly-calibrated relations
+presented here, as long as the posterior samples are published.
+After producing new posterior PDF samples we then add up the individual
+log-likelihoods of each sample to compute the total marginalized likelihood
+for a star.
+The marginalized likelihood of the whole dataset is then computed as the sum
+of the log-likelihoods of each star.
+The importance sampling method allows us to approximate the integral over the
+latent variables.
+
+Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
+model.
+The faint colored lines show the model where the parameters have been drawn
+from their posterior PDFs.
 
@@
 \caption{
-TBD
+Stellar density vs. flicker. This figure shows the model, conditioned on the
+data. The solid black line shows the median model values, and the pink lines
+are random draws from the posterior PDFs of the model parameters; $\alpha$,
+$\beta$ and $\sigma_{\rho}$.
 }
@@
 \caption{
-TBD
+$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
+the model, conditioned on the data. The solid black line shows the median model
+values, and the blue lines are random draws from the posterior PDFs of the
+model parameters; $\gamma$, $\delta$ and $\sigma_g$.
 }
@@
 \end{figure}
+
+\begin{table}
+\caption{Median parameter values with their 16th and 84th percentile
+	 uncertainties.}
+\begin{tabular}{lcc}
+\hline\hline
+    Parameter & Median value \\
+    \hline
+    $\alpha$         &   \\
+    $\beta$          &   \\
+    $\sigma_{\rho}$  &   \\
+    $\gamma$         &   \\
+    $\delta$         &   \\
+    $\sigma_{g}$     &   \\
+    \hline
+\end{tabular}
+\end{table}

### 2015-06-05 version1.0/introduction.tex
making some corrections to the text

@@
 stellar parameters to varying degrees.
-Empirically-derived and reliable estimates are of particular value for
-increasing our confidence in the end-product results built upon these inputs.
+% Empirically-derived and reliable estimates are of particular value for
+% increasing our confidence in the end-product results built upon these inputs.
 
@@
  astronomy and particularly within the exoplanet community.
-For example, \citet{wolfgang:2015} considere the mass-radius relationship
+For example, \citet{wolfgang:2015} considered the mass-radius relationship
 of exoplanets to be probabilistic, since a planet of a particular radius could
@@
 
-The demand for probilistic stellar parameters is not only motivated by the fact
-that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, it also has a practical purpose.
+The demand for probabilistic stellar parameters is not only motivated by the
+fact that probability distributions are far more representative of our
+`beliefs' about astrophysical parameters, it also has a practical purpose.
 When using data published in the astronomical literature to, for example, infer

### 2015-06-05 version1.0/manuscript.tex
making some corrections to the text

--- 2015-06-03 version1.0/manuscript.tex
@@
 
-TBD
+% aims
+The relations between observable stellar parameters are usually assumed to be
+deterministic.
+That is, given an infinitely precise measurement of independant
+variable, `$x$', and some model, the value of dependant variable, `y' can be
+known exactly.
+In practise this assumption is rarely valid and intrinsic stochasticity means
+that two stars with exactly the same `$x$', will have slightly different
+`$y$`s.
+% methods
+The relation between short-timescale brightness fluctuations (flicker) of stars
+and both surface gravity \citep{bastien:2013} and stellar density
+\citep{kipping:2014} are two such stochastic relations that have, until now,
+been treated as deterministic ones.
+We recalibrate these relations in a probabilistic framework, using
+Hierarchical Bayesian Modelling (HBM) and constrain the instrinsic scatter in
+the relations.
+Our treatment also fully accounts for the observational uncertainties on all
+variables.
+% results
+These new relations provide probability distributions over values of $\log(g)$
+or \rhostar for given measurements of flicker.
+Such probability distributions will be useful for subsequent studies using
+probabilistic inference.
 
@@
 displays a remarkable correlation to the asteroseismically determined parent
-star surface gravities (\logg). Turning this round, the observation implies
+star surface gravities (\logg). Turning this around, the observation implies
 that flicker can be used to blindly infer surface gravities at the level of
@@
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
+Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
 \bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}

### 2015-06-08 code/plots.py
plotting sqrt

@@ -17,7 +17,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     assert np.shape(samples)[0] < np.shape(samples)[1], \
             "samples is wrong shape"
     beta, alpha, tau = np.median(samples, axis=1)
-    sigma = abs(tau)
+    sigma = (np.sqrt(abs(tau))
     pars = [beta, alpha, sigma]
 
     print alpha, beta, tau, sigma
@@ -46,6 +46,10 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
+        plt.plot(xs, model1(pars, xs)+sigma-3, "k--")
+        plt.plot(xs, model1(pars, xs)-sigma-3, "k--")
+        plt.axhline(sigma, color="k")
+        plt.axhline(-sigma, color="k")
 
     elif fname == "logg":
         plt.ylim(3, 5)
@@ -61,6 +65,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
         plt.plot(xs, model1(pars, xs), ".2", linewidth=1)
         plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0,
                      alpha=.5, ecolor=".5", mec=".2")
+        plt.plot(xs, model1(pars, xs)+sigma, "k--")
+        plt.plot(xs, model1(pars, xs)-sigma, "k--")
 
 #     plt.fill_between(xs, model1(pars, xs)+sigma,
 #                      model1(pars, xs)+sigma, color=col, alpha=.3,

### 2015-06-08 code/main.py
small changes

@@ -59,8 +59,8 @@ def MCMC(whichx, nsamp):
     # load data
     fr, frerr, r, rerr = np.genfromtxt("../data/flickers.dat").T
     fl, flerr, l, lerr, t, terr = np.genfromtxt("../data/log.dat").T
-    nd = len(fr)
-#     nd = 50
+#     nd = len(fr)
+    nd = 100
     x, xerr, y, yerr = fl[:nd], flerr[:nd], l[:nd], lerr[:nd]
     if whichx == "rho":
         x, xerr, y, yerr = fr[:nd], frerr[:nd], r[:nd], rerr[:nd]
@@ -163,5 +163,5 @@ def make_plots(whichx):
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    MCMC(whichx, 500)
+    MCMC(whichx, 10)
     make_plots(whichx)

### 2015-06-08 code/plots.py
fixing typo

@@ -17,7 +17,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
     assert np.shape(samples)[0] < np.shape(samples)[1], \
             "samples is wrong shape"
     beta, alpha, tau = np.median(samples, axis=1)
-    sigma = (np.sqrt(abs(tau))
+    sigma = (np.sqrt(abs(tau)))
     pars = [beta, alpha, sigma]
 
     print alpha, beta, tau, sigma

### 2015-06-08 code/plots.py
plotting on bbq

@@ -1,8 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import triangle
-from colours import plot_colours
-cols = plot_colours()
 import scipy.interpolate as spi
 import h5py
 import sys
@@ -34,7 +32,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
 
     if fname == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
-        col = cols.pink
+        col = "#FF33CC"
         plt.text(1.55, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
                  (\\alpha + \\beta \log_{10}(F_8), \sigma_{\\rho})$")
         plt.text(1.95, .22, "$\\alpha = %.3f$" % (alpha-3))
@@ -48,12 +46,10 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
                              alpha=.5, ecolor=".5", mec=".2")
         plt.plot(xs, model1(pars, xs)+sigma-3, "k--")
         plt.plot(xs, model1(pars, xs)-sigma-3, "k--")
-        plt.axhline(sigma, color="k")
-        plt.axhline(-sigma, color="k")
 
     elif fname == "logg":
         plt.ylim(3, 5)
-        col = cols.blue
+        col = "#0066CC"
         plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
         plt.text(1.6, 4.7, "$\log(g) \sim \mathcal{N} \
                  (\\gamma + \\delta \log_{10}(F_8), \\sigma_g)$")
@@ -79,8 +75,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
 
     plt.xlim(1, 2.4)
     plt.xlabel("$\log_{10}\mathrm{(F}_8~\mathrm{[ppm]})$")
-    print "/Users/angusr/Dropbox/Flickerhackday/figs/%s_vs_flicker.pdf" % fname
-    plt.savefig("/Users/angusr/Dropbox/Flickerhackday/figs/%s_vs_flicker.pdf"
+    print "..figs/%s_vs_flicker.pdf" % fname
+    plt.savefig("../figs/%s_vs_flicker.pdf"
                 % fname)
     plt.savefig("flicker_inv_%s" % fname)
 

### 2015-06-08 version1.0/HBM.tex
a couple more small corrections
ruth's changes

@@
 
-For our calibration of \rhostar-$F_8$, we used a sample of Kepler stars with
+For our calibration data, we used a sample of \Kepler\ stars with
 both asteroseismic and flicker measurements available. \citet{chaplin:2014}
 report asteroseismic \rhostar\ estimates (and the associated uncertainties) for
-518 Kepler stars. The authors report three different sets of results, depending
-on the choice of \Teff\ and \FeH, and in this work we elected to use values
-reported in their Table 6 over Table 5, and Table 5 over Table 4. We
+518 \Kepler\ stars. The authors report three different sets of results, 
+depending on the choice of \Teff\ and \FeH, and in this work we elected to use 
+values reported in their Table 6 over Table 5, and Table 5 over Table 4. We
 additionally used the 71 additional planet hosting stars with asteroseismology
 reported in \citet{huber:2013} but not reported in \citet{chaplin:2014}. Values
-for flicker and range were taken from \citet{kipping:2014}, based upon the
+for flicker and ``range'' were taken from \citet{kipping:2014}, based upon the
 methods described in \citet{bastien:2013}. Following \citet{kipping:2014} and
@@
 masses and radii, this could only be done under the incorrect assumption of
-zero covariance between $M_{\star}$ and $R_{\star}$.}
+zero covariance between $M_{\star}$ and $R_{\star}$.}.
 
@@
 
-We model the stochastic relationship between $F_8$, \logg\ and \rhostar,
+We model the stochastic relationship between $F_8$ and \logg\ or \rhostar,
 accounting for the fact that there exists some intrinsic scatter in
@@
 
-The two models describing the relationship between $F_8$, \logg and \rhostar
+The two models describing the relationship between $F_8$, \logg\ and \rhostar\
 density can be written as
@@
 this data set, we marginalize over the latent, `true' values of $F_8$,
-\rhostar and \logg using the importance sampling method of
+\rhostar\ and \logg\ using the importance sampling method of
 \citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
@@
 be negligable.}.
-The observed values of the variables; $F_8$, \rhostar and \logg can be thought
+In what follows we give an overview of this procedure, but encourage the reader
+to refer back  to this original reference for a more detailed description of
+the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be thought
 of as a single `draw' from an underlying (assumed Gaussian) probability
@@
 Such an approximation {\it will not be necessary} for those studies using
-\rhostar or \logg calculated using the newly-calibrated relations
+\rhostar\ or \logg\ calculated using the newly-calibrated relations
 presented here, as long as the posterior samples are published.

### 2015-06-08 version1.0/introduction.tex
ruth's changes

--- 2015-05-12 version1.0/introduction.tex
@@
 %%
-
-% Stellar characterization techniaues overview, traditionally frequentist
-
-% who cares about characterization?
 
@@
 Basic stellar parameters, such as effective temperature and surface gravity, 
-can be inferred using one (or more) of several types of observations, e.g. 
+can be inferred using one (or more) of several types of observations, such as
 spectroscopy, photometry, interferometry, etc. This inference can be done by
 invoking theoretical models or by building an empirical calibration library.
-As an example, an observed stellar spectrum could be matched against a library of 
+For example, an observed stellar spectrum could be matched against a library of 
 theoretical spectra generated using stellar atmosphere models, or, against a 
@@
 the mass-radius relationship of exoplanets is probabilistic, since a particular
-sized planet could be due to a range of planet masses, thanks to variances in
-compositions, environment and other circumstances. These recent demonstrations
-in exoplanetary science have prompted us to consider the need for treating the 
-parent stars in the same probabilistic framework, with potential applications 
-spanning many fields of astronomy.
+sized planet could be due to a range of planet masses, cdue to unmodeled 
+variances in compositions, environment and other complications. These recent 
+demonstrations in exoplanetary science have prompted us to consider the need for 
+treating the parent stars in the same probabilistic framework, with potential 
+applications spanning many fields of astronomy.
 
+\textbf{
 The demand for probilistic stellar parameters is not only motivated by the fact
@@
 about astrophysical parameters, but also... (ruth will continue this sentence)
+}
 
@@
 %relation for sub-Neptunes and infer the level of astrophysical dispersion
+
+% RA text...
+
+One of the more recent tools developed to characterize stars is known as
+``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
+8-hour timescale (denoted as $F_8$) in a broad visible bandpass time series
+photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
+mission. A more detailed account of the proceedure to calculate flicker is
+described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
+displays a remarkable correlation to the asteroseismically determined parent
+star surface gravities (\logg). Turning this around, the observation implies
+that flicker can be used to blindly infer surface gravities at the level of
+$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
+curves available through the array of exoplanet transit missions flying and
+scheduled to launch.
+
+\citet{cranmer:2014} demonstrated that models of stellar surface granulation are 
+indeed reproduce a flicker effect in close agreement with that observed by
+\citet{bastien:2013}, providing a physically-plausible explanation. Since
+surface gravity is highly correlated to mean stellar density (\rhostar) on
+evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
+be used to infer \rhostar, which is more useful for exoplanet transit
+analysis \citep{seager:2003}.
+
+Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
+of the problem which are attractive for our purposes of a simple demonstration
+of probabilistic inference of stellar parameters. Firstly, in log-log space the
+relationship is very simple, appearing to be linear \citep{kipping:2014}. 
+Secondly, there is a sufficiently large number of points in the sample (439 
+stars) to constrain a population-based model. Thirdly, there is significant 
+excess scatter around the best-fitting relation implying that a deterministic 
+model is inadequate. This is not surprising given that granulation is a complex 
+and messy process for which one should not expect any parametric model to 
+provide a perfect description. For these reasons, we identify the calibration of 
+flicker to \logg\ and \rhostar\ as a well-posed problem to first demonstrate 
+probabilistic inference in the arena of stellar characterization.

### 2015-06-08 version1.0/manuscript.tex
adding text
ruth's changes

@@
 \author{
-	{\bf	R.~Angus\altaffilmark{1,2},
-		D.~M.~Kipping\altaffilmark{2,3},
+	{\bf	Ruth~Angus\altaffilmark{1,2},
+		David.~M.~Kipping\altaffilmark{2,3},
 	}
@@
 % results
-These new relations provide probability distributions over values of $\log(g)$
-or \rhostar for given measurements of flicker.
+These new relations provide probability distributions over values of \logg\
+or \rhostar\ for given measurements of flicker.
 Such probability distributions will be useful for subsequent studies using
@@
 
-One of the more recent tools developed to characterize stars is known as
-``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
-8-hour timescale (dubbed $F_8$) in a broad visible bandpass time series
-photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
-mission. A more detailed account of the proceedure to calculate flicker is
-described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
-displays a remarkable correlation to the asteroseismically determined parent
-star surface gravities (\logg). Turning this around, the observation implies
-that flicker can be used to blindly infer surface gravities at the level of
-$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
-curves available through the array of exoplanet transit missions flying and
-scheduled to launch.
-
-\citet{cranmer:2014} demonstrated that stellar surface granulation is expected
-to reproduce a flicker effect in close agreement with that observed by
-\citet{bastien:2013}, providing a physically-plausible explanation. Since
-surface gravity is highly correlated to mean stellar density (\rhostar) on
-evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
-be used to infer \rhostar, which is more useful for exoplanet transit
-analysis.
-
-Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
-of the problem which are attractive for our purposes of a simple demonstration
-of probabilistic inference of stellar parameters. Firstly, in log-log space the
-relationship is very simple, appearing to be linear. Secondly, there is a
-sufficiently number of points in the sample (439 stars) to constrain a
-population-based model. Thirdly, there is significant excess scatter around the
-best-fitting relation implying that a deterministic model is inadequate. This is
-not surprising given that granulation is a complex and messy process for which
-one should not expect any parametric model to provide a perfect description. For
-these reasons, we identify the calibration of flicker to \logg\ and \rhostar\
-as a well-posed problem to first demonstrate probabilistic inference in the
-arena of stellar characterization.
-
 \section{PROBABILISTIC CALIBRATION}
@@
 \input{HBM.tex}
-
-\section{APPLICATION}
-\label{sec:appication}
-
-\input{application.tex}
 
@@
 Rogers, L.~A., 2015, ApJ, 801, 41
+\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]{seager:2003} 
+Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
 \bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}

### 2015-06-08 code/noisy_plane.py

@@ -15,6 +15,17 @@ def generate_samples(obs, u, N):
                 zip(obs[i, :], u[i, :])])
     return samples
 
+def vanilla_lnlike(pars, obs, u):
+    '''
+    Generic likelihood function
+    '''
+    yobs, xobs = obs[0, :], obs[1, :]
+    yerr, xerr = u[0, :], u[1, :]
+    ypred = model1(pars, xobs)
+    inv_sigma2 = 1./(yerr**2 + pars[2]**2)
+    chi2 = -.5*(((yobs - ypred)**2 * inv_sigma2) - np.log(inv_sigma2))
+    return np.sum(chi2)
+
 def lnlike(pars, samples, obs, u):
     '''
     Generic likelihood function for importance sampling with any number of
@@ -30,7 +41,6 @@ def lnlike(pars, samples, obs, u):
         ll[i, :] = -.5*((zobs[i] - zpred)/zerr[i])**2
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     return loglike
-#     return np.logaddexp.reduce(loglike, axis=0)
 
 # n-D, hierarchical log-likelihood
 def lnlikeH(pars, samples, obs, u):

### 2015-06-08 version1.0/introduction.tex
test
v minor changes

@@
 %%
+
+TEST
 
@@
 
-\citet{cranmer:2014} demonstrated that models of stellar surface granulation are 
+\citet{cranmer:2014} demonstrated that models of stellar surface granulation 
 indeed reproduce a flicker effect in close agreement with that observed by

### 2015-06-08 code/plots.py
including scatter

@@ -57,7 +57,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, plot_samp=False):
         plt.text(1.95, 4.42, "$\\delta = %.3f$" % beta)
         plt.text(1.95, 4.32, "$\\sigma_g = %.3f$" % tau**.5)
         for i in range(ndraws):
-            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs), col, alpha=.01)
+            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)+s_samp, col, alpha=.01)
         plt.plot(xs, model1(pars, xs), ".2", linewidth=1)
         plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0,
                      alpha=.5, ecolor=".5", mec=".2")

### 2015-07-09 code/main.py
using bigdata

@@ -19,7 +19,7 @@ def lnpriorHF(pars):
 def lnprob(pars, samples, obs, u, mm=False):
     return lnlikeH(pars, samples, obs, u) + lnpriorHF(pars)
 
-def MCMC(whichx, nsamp, fname, nd, bigdata):
+def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
 
     rho_pars = [-2., 6., .0065]
     logg_pars = [-1.850, 7., .0065]
@@ -27,16 +27,7 @@ def MCMC(whichx, nsamp, fname, nd, bigdata):
     if whichx == "rho":
         pars_init = rho_pars
 
-    plt.clf()
-    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=False)
-    plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="r.", capsize=0)
-    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=bigdata)
-    plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0)
-    plt.ylim(3, 5)
-    plt.xlim(1, 2.4)
-    plt.savefig("Test.pdf")
-
-    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=False)
+    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=True)
 
     # format data and generate samples
     obs = np.vstack((x, y))
@@ -51,10 +42,10 @@ def MCMC(whichx, nsamp, fname, nd, bigdata):
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
                                     args=(s, obs, u))
     print "burning in..."
-    pos, _, _, = sampler.run_mcmc(pos, 500)
+    pos, _, _, = sampler.run_mcmc(pos, burnin)
     sampler.reset()
     print "production run..."
-    sampler.run_mcmc(pos, 1000)
+    sampler.run_mcmc(pos, run)
     samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
     m, c, sig = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))
@@ -98,7 +89,7 @@ def make_plots(whichx, fname):
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
     fname = "test"
[…]

### 2015-07-09 code/main.py
adding priors

@@ -9,7 +9,9 @@ import sys
 import h5py
 
 def lnprior(pars, mm=False):
-    return 0.
+    if -10 < pars[0] < 10 and -10 < pars[1] < 10 and 0 < pars[2] < 1:
+        return 0.
+    return -np.inf
 
 def lnpriorHF(pars):
     if pars[2] < 0:
@@ -34,7 +36,13 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
     u = np.vstack((xerr, yerr))
     up = np.vstack((xerr, yerr))
     um = np.vstack((xerr*.5, yerr*.5))
-    s = generate_samples_log(obs, up, um, nsamp)
+#     s = generate_samples_log(obs, up, um, nsamp) # FIXME
+    s = generate_samples(obs, u, nsamp) # FIXME
+#     if nsamp == 1:
+#         s[0, :, :] = x
+#         s[1, :, :] = y
+#     print np.shape(s)
+#     assert 0
 
     # set up and run emcee
     ndim, nwalkers = len(pars_init), 32
@@ -88,8 +96,8 @@ def make_plots(whichx, fname):
 
 if __name__ == "__main__":
     whichx = str(sys.argv[1])
-    fname = "test"
-    MCMC(whichx, 50, fname, 0, bigdata=True)
+    fname = "f"
+    MCMC(whichx, 10, fname, 0, bigdata=True)
     make_plots(whichx, fname)
 
 #     # load data

### 2015-07-09 code/main.py
using fractional lhf

@@ -19,7 +19,7 @@ def lnpriorHF(pars):
     return 0
 
 def lnprob(pars, samples, obs, u, mm=False):
-    return lnlikeH(pars, samples, obs, u) + lnpriorHF(pars)
+    return lnlikeHF(pars, samples, obs, u) + lnpriorHF(pars)
 
 def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
 

### 2015-07-09 code/plots.py
adding fractional option

@@ -20,7 +20,7 @@ def fit_straight_line(x, y, yerr):
     return np.linalg.solve(ATCA, ATCy)
 
 def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
-                              plot_samp=False):
+                              plot_samp=False, fractional=True):
 
     # fit straight line
     lim = 200
@@ -54,7 +54,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
         plt.text(1.95, -.08, "$\\sigma_{\\rho} = %.3f$" % tau**.5)
         plt.ylim(-2, 1)
         for i in range(ndraws):
-            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)-3, col, alpha=.01)
+            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)-3, col, alpha=.1)
         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
@@ -71,23 +71,32 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
         plt.text(1.95, 4.42, "$\\delta = %.3f$" % beta)
         plt.text(1.95, 4.32, "$\\sigma_g = %.3f$" % tau**.5)
         for i in range(ndraws):
-            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)+s_samp[i], col, alpha=.01)
+            ys = model1([b_samp[i], a_samp[i]], xs)
+            if fractional:
+                plt.plot(xs, ys + ys * s_samp[i], col, alpha=.05)
+            else:
+                plt.plot(xs, ys + s_samp[i], col, alpha=.05)
         plt.plot(xs, model1(pars, xs), ".2", linewidth=1)
         plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0,
-                     alpha=.5, ecolor="k", mec=".2")
-        plt.plot(xs, model1(pars, xs)+sigma, "k--")
-        plt.plot(xs, model1(pars, xs)-sigma, "k--")
[…]

### 2015-07-09 code/model.py
changing axis limits

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import scipy as sp
 import scipy.optimize as spo
 
-def load_data(whichx, nd=0, bigdata=False, huber=False):
+def load_data(whichx, nd=0, bigdata=True, huber=False):
 
     if bigdata:
         data = np.genfromtxt("../data/BIGDATA.nohuber.filtered.dat").T
@@ -52,10 +52,6 @@ def load_data(whichx, nd=0, bigdata=False, huber=False):
 def model1(pars, x):
     return pars[0]*x + pars[1]
 
-def deming_model(pars, x, y):
-    b, a = pars
-    return (a*b + x + b*y) / (1 + b**2), (a + b*(x + b*y)) / (1 + b**2)
-
 def model(pars, samples):
     xs = samples[0, :, :]
     x = np.reshape(xs, np.shape(xs)[0]* np.shape(xs)[1])
@@ -86,39 +82,29 @@ def nll(pars, x, y, yerr, model):
     chi2 = -.5 * ((y-mod)**2 * inv_sig2)
     return - np.logaddexp.reduce(chi2, axis=0)
 
-def deming_nll(pars, x, y, yerr, model):
-    inv_sig2 = 1./(yerr**2 + xerr**2)
-    xc, yc = deming_model(pars, x, y)
-    chi2 = ((x-xc)**2 + (y-yc)**2) * inv_sig2
-    return - np.logaddexp.reduce(chi2, axis=0)
-
 if __name__ == "__main__":
 
-    x, y, xerr, yerr = load_data("logg", bigdata=True)
+    x, y, xerr, yerr = load_data("logg")
 
     pars_init = [1.8, 5.6, -1, 5.95, -1.2]
     res = spo.minimize(nll, pars_init, args=(x, y, yerr, broken_power_law1))
     print res.x
 
     plt.clf()
-    plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0, ecolor=".7")
+    plt.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="k.", capsize=0, ecolor=".7",
+                 label="bigdata")
 
-#     plt.xlim(1.6, 2.2)
-#     plt.plot(x, broken_power_law1(res.x, x), "m")
+    plt.xlim(1., 2.4)
+    plt.ylim(3., 5.)
+    plt.plot(x, broken_power_law1(res.x, x), "m")
 
     pars_init = [5.6, -1]
[…]

### 2015-07-09 code/noisy_plane.py
fractional sigma

@@ -80,9 +80,7 @@ def lnlikeHF(pars, samples, obs, u):
     yerr = u[1, :]
     ll = np.zeros((nobs, nsamp*nobs))
     for i in range(nobs):
-        inv_sigma2 = 1.0/(yerr[i]**2 + pars[2]**2 + pars[3]*model1(pars, xobs[i]))
-#         inv_sigma2 = 1.0/(yerr[i]**2 + (m*model1(pars, xobs[i]))**2)
-#         inv_sigma2 = 1.0/(yerr[i]**2 + m/model1(pars, xobs[i]))
+        inv_sigma2 = 1.0/(yerr[i]**2 + (pars[2]*model1(pars, xobs[i]))**2)
         ll[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     if np.isfinite(loglike):

### 2015-07-09 code/plots.py
almost finished new plots

@@ -54,7 +54,11 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
         plt.text(1.95, -.08, "$\\sigma_{\\rho} = %.3f$" % tau**.5)
         plt.ylim(-2, 1)
         for i in range(ndraws):
-            plt.plot(xs, model1([b_samp[i], a_samp[i]], xs)-3, col, alpha=.1)
+            ys = model1([b_samp[i], a_samp[i]], xs)
+            if fractional:
+                plt.plot(xs, ys + ys * s_samp[i], col, alpha=.05)
+            else:
+                plt.plot(xs, ys + s_samp[i], col, alpha=.05)
         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")

### 2015-07-10 code/noisy_plane.py
adding extra intercept

@@ -64,7 +64,7 @@ def lnlikeH(pars, samples, obs, u):
     return loglike
 
 # n-D, hierarchical log-likelihood with sigma as a function of y
-def lnlikeHF(pars, samples, obs, u):
+def lnlikeHF(pars, samples, obs, u, extra=False):
     '''
     Generic likelihood function for importance sampling with any number of
     dimensions.
@@ -72,6 +72,7 @@ def lnlikeHF(pars, samples, obs, u):
     obs should be a 2d array of observations. shape = (ndims, nobs)
     u should be a 2d array of uncertainties. shape = (ndims, nobs)
     samples is a 3d array of samples. shape = (ndims, nobs, nsamp)
+    if extra == True, the sigma has both a slope and intercept
     '''
     ndims, nobs, nsamp = samples.shape
     ypred = model(pars, samples)
@@ -80,7 +81,11 @@ def lnlikeHF(pars, samples, obs, u):
     yerr = u[1, :]
     ll = np.zeros((nobs, nsamp*nobs))
     for i in range(nobs):
-        inv_sigma2 = 1.0/(yerr[i]**2 + (pars[2]*model1(pars, xobs[i]))**2)
+        if extra:
+            inv_sigma2 = 1.0/(yerr[i]**2 + \
+                    (pars[2] + pars[3] * model1(pars, xobs[i]))**2)
+        else:
+            inv_sigma2 = 1.0/(yerr[i]**2 + (pars[2]*model1(pars, xobs[i]))**2)
         ll[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     if np.isfinite(loglike):

### 2015-07-10 code/extra_main.py
adding extra_main.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from noisy_plane import generate_samples, generate_samples_log, \
+        lnlike, lnlikeH, lnlikeHF
+from model import model1, model, load_data
+import emcee
+import triangle
+import sys
+import h5py
+
+def lnprior(pars, mm=False):
+    if -10 < pars[0] < 10 and -10 < pars[1] < 10 and -10 < pars[2] < 10 \
+            and -10 < pars[3] < -10:
+        return 0.
+    return -np.inf
+
+def lnprob(pars, samples, obs, u, mm=False):
+    return lnlike(pars, samples, obs, u) + lnprior(pars)
+
+def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
+
+    rho_pars = [-2., 6., .0065, .05]
+    logg_pars = [-1.850, 7., .0065, .05]
+    pars_init = logg_pars
+    if whichx == "rho":
+        pars_init = rho_pars
+
+    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=True)
+
+    # format data and generate samples
+    obs = np.vstack((x, y))
+    u = np.vstack((xerr, yerr))
+    up = np.vstack((xerr, yerr))
+    um = np.vstack((xerr*.5, yerr*.5))
+    s = generate_samples(obs, u, nsamp) # FIXME
+
+    # set up and run emcee
+    ndim, nwalkers = len(pars_init), 32
+    pos = [pars_init + 1e-4*np.random.randn(ndim) for i in range(nwalkers)]
+    sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
+                                    args=(s, obs, u))
+    print "burning in..."
+    pos, _, _, = sampler.run_mcmc(pos, burnin)
+    sampler.reset()
+    print "production run..."
+    sampler.run_mcmc(pos, run)
+    samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
+    m, c, sig, f = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
+               zip(*np.percentile(samp, [16, 50, 84], axis=0)))
+    pars = [m[0], c[0], sig[0], f[0]]
+
+    # save samples
+    f = h5py.File("%s_samples_%s.h5" % (whichx, fname), "w")
[…]

### 2015-07-10 code/noisy_plane.py
adding mixture model

@@ -92,6 +92,42 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
         return loglike
     return -np.inf
 
+# n-D, hierarchical log-likelihood with sigma as a function of y
+def lnlikeHFM(pars, samples, obs, u, extra=False):
+    '''
+    Generic likelihood function for importance sampling with any number of
+    dimensions.
+    Now with added jitter parameter (hierarchical)
+    obs should be a 2d array of observations. shape = (ndims, nobs)
+    u should be a 2d array of uncertainties. shape = (ndims, nobs)
+    samples is a 3d array of samples. shape = (ndims, nobs, nsamp)
+    if extra == True, the sigma has both a slope and intercept
+    Now with a mixture model!
+    '''
+    ndims, nobs, nsamp = samples.shape
+    ypred = model(pars, samples)
+    yobs = obs[1, :]
+    xobs = obs[0, :]
+    yerr = u[1, :]
+    ll1 = np.zeros((nobs, nsamp*nobs))
+    ll2 = np.zeros((nobs, nsamp*nobs))
+    Y, V, P = pars[4], pars[5], pars[6]
+    for i in range(nobs):
+        if extra:
+            inv_sigma2 = 1.0/(yerr[i]**2 + \
+                    (pars[2] + pars[3] * model1(pars, xobs[i]))**2 + V)
+        else:
+            inv_sigma2 = 1.0/(yerr[i]**2 + \
+                    (pars[2]*model1(pars, xobs[i]))**2 + V)
+        ll1[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
+        ll2[i, :] = -.5*((yobs[i] - Y)**2*inv_sigma2) + np.log(inv_sigma2)
+    loglike1 = np.logaddexp.reduce(ll, axis=1)
+    loglike2 = np.logaddexp.reduce(ll, axis=1)
+    loglike = np.sum(np.logaddexp(np.log(1-P) + lnlike1, np.log(P) + lnlike2))
+    if np.isfinite(loglike):
+        return loglike
+    return -np.inf
+
 def generate_samples_log(obs, up, um, N):
     '''
     obs is a 2darray of observations. shape = (ndims, nobs)

### 2015-07-10 version1.0/manuscript.tex
v minor changes
Make link, citation, etc. colors not terrible
Please!

@@
 \usepackage{epsfig}
+\usepackage{color}
+\definecolor{linkcolor}{rgb}{0,0,0.5}
 \usepackage[%
 	colorlinks=true,	% false: boxed links; true: colored links
-	urlcolor=blue,		% color of external links
 	pdfpagelabels=true,
@@
 %	pdfnewwindow=true,      % links in new window
-	linkcolor=red,          % color of internal links (change box color with linkbordercolor)
-	citecolor=green,        % color of links to bibliography
-%	filecolor=magenta,      % color of file links
+	urlcolor=linkcolor,     % color of external links
+	linkcolor=linkcolor,    % color of internal links (change box color with linkbordercolor)
+	citecolor=linkcolor,    % color of links to bibliography
+	filecolor=linkcolor,    % color of file links
         ]{hyperref}
@@
 %%
-Based on archival data of the \emph{Kepler} telescope.
+Based on archival data of the \Kepler\ telescope.
 }

### 2015-07-10 version1.0/introduction.tex
changed data files

@@
 %%
-
-TEST
 
@@
 
-\textbf{
 The demand for probilistic stellar parameters is not only motivated by the fact
 that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, but also... (ruth will continue this sentence)
-}
+about astrophysical parameters, is also has a practical purpose.
+When using data published in the astronomical literature to, for example, infer
+relationships between observed parameters, that inference can be performed as
+the final stage in a hierarchical treatment.
+This is useful if one wants to account for multi-dimensional uncertainties by
+marginalizing over the `true' parameter values upon which the noisy
+observations are dependent.
+This point is discussed further in \textsection\ref{sec:HBM}.
 

### 2015-07-10 version1.0/manuscript.tex
changed data files

--- 2015-06-08 version1.0/manuscript.tex
@@
 that two stars with exactly the same `$x$', will have slightly different
-`$y$`s.
+`$y$'s.
 % methods
@@
 % results
-These new relations provide probability distributions over values of \logg\
-or \rhostar\ for given measurements of flicker.
-Such probability distributions will be useful for subsequent studies using
-probabilistic inference.
+These new relations will provide probability distributions over values of
+\logg\ or \rhostar\ for given measurements of flicker.
 

### 2015-07-10 code/extra.py
running extra slope and intercept version and mixture model version

@@ -15,7 +15,7 @@ def lnprior(pars, mm=False):
     return -np.inf
 
 def lnprob(pars, samples, obs, u, mm=False):
-    return lnlike(pars, samples, obs, u) + lnprior(pars)
+    return lnlikeHF(pars, samples, obs, u, extra=True) + lnprior(pars)
 
 def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
 

### 2015-07-10 code/main.py

@@ -71,7 +71,7 @@ def make_plots(whichx, fname):
 
     x, y, xerr, yerr = load_data(whichx)
 
-    with h5py.File("%s_samples.h5" % whichx) as f:
+    with h5py.File("%s_samples_%s.h5" % (whichx, fname)) as f:
         samp = f["samples"][...]
     m, c, sig = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))

### 2015-07-10 code/noisy_plane.py

@@ -111,7 +111,7 @@ def lnlikeHFM(pars, samples, obs, u, extra=False):
     yerr = u[1, :]
     ll1 = np.zeros((nobs, nsamp*nobs))
     ll2 = np.zeros((nobs, nsamp*nobs))
-    Y, V, P = pars[4], pars[5], pars[6]
+    Y, V, P = pars[3], pars[4], pars[5]
     for i in range(nobs):
         if extra:
             inv_sigma2 = 1.0/(yerr[i]**2 + \
@@ -121,8 +121,8 @@ def lnlikeHFM(pars, samples, obs, u, extra=False):
                     (pars[2]*model1(pars, xobs[i]))**2 + V)
         ll1[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
         ll2[i, :] = -.5*((yobs[i] - Y)**2*inv_sigma2) + np.log(inv_sigma2)
-    loglike1 = np.logaddexp.reduce(ll, axis=1)
-    loglike2 = np.logaddexp.reduce(ll, axis=1)
+    lnlike1 = np.logaddexp.reduce(ll1, axis=1)
+    lnlike2 = np.logaddexp.reduce(ll2, axis=1)
     loglike = np.sum(np.logaddexp(np.log(1-P) + lnlike1, np.log(P) + lnlike2))
     if np.isfinite(loglike):
         return loglike

### 2015-07-10 code/plots.py

@@ -30,20 +30,18 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
     assert np.shape(samples)[0] < np.shape(samples)[1], \
             "samples is wrong shape"
     beta, alpha, tau = np.median(samples, axis=1)
-    sigma = (np.sqrt(abs(tau)))
+    sigma = abs(tau)**.5
     pars = [beta, alpha, sigma]
 
     print alpha, beta, tau, sigma
-    ys = alpha + beta * x
-    sig_tot = sigma + yerr
-    ys = np.linspace(min(y)-5, max(y)+5, 1000)
-    xs = np.linspace(1., 2.4, 100)
-    plt.clf()
 
     b_samp = np.random.choice(samples[0, :], ndraws)
     a_samp = np.random.choice(samples[1, :], ndraws)
-    s_samp = np.random.choice(samples[2, :], ndraws) * np.random.randn(ndraws)
+    t_samp = np.random.choice(samples[2, :], ndraws)
+    s_samp = abs(t_samp)**.5 * np.random.randn(ndraws)
 
+    plt.clf()
+    xs = np.linspace(1., 2.4, 100)
     if whichx == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = "#FF33CC"
@@ -56,14 +54,19 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, ndraws,
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             if fractional:
-                plt.plot(xs, ys + ys * s_samp[i], col, alpha=.05)
+                plt.plot(xs, ys + ys * s_samp[i] - 3, col, alpha=.05)
             else:
-                plt.plot(xs, ys + s_samp[i], col, alpha=.05)
+                plt.plot(xs, ys + s_samp[i] - 3, col, alpha=.05)
         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
-        plt.plot(xs, model1(pars, xs)+sigma-3, "k--")
-        plt.plot(xs, model1(pars, xs)-sigma-3, "k--")
+        ys = model1(pars, xs)
+        if fractional:
[…]

### 2015-08-24 code/extra.py
no log prior on f

@@ -10,7 +10,7 @@ import h5py
 
 def lnprior(pars, mm=False):
     if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -10 < pars[2] < 10 \
-            and -10 < pars[3] < 10:
+            and -100 < pars[3] < 100:
         return 0.
     return -np.inf
 
@@ -28,8 +28,8 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
     """
 
     # set initial parameters
-    rho_pars = [-2., 6., np.log(.0065), np.log(.05)]
-    logg_pars = [-1.850, 7., np.log(.0065), np.log(.05)]
+    rho_pars = [-2., 6., np.log(.0065), .05]
+    logg_pars = [-1.850, 7., np.log(.0065), .05]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars

### 2015-08-24 code/plots.py
saving likelihoods and plotting regions

@@ -29,12 +29,19 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
 
     assert np.shape(samples)[0] < np.shape(samples)[1], \
             "samples is wrong shape"
+
+    # take medians
     results = np.median(samples, axis=1)
     print 'results = ', results
     beta, alpha, tau = results[:3]
     if extra:
         print np.shape(results), "shape"
         beta, alpha, tau, f = results
+
+    # use highest likelihood samples
+    lls = samples[:, -1]
+    m = lls == max(lls)
+    beta, alpha, tau = [samples[i, m] for i in range(np.shape(samples)[0]-1)]
 #     sigma = abs(tau)**.5
     sigma = tau
     pars = [beta, alpha, sigma]
@@ -60,6 +67,7 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         plt.text(1.95, .07, "$\\beta = %.3f$" % beta)
         plt.text(1.95, -.08, "$\\sigma_{\\rho} = %.3f$" % sigma)
         plt.ylim(-2, 1)
+        lines = []
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             if fractional:
@@ -68,10 +76,18 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
                 plt.plot(xs, f_samp[i] * ys + s_samp[i] - 3, col,
                          alpha=.05)
             else:
-                plt.plot(xs, ys + s_samp[i] - 3, col, alpha=.05)
-#         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
+#                 plt.plot(xs, ys + s_samp[i] - 3, col, alpha=.05)
+#                 plt.plot(xs, ys + np.random.randn(1)*s_samp[i] - 3, col,
+#                          alpha=.05)
+                lines.append(ys + np.random.randn(1)*s_samp[i] - 3) #FIXME: opt
+        plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
+
[…]

### 2015-08-24 code/simple_flicker.py

@@ -25,13 +25,16 @@ sampler = emcee.EnsembleSampler(nwalkers, ndim, lnlike)
 
 p0 = np.array([m_init, b_init, lnsy_init]) \
         + 1e-5 * np.random.randn(nwalkers, ndim)
-p0, _, _, _ = sampler.run_mcmc(p0, 200)
+p0, _, _, _ = sampler.run_mcmc(p0, 5000)
 sampler.reset()
-sampler.run_mcmc(p0, 500)
+sampler.run_mcmc(p0, 5000)
 
 lls = sampler.blobs
 flat_lls = np.reshape(lls, (np.shape(lls)[0]*np.shape(lls)[1]))
 samp = np.vstack((sampler.chain[:, :, :].reshape(-1, ndim).T, flat_lls)).T
+
+triangle.corner(sampler.flatchain, truths=[1.7, 2.013, np.log(.022)])
+plt.savefig("triangle_%s_%s" % (whichx, fname))
 
 f = h5py.File("%s_samples_%s.h5" % (whichx, fname), "w")
 data = f.create_dataset("samples", np.shape(samp))

### 2015-08-24 code/extra.py
saving likelihoods

@@ -11,13 +11,13 @@ import h5py
 def lnprior(pars, mm=False):
     if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -10 < pars[2] < 10 \
             and -100 < pars[3] < 100:
-        return 0.
-    return -np.inf
+        return 0., 0.
+    return -np.inf, None
 
 def lnprob(pars, samples, obs, u):
     return lnlikeHF(pars, samples, obs, u) + lnprior(pars)
 
-def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
+def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     """
     nsamp (int) = number of samples.
     whichx (str) = logg or rho.
@@ -54,10 +54,11 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
     sampler.reset()
     print "production run..."
     sampler.run_mcmc(pos, run)
-    samp = sampler.chain[:, 50:, :].reshape((-1, ndim))
-    m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
-               zip(*np.percentile(samp, [16, 50, 84], axis=0)))
-    pars = np.exp([m[0], c[0], ln_sig[0], lnf[0]])
+
+    # load likelihood
+    lls = sampler.blobs
+    flat_lls = np.reshape(lls, (np.shape(lls)[0]*np.shape(lls)[1]))
+    samp = np.vstack((sampler.chain[:, :, :].reshape(-1, ndim).T, flat_lls)).T
 
     # save samples
     f = h5py.File("%s_samples_%s.h5" % (whichx, fname), "w")
@@ -66,6 +67,7 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=500):
     data[:, 1] = np.exp(samp[:, 1])
     data[:, 2] = np.exp(samp[:, 2])
     data[:, 3] = np.exp(samp[:, 3])
+    data[:, 4] = np.exp(samp[:, 4])
     f.close()
 
 def make_plots(whichx, fname):
@@ -88,5 +90,5 @@ if __name__ == "__main__":
     whichx = str(sys.argv[1])
     fname = "f_extra"
     nd = 0 # set to zero to use all the data
-    MCMC(whichx, 500, fname, nd, bigdata=True)
[…]

### 2015-08-24 code/noisy_plane.py
saving likelihoods

@@ -91,8 +91,8 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
         ll[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     if np.isfinite(loglike):
-        return loglike
-    return -np.inf
+        return loglike, loglike
+    return -np.inf, None
 
 # n-D, hierarchical log-likelihood with sigma as a function of y
 def lnlikeHFM(pars, samples, obs, u, extra=False):

### 2015-08-24 code/extra.py
getting ll samples to work

@@ -50,7 +50,7 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
                                     args=(s, obs, u))
     print "burning in..."
-    pos, _, _, = sampler.run_mcmc(pos, burnin)
+    pos, _, _, _ = sampler.run_mcmc(pos, burnin)
     sampler.reset()
     print "production run..."
     sampler.run_mcmc(pos, run)
@@ -75,7 +75,7 @@ def make_plots(whichx, fname):
     x, y, xerr, yerr = load_data(whichx)
 
     with h5py.File("%s_samples_%s.h5" % (whichx, fname)) as f:
-        samp = f["samples"][...]
+        samp = f["samples"][:, 3]
     m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))
     pars = [m[0], c[0], ln_sig[0], lnf[0]]

### 2015-08-24 code/plots.py

@@ -42,8 +42,9 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
     lls = samples[:, -1]
     m = lls == max(lls)
     beta, alpha, tau = [samples[i, m] for i in range(np.shape(samples)[0]-1)]
-#     sigma = abs(tau)**.5
-    sigma = tau
+    sigma = abs(tau)**.5
+    if fname == "simple":
+        sigma = tau
     pars = [beta, alpha, sigma]
 
     print alpha, beta, tau, sigma
@@ -51,7 +52,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
     b_samp = np.random.choice(samples[0, :], ndraws)
     a_samp = np.random.choice(samples[1, :], ndraws)
     t_samp = np.random.choice(samples[2, :], ndraws)
-#     s_samp = abs(t_samp)**.5 * np.random.randn(ndraws)
+    s_samp = abs(t_samp)**.5 * np.random.randn(ndraws)
+    if fname == "simple"
     s_samp = t_samp
     if extra:
         f_samp = np.random.choice(samples[3, :], ndraws)
@@ -71,10 +73,14 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             if fractional:
-                plt.plot(xs, ys + ys * s_samp[i] - 3, col, alpha=.05)
+#                 plt.plot(xs, ys + ys * s_samp[i] - 3, col, alpha=.05)
+                lines.append(ys + ys * np.random.randn(1)*s_samp[i]
+                             - 3) #FIXME: opt
             elif extra:
-                plt.plot(xs, f_samp[i] * ys + s_samp[i] - 3, col,
-                         alpha=.05)
+#                 plt.plot(xs, f_samp[i] * ys + s_samp[i] - 3, col,
+#                          alpha=.05)
+                lines.append(ys*f_samp[i] + np.random.randn(1)*s_samp[i]
+                             - 3) #FIXME: opt
             else:
 #                 plt.plot(xs, ys + s_samp[i] - 3, col, alpha=.05)
 #                 plt.plot(xs, ys + np.random.randn(1)*s_samp[i] - 3, col,
[…]

### 2015-08-24 code/extra.py
dubbuging plotting

@@ -63,11 +63,11 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     # save samples
     f = h5py.File("%s_samples_%s.h5" % (whichx, fname), "w")
     data = f.create_dataset("samples", np.shape(samp))
-    data[:, 0] = np.exp(samp[:, 0])
-    data[:, 1] = np.exp(samp[:, 1])
+    data[:, 0] = samp[:, 0]
+    data[:, 1] = samp[:, 1]
     data[:, 2] = np.exp(samp[:, 2])
-    data[:, 3] = np.exp(samp[:, 3])
-    data[:, 4] = np.exp(samp[:, 4])
+    data[:, 3] = samp[:, 3]
+    data[:, 4] = samp[:, 4]
     f.close()
 
 def make_plots(whichx, fname):
@@ -75,7 +75,7 @@ def make_plots(whichx, fname):
     x, y, xerr, yerr = load_data(whichx)
 
     with h5py.File("%s_samples_%s.h5" % (whichx, fname)) as f:
-        samp = f["samples"][:, 3]
+        samp = f["samples"][:, :-1]
     m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                zip(*np.percentile(samp, [16, 50, 84], axis=0)))
     pars = [m[0], c[0], ln_sig[0], lnf[0]]
@@ -90,5 +90,5 @@ if __name__ == "__main__":
     whichx = str(sys.argv[1])
     fname = "f_extra"
     nd = 0 # set to zero to use all the data
-    MCMC(whichx, 5, fname, nd, bigdata=True, burnin=100, run=100)
+    MCMC(whichx, 500, fname, nd, bigdata=True)
     make_plots(whichx, fname)

### 2015-08-24 code/plots.py

@@ -91,19 +91,23 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
                              alpha=.5, ecolor=".5", mec=".2")
 
         quantiles = np.percentile(lines, [16, 84], axis=0)
+        diff1 = model1(pars, xs) - 3 - quantiles[0]
+        diff2 = quantiles[1] - model1(pars, xs) - 3
         plt.fill_between(xs, quantiles[0], quantiles[1], color=col,
+                         alpha=.5)
+        plt.fill_between(xs, quantiles[0]-diff1, quantiles[1]+diff2, color=col,
                          alpha=.5)
 
         ys = model1(pars, xs)
-        if fractional:
-            plt.plot(xs, ys + ys * sigma - 3 , "k--")
-            plt.plot(xs, ys - ys * sigma - 3, "k--")
-        elif extra:
-            plt.plot(xs, ys + f * ys + sigma - 3 , "k--")
-            plt.plot(xs, ys - f * ys + sigma - 3, "k--")
-        else:
-            plt.plot(xs, ys + sigma - 3 , "k--")
-            plt.plot(xs, ys - sigma - 3, "k--")
+#         if fractional:
+#             plt.plot(xs, ys + ys * sigma - 3 , "k--")
+#             plt.plot(xs, ys - ys * sigma - 3, "k--")
+#         elif extra:
+#             plt.plot(xs, ys + f * ys + sigma - 3 , "k--")
+#             plt.plot(xs, ys - f * ys + sigma - 3, "k--")
+#         else:
+#             plt.plot(xs, ys + sigma - 3 , "k--")
+#             plt.plot(xs, ys - sigma - 3, "k--")
 
     elif whichx == "logg":
         plt.ylim(3, 5)
@@ -130,25 +134,28 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
 #                 plt.plot(xs, ys + np.random.randn(1)*s_samp[i], col, alpha=.05)
                 lines.append(ys + np.random.randn(1)*s_samp[i]) #FIXME: opt
 
-
-        quantiles = np.percentile(lines, [16, 84], axis=0)
-        plt.fill_between(xs, quantiles[0], quantiles[1], color=col,
-                         alpha=.5)
-
         plt.plot(xs, model1(pars, xs), ".2", linewidth=1)
[…]

### 2015-08-26 code/extra.py
not log

@@ -8,9 +8,14 @@ import triangle
 import sys
 import h5py
 
-def lnprior(pars, mm=False):
+def lnprior_extra(pars, mm=False):
     if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -10 < pars[2] < 10 \
             and -100 < pars[3] < 100:
+        return 0., 0.
+    return -np.inf, None
+
+def lnprior(pars, mm=False):
+    if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -100 < pars[2] < 100:
         return 0., 0.
     return -np.inf, None
 
@@ -28,8 +33,12 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     """
 
     # set initial parameters
-    rho_pars = [-2., 6., np.log(.0065), .05]
-    logg_pars = [-1.850, 7., np.log(.0065), .05]
+    if fname == "f_extra":
+        rho_pars = [-2., 6., np.log(.0065), .05]
+        logg_pars = [-1.850, 7., np.log(.0065), .05]
+    else:
+        rho_pars = [-2., 6., .0065]
+        logg_pars = [-1.850, 7., .0065]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -67,7 +76,8 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     data[:, 1] = samp[:, 1]
     data[:, 2] = np.exp(samp[:, 2])
     data[:, 3] = samp[:, 3]
-    data[:, 4] = samp[:, 4]
+    if fname == "f_extra":
+        data[:, 4] = samp[:, 4]
     f.close()
 
 def make_plots(whichx, fname):
@@ -76,19 +86,27 @@ def make_plots(whichx, fname):
 
     with h5py.File("%s_samples_%s.h5" % (whichx, fname)) as f:
         samp = f["samples"][:, :-1]
-    m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
-               zip(*np.percentile(samp, [16, 50, 84], axis=0)))
-    pars = [m[0], c[0], ln_sig[0], lnf[0]]
+
+    if fname == "f_extra":
+        m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
[…]

### 2015-08-26 code/noisy_plane.py

@@ -87,7 +87,7 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
                     np.exp(pars[3]) * model1(pars, xobs[i]) )**2)
         else:
             inv_sigma2 = 1.0/(yerr[i]**2 + \
-                    (np.exp(pars[2])*model1(pars, xobs[i]))**2)
+                    (pars[2]*model1(pars, xobs[i]))**2)
         ll[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     if np.isfinite(loglike):

### 2015-08-26 code/plots.py

@@ -53,8 +53,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
     a_samp = np.random.choice(samples[1, :], ndraws)
     t_samp = np.random.choice(samples[2, :], ndraws)
     s_samp = abs(t_samp)**.5 * np.random.randn(ndraws)
-    if fname == "simple"
-    s_samp = t_samp
+    if fname == "simple":
+        s_samp = t_samp
     if extra:
         f_samp = np.random.choice(samples[3, :], ndraws)
 
@@ -90,13 +90,11 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
 
-        quantiles = np.percentile(lines, [16, 84], axis=0)
-        diff1 = model1(pars, xs) - 3 - quantiles[0]
-        diff2 = quantiles[1] - model1(pars, xs) - 3
-        plt.fill_between(xs, quantiles[0], quantiles[1], color=col,
-                         alpha=.5)
-        plt.fill_between(xs, quantiles[0]-diff1, quantiles[1]+diff2, color=col,
-                         alpha=.5)
+        quantiles = np.percentile(lines, [2, 16, 84, 98], axis=0)
+        plt.fill_between(xs, quantiles[1], quantiles[2], color=col,
+                         alpha=.4)
+        plt.fill_between(xs, quantiles[0], quantiles[3], color=col,
+                         alpha=.2)
 
         ys = model1(pars, xs)
 #         if fractional:
@@ -139,13 +137,11 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
                      alpha=.5, ecolor=".5", mec=".2")
         ys = model1(pars, xs)
 
-        quantiles = np.percentile(lines, [16, 84], axis=0)
-        plt.fill_between(xs, quantiles[0], quantiles[1], color=col,
-                         alpha=.5)
-        diff1 = model1(pars, xs) - 3 - quantiles[0]
-        diff2 = quantiles[1] - model1(pars, xs) - 3
-        plt.fill_between(xs, quantiles[0]-diff1, quantiles[1]+diff2, color=col,
-                         alpha=.5)
[…]

### 2015-08-26 version1.0/HBM.tex
adding equation

@@
 these probability distributions through to subsequent analyses.
-Several recent studies have required posterior Probability Distribution
+Several recent studies have required posterior Probability Density
 Function (PDF) samples, in order to conduct their (hierarchical)
-analyses \citep[e.g.][]{rogers:2015, foreman-mackey:2014, angus:2015}.
+analyses \citep[e.g.][]{foreman-mackey:2014, rogers:2015, angus:2015}.
 
@@
 that is biased towards zero
-(e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
-For a demonstration of the effects of neglecting intrinsic scatter and
-two-dimensional uncertainties, see Kelly (2007).
+\citep[e.g.][]{fuller:1987, fox:1997}.
+For a demonstration of the affects of neglecting intrinsic scatter and
+two-dimensional uncertainties, see \citet{kelly:2007}.
 
-The two models describing the relationship between $F_8$, \logg\ and \rhostar\
-density can be written as
+We write the two models describing the relationship between $F_8$, \logg\ and
+\rhostar\ as
 \begin{equation}
@@
 	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
-	\sigma = \sigma_g)
+	\sigma = \sigma_g).
 \end{equation}
@@
 \rhostar\ or \logg\ calculated using the newly-calibrated relations
-presented here, as long as the posterior samples are published.
+presented here, since we have published our posterior samples.
 After producing new posterior PDF samples we then add up the individual
@@
 latent variables.
+We used flat priors for all except the $\sigma$ parameters, for which we used
+priors which were flat in log-space.
 
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
-model.
-The faint colored lines show the model where the parameters have been drawn
-from their posterior PDFs.
+models.
+The shaded regions show the 1 and 2$\sigma$ confidence interval.
+These intervals are representative of the uncertainty in the flicker relations,
+but also of the intrinsic scatter in those relations.
+% Given a measurement of flicker, the predictive probability distribution of
+% \logg and \rhostar will be a Gaussian distribution
+
+We also used an analytic form of the marginal likelihood, which accounts for
+2-d uncertainties but does not allow for the modelling of the intrinsic scatter
+as a function of $x$ or $y$.
+For the relation between flicker and \rhostar This likelihood function can be
+written as
+\begin{eqnarray}
+	p(\rho_*|\alpha, \beta, \sigma_{\rho}) = &
+	\exp \left[-\frac{1}{2} \sum_{n=1}^N\frac{[y_n-(\alpha + \beta x_n)]^2}
+	{\left[\beta \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2\right]}\right]
+	\\ \nonumber
+	& + \beta^2 \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2,
+\end{eqnarray}
+and similarly for \logg.
+The parameter values found using this simpler likelihood function are included
+in table \ref{tab:results2}.
 
@@
 \caption{
-Stellar density vs. flicker. This figure shows the model, conditioned on the
-data. The solid black line shows the median model values, and the pink lines
-are random draws from the posterior PDFs of the model parameters; $\alpha$,
-$\beta$ and $\sigma_{\rho}$.
-}
+Stellar density vs. flicker.
+This figure shows the model, conditioned on the data.
+The solid black line shows the highest-likelihood sample, and the pink shaded
+regions show the 1 and 2$\sigma$ confidence intervals.}
 \label{fig:rhostar}
@@
 \caption{
-$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
-the model, conditioned on the data. The solid black line shows the median model
-values, and the blue lines are random draws from the posterior PDFs of the
-model parameters; $\gamma$, $\delta$ and $\sigma_g$.
-}
+$\log(g)$ vs. flicker.
+As in \ref{fig:rhostar} this figure shows the model, conditioned on the data.
+The solid black line shows the highest-likelihood sample, and the blue shaded
+regions show the 1 and 2$\sigma$ confidence intervals.}
 \label{fig:logg}
@@
 \begin{table}
-\caption{Median parameter values with their 16th and 84th percentile
-	 uncertainties.}
+\caption{Highest-likelihood parameter values with their 16th and 84th
+	percentile uncertainties.}
 \begin{tabular}{lcc}
 \hline\hline
-    Parameter & Median value \\
+    Parameter & highest-likelihood value \\
     \hline

### 2015-08-26 version1.0/manuscript.tex
adding equation

@@
 
-RA thanks...
+RA would like to thank Dan Foreman-Mackey for his extremely helpful suggestions
+and comments, and Angie Wolfgang and John Johnson for their fantastic advice.
+RA would also like to thank the SAMSI institute and the other members of the
+SAMSI {\it Kepler} exoplanet statistics working groups.
 DMK thanks...
@@
 ApJ, 795, 64
+\bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}
+Fox, J., 1997, Applied Regression Analysis, Linear Models, and Related Methods
+(Thousand Oaks:Sage Publications, Inc.)
+\bibitem[\protect\citeauthoryear{Fuller}{1987}]{fuller:1987}
+Fuller, W.~A., 1987, Wiley Series in Probability and Mathematical Statistics
 \bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
@@
 127
+\bibitem[\protect\citeauthoryear{Kelly}{2007}]{kelly:2007} Kelly, B.~C., 2007,
+ApJ, 665, 1489
 \bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
@@
 Rogers, L.~A., 2015, ApJ, 801, 41
-\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]{seager:2003}
-Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
+\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]
+{seager:2003}Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
 \bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}

### 2015-08-26 code/noisy_plane.py
speeding up

@@ -75,20 +75,17 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
     if extra == True, the sigma has both a slope and intercept
     '''
     ndims, nobs, nsamp = samples.shape
-    ypred = model(pars, samples)
-    yobs = obs[1, :]
-    xobs = obs[0, :]
-    yerr = u[1, :]
     ll = np.zeros((nobs, nsamp*nobs))
     for i in range(nobs):
         if extra:
-            inv_sigma2 = 1.0/(yerr[i]**2 + \
+            inv_sigma2 = 1.0/(u[1, :][i]**2 + \
                     (np.exp(pars[2]) + \
-                    np.exp(pars[3]) * model1(pars, xobs[i]) )**2)
+                    np.exp(pars[3]) * model1(pars, obs[0, :][i]) )**2)
         else:
-            inv_sigma2 = 1.0/(yerr[i]**2 + \
-                    (pars[2]*model1(pars, xobs[i]))**2)
-        ll[i, :] = -.5*((yobs[i] - ypred)**2*inv_sigma2) + np.log(inv_sigma2)
+            inv_sigma2 = 1.0/(u[1, :][i]**2 + \
+                    (pars[2]*model1(pars, obs[0, :][i]))**2)
+        ll[i, :] = -.5*((obs[1, :][i] - model(pars, samples))**2*inv_sigma2) \
+                + np.log(inv_sigma2)
     loglike = np.sum(np.logaddexp.reduce(ll, axis=1))
     if np.isfinite(loglike):
         return loglike, loglike

### 2015-08-27 code/extra.py
bug fixing, running extra and asym prior

@@ -9,18 +9,18 @@ import sys
 import h5py
 
 def lnprior_extra(pars, mm=False):
-    if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -10 < pars[2] < 10 \
-            and -100 < pars[3] < 100:
+    if -1e6 < pars[0] < 1e6 and -1e6 < pars[1] < 1e6 and -1e6 < pars[2] < 1e6 \
+            and 0 < pars[3] < 1e6:
         return 0., 0.
     return -np.inf, None
 
 def lnprior(pars, mm=False):
-    if -100 < pars[0] < 100 and -100 < pars[1] < 100 and -100 < pars[2] < 100:
+    if -1e6 < pars[0] < 1e6 and -1e6 < pars[1] < 1e6 and -1e6 < pars[2] < 1e6:
         return 0., 0.
     return -np.inf, None
 
 def lnprob(pars, samples, obs, u):
-    return lnlikeHF(pars, samples, obs, u) + lnprior(pars)
+    return lnlikeHF(pars, samples, obs, u, extra=True) + lnprior_extra(pars)
 
 def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
     """
@@ -34,11 +34,11 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
 
     # set initial parameters
     if fname == "f_extra":
-        rho_pars = [-2., 6., np.log(.0065), .05]
-        logg_pars = [-1.850, 7., np.log(.0065), .05]
+        rho_pars = [-1.69293833, 5.1408906, .0065, .05]
+        logg_pars = [-1.05043614, 5.66819525, .0065, .05]
     else:
-        rho_pars = [-2., 6., .0065]
-        logg_pars = [-1.850, 7., .0065]
+        rho_pars = [-1.69293833, 5.1408906, .0065]
+        logg_pars = [-1.05043614, 5.66819525, .0065]
     pars_init = logg_pars
     if whichx == "rho":
         pars_init = rho_pars
@@ -91,7 +91,7 @@ def make_plots(whichx, fname):
         m, c, ln_sig, lnf = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
                    zip(*np.percentile(samp, [16, 50, 84], axis=0)))
         pars = [m[0], c[0], ln_sig[0], lnf[0]]
[…]

### 2015-08-27 code/noisy_plane.py

@@ -79,8 +79,7 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
     for i in range(nobs):
         if extra:
             inv_sigma2 = 1.0/(u[1, :][i]**2 + \
-                    (np.exp(pars[2]) + \
-                    np.exp(pars[3]) * model1(pars, obs[0, :][i]) )**2)
+                    (pars[2] + pars[3] * model1(pars, obs[0, :][i]) )**2)
         else:
             inv_sigma2 = 1.0/(u[1, :][i]**2 + \
                     (pars[2]*model1(pars, obs[0, :][i]))**2)

### 2015-08-27 code/plots.py

@@ -52,7 +52,8 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
     b_samp = np.random.choice(samples[0, :], ndraws)
     a_samp = np.random.choice(samples[1, :], ndraws)
     t_samp = np.random.choice(samples[2, :], ndraws)
-    s_samp = abs(t_samp)**.5 * np.random.randn(ndraws)
+    if fname == "f":
+        s_samp = (abs(t_samp)**.5 - 1) * np.random.randn(ndraws) + 1
     if fname == "simple":
         s_samp = t_samp
     if extra:
@@ -68,14 +69,16 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         plt.text(1.95, .22, "$\\alpha = %.3f$" % (alpha-3))
         plt.text(1.95, .07, "$\\beta = %.3f$" % beta)
         plt.text(1.95, -.08, "$\\sigma_{\\rho} = %.3f$" % sigma)
-        plt.ylim(-2, 1)
+#         plt.ylim(-2, 1)
+
         lines = []
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             if fractional:
-#                 plt.plot(xs, ys + ys * s_samp[i] - 3, col, alpha=.05)
-                lines.append(ys + ys * np.random.randn(1)*s_samp[i]
-                             - 3) #FIXME: opt
+                plt.plot(xs, ys-3 + (ys-3) * (s_samp[i]), col, alpha=.05)
+                plt.plot(xs, ys-3 , "b", alpha=.05)
+#                 lines.append(ys-3 + ys * (np.random.randn(1)*s_samp[i] - 1))
+#                              - 3) #FIXME: opt
             elif extra:
 #                 plt.plot(xs, f_samp[i] * ys + s_samp[i] - 3, col,
 #                          alpha=.05)
@@ -89,14 +92,14 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         plt.plot(xs, model1(pars, xs)-3, ".2", linewidth=1)
         plt.errorbar(x, y-3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
                              alpha=.5, ecolor=".5", mec=".2")
+#
+#         quantiles = np.percentile(lines, [2, 16, 84, 98], axis=0)
[…]

### 2015-08-29 code/extra.py
running for longer

@@ -34,8 +34,8 @@ def MCMC(whichx, nsamp, fname, nd, bigdata, burnin=500, run=1000):
 
     # set initial parameters
     if fname == "f_extra":
-        rho_pars = [-1.69293833, 5.1408906, .0065, .05]
-        logg_pars = [-1.05043614, 5.66819525, .0065, .05]
+        rho_pars = [-1.69293833, 5.1408906, 1., -.001]
+        logg_pars = [-1.05043614, 5.66819525, 1., .06]
     else:
         rho_pars = [-1.69293833, 5.1408906, .0065]
         logg_pars = [-1.05043614, 5.66819525, .0065]
@@ -108,5 +108,5 @@ if __name__ == "__main__":
     whichx = str(sys.argv[1])
     fname = "f_extra"
     nd = 0 # set to zero to use all the data
-    MCMC(whichx, 2, fname, nd, bigdata=True, burnin=10, run=50)
+    MCMC(whichx, 500, fname, nd, bigdata=True, burnin=500, run=1000)
     make_plots(whichx, fname)

### 2015-10-13 code/noisy_plane.py
what the hell was that lhf about?

@@ -81,7 +81,7 @@ def lnlikeHF(pars, samples, obs, u, extra=False):
     for i in range(nobs):
         if extra:
             inv_sigma2 = 1.0/(u[1, :][i]**2 + \
-                    pars[2]**2 + pars[3] * model1(pars, obs[0, :][i]))**2
+                    pars[2]**2 + pars[3] * obs[0, :][i])
         else:
             inv_sigma2 = 1.0/(u[1, :][i]**2 + \
                     (pars[2]*model1(pars, obs[0, :][i]))**2)

### 2015-10-13 code/extra.py
correctifying

@@ -9,8 +9,8 @@ import sys
 import h5py
 
 def lnprior_extra(pars, mm=False):
-    if -1e6 < pars[0] < 0 and 0 < pars[1] < 1e6 and -1e6 < pars[2] < 1e6 \
-            and 0 < pars[3] < 1e6:
+    if -100 < pars[0] < 0 and 0 < pars[1] < 100 and 0 < pars[2] < 100 \
+            and -100 < pars[3] < 100:
         return 0., 0.
     return -np.inf, None
 
@@ -24,6 +24,7 @@ def lnprob(pars, samples, obs, u, extra, f):
         return lnlikeHF(pars, samples, obs, u, extra=extra) + \
                 lnprior_extra(pars)
     elif f:
+        print extra, f
         return lnlikeHF(pars, samples, obs, u, extra=extra) + lnprior(pars)
     else:
         return lnlikeH(pars, samples, obs, u) + lnprior(pars)
@@ -118,6 +119,7 @@ if __name__ == "__main__":
     elif fname == "f": extra, f = False, True
     else: extra, f = False, False
     nd = 0 # set to zero to use all the data
-    ns, bi, r = 50, 100, 500
+#     ns, bi, r = 50, 100, 500
+    ns, bi, r = 2, 50, 200
     MCMC(whichx, ns, fname, nd, extra, f, bigdata=False, burnin=bi, run=r)
     make_plots(whichx, fname)

### 2015-10-13 code/plots.py

@@ -87,12 +87,14 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             y3 = ys - 3
-            if fractional:
-                plt.plot(xs, ym + ym * np.random.randn(1)*sigma - 3 ,
-                        "b", alpha=.05)
+#             if fractional:
+#                 plt.plot(xs, ym + ym * np.random.randn(1)*sigma - 3 ,
+#                         "b", alpha=.05)
 #                 lines.append(ys-3 + ys * (np.random.randn(1)*s_samp[i] - 1))
 #                 lines.append(ym + ym * np.random.randn(1)*(sigma-1) - 3)
-            elif extra:
+            if extra:
+                plt.plot(xs, y3 + np.random.randn(1)*np.median(s_samp) +
+                        np.random.randn(1)*np.median(f_samp)*xs, col, alpha=.05)
 #                 plt.plot(xs, y3 * (1 + f_samp[i]), "k", alpha=.05)
 #                 plt.plot(xs, ys - 3, col, alpha=.05)
 #                 lines.append(y3 * f_samp[i] + s_samp[i]) #FIXME: opt
@@ -136,11 +138,13 @@ def make_inverse_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         ym = model1([np.median(b_samp), np.median(a_samp)], xs)
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
-            if fractional:
-                plt.plot(xs, ym + ym * np.random.randn(1)*sigma - 3 ,
-                        "b", alpha=.05)
+#             if fractional:
+#                 plt.plot(xs, ym + ym * np.random.randn(1)*sigma - 3 ,
+#                         "b", alpha=.05)
 #                 lines.append(ys + ys*np.random.randn(1)*s_samp[i])
-            elif extra:
+            if extra:
+                plt.plot(xs, ys + np.random.randn(1)*np.median(s_samp) +
+                        np.random.randn(1)*np.median(f_samp)*xs, col, alpha=.05)
 #                 plt.plot(xs, ys + np.median(f_samp-1)*np.random.randn(1) * \
 #                         ys + np.median(s_samp-1)*np.random.randn(1), col, alpha=.05)
 #                 plt.plot(xs, ys + np.median(f_samp)*np.random.randn(1)*ys

### 2015-10-13 code/extra.py
new, simpler plots

@@ -9,7 +9,7 @@ import sys
 import h5py
 
 def lnprior_extra(pars, mm=False):
-    if -100 < pars[0] < 0 and 0 < pars[1] < 100 and 0 < pars[2] < 100 \
+    if -100 < pars[0] < 0 and 0 < pars[1] < 100 and -100 < pars[2] < 100 \
             and -100 < pars[3] < 100:
         return 0., 0.
     return -np.inf, None
@@ -51,7 +51,7 @@ def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
         pars_init = rho_pars
 
     # load the data
-    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=True)
+    x, y, xerr, yerr = load_data(whichx, nd=nd, bigdata=False)
 
     # format data and generate samples
     obs = np.vstack((x, y))
@@ -120,6 +120,6 @@ if __name__ == "__main__":
     else: extra, f = False, False
     nd = 0 # set to zero to use all the data
 #     ns, bi, r = 50, 100, 500
-    ns, bi, r = 2, 50, 200
+    ns, bi, r = 2, 50, 500
     MCMC(whichx, ns, fname, nd, extra, f, bigdata=False, burnin=bi, run=r)
     make_plots(whichx, fname)

### 2015-10-13 code/plots.py

@@ -19,6 +19,71 @@ def fit_straight_line(x, y, yerr):
     ATCy = np.dot(np.dot(AT, C), y)
     return np.linalg.solve(ATCA, ATCy)
 
+def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
+                          fractional=False, extra=False):
+
+    # draw samples from post
+    b_samp = np.random.choice(samples[0, :], ndraws)
+    a_samp = np.random.choice(samples[1, :], ndraws)
+    s_samp = np.random.choice(samples[2, :], ndraws)
+    f_samp = np.random.choice(samples[3, :], ndraws)
+
+    # rho plot
+    plt.clf()
+    xs = np.linspace(.7, 2.5, 100)
+    if whichx == "rho":
+        plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
+        col = "#FF33CC"
+        plt.ylim(-2, 1)
+
+        # plot line draws
+        for i in range(ndraws):
+            ys = model1([b_samp[i], a_samp[i]], xs)
+            y3 = ys - 3
+            line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
+                    np.random.randn(1)*np.median(f_samp)*xs)**.5 - 3
+            plt.plot(xs, line, col, alpha=.05)
+
+        # plot best fit and data
+        ym = model1([np.median(b_samp), np.median(a_samp)], xs)
+        plt.plot(xs, ym - 3, ".2", linewidth=1)
+        plt.errorbar(x, y - 3, xerr=xerr, yerr=xerr, fmt="k.", capsize=0,
+                             alpha=.5, ecolor=".5", mec=".2")
+        plt.savefig("new_rho")
+
+    # logg plot
+    elif whichx == "logg":
+        plt.ylim(3, 5)
+        col = "#0066CC"
+        plt.ylabel("$\log_{10}(g [\mathrm{cm~s}^{-2}])$")
+
+        # plot line draws
+#         lines = np.zeros((ndraws, len(xs)))
+        lines = []
+        for i in range(ndraws):
+            ys = model1([b_samp[i], a_samp[i]], xs)
+            line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
+                    np.random.randn(1)*np.median(f_samp)*xs)**.5
+            plt.plot(xs, line, col, alpha=.05)
+#             lines[i, :] = line
+            lines.append(line)
+
+        # plot regions
[…]

### 2015-10-13 code/extra.py
trying to correct priors

@@ -23,9 +23,9 @@ def lnprob(pars, samples, obs, u, extra, f):
     if extra:
         return lnlikeHF(pars, samples, obs, u, extra=extra) + \
                 lnprior_extra(pars)
-    elif f:
-        print extra, f
-        return lnlikeHF(pars, samples, obs, u, extra=extra) + lnprior(pars)
+#     elif f:
+#         print extra, f
+#         return lnlikeHF(pars, samples, obs, u, extra=extra) + lnprior(pars)
     else:
         return lnlikeH(pars, samples, obs, u) + lnprior(pars)
 

### 2015-10-14 code/extra.py
adjusting priors again and changing initialisation

@@ -9,25 +9,31 @@ import sys
 import h5py
 
 def lnprior_extra(pars, mm=False):
-    if -100 < pars[0] < 0 and 0 < pars[1] < 100 and -100 < pars[2] < 100 \
+    if -10 < pars[0] < 0 and 0 < pars[1] < 10 and -100 < pars[2] < 100 \
             and -100 < pars[3] < 100:
         return 0., 0.
     return -np.inf, None
 
-def lnprior(pars, mm=False):
-    if -1e6 < pars[0] < 1e6 and -1e6 < pars[1] < 1e6 and -1e6 < pars[2] < 1e6:
-        return 0., 0.
-    return -np.inf, None
+# def lnprior(pars, mm=False):
+#     if -1e6 < pars[0] < 1e6 and -1e6 < pars[1] < 1e6 and -1e6 < pars[2] < 1e6:
+#         return 0., 0.
+#      return -np.inf, None
 
 def lnprob(pars, samples, obs, u, extra, f):
-    if extra:
-        return lnlikeHF(pars, samples, obs, u, extra=extra) + \
-                lnprior_extra(pars)
+#     if extra:
+    print pars
+    pars[-2] = 1.2
+    print pars
+    print lnlikeHF(pars, samples, obs, u, extra=extra) + \
+            lnprior_extra(pars)
+    raw_input('tetner')
+    return lnlikeHF(pars, samples, obs, u, extra=extra) + \
+            lnprior_extra(pars)
 #     elif f:
 #         print extra, f
 #         return lnlikeHF(pars, samples, obs, u, extra=extra) + lnprior(pars)
-    else:
-        return lnlikeH(pars, samples, obs, u) + lnprior(pars)
+#     else:
+#         return lnlikeH(pars, samples, obs, u) + lnprior(pars)
 
 def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
     """
@@ -41,8 +47,8 @@ def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
 
     # set initial parameters
     if fname == "f_extra":
-        rho_pars = [-1.69293833, 5.1408906, 1., -.001]
-        logg_pars = [-1.05043614, 5.66819525, 1., .06]
+        rho_pars = [-1.69293833, 5.1408906, .25, -.001]
+        logg_pars = [-1.05043614, 5.66819525, .25, .06]
     else:
[…]

### 2015-10-14 code/extra.py
running for a bit longer to check result

@@ -20,13 +20,13 @@ def lnprior_extra(pars, mm=False):
 #      return -np.inf, None
 
 def lnprob(pars, samples, obs, u, extra, f):
-#     if extra:
-    print pars
-    pars[-2] = 1.2
-    print pars
-    print lnlikeHF(pars, samples, obs, u, extra=extra) + \
-            lnprior_extra(pars)
-    raw_input('tetner')
+# #     if extra:
+#     print pars
+#     pars[-2] = 1.2
+#     print pars
+#     print lnlikeHF(pars, samples, obs, u, extra=extra) + \
+#             lnprior_extra(pars)
+#     raw_input('tetner')
     return lnlikeHF(pars, samples, obs, u, extra=extra) + \
             lnprior_extra(pars)
 #     elif f:
@@ -126,6 +126,6 @@ if __name__ == "__main__":
     else: extra, f = False, False
     nd = 0 # set to zero to use all the data
 #     ns, bi, r = 50, 100, 500
-    ns, bi, r = 2, 50, 500
+    ns, bi, r = 2, 50, 1000
     MCMC(whichx, ns, fname, nd, extra, f, bigdata=False, burnin=bi, run=r)
     make_plots(whichx, fname)

### 2015-10-14 code/extra.py
full run with better initialisation

@@ -47,8 +47,8 @@ def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
 
     # set initial parameters
     if fname == "f_extra":
-        rho_pars = [-1.69293833, 5.1408906, .25, -.001]
-        logg_pars = [-1.05043614, 5.66819525, .25, .06]
+        rho_pars = [-1.793214679, 5.34215688, 0.02334097, -.0002600777]
+        logg_pars = [-1.02143776, 5.69156647, .24239756, .049233887]
     else:
         rho_pars = [-1.69293833, 5.1408906, .0065]
         logg_pars = [-1.05043614, 5.66819525, .0065]
@@ -126,6 +126,6 @@ if __name__ == "__main__":
     else: extra, f = False, False
     nd = 0 # set to zero to use all the data
 #     ns, bi, r = 50, 100, 500
-    ns, bi, r = 2, 50, 1000
+    ns, bi, r = 500, 100, 5000
     MCMC(whichx, ns, fname, nd, extra, f, bigdata=False, burnin=bi, run=r)
     make_plots(whichx, fname)

### 2015-10-14 version1.0/HBM.tex
switching to simple model and back again

@@
 $\gamma$, $\delta$, and $\sigma_g$.
-These relationships can be described as Gaussian distributions with means
+These relations are Gaussian distributions with means
 given by the equation of a straight line, and standard deviations which
@@
 likelhood, we expect any bias introduced by the importance sampling method to
-be negligable.}.
-In what follows we give an overview of this procedure, but encourage the reader
-to refer back  to this original reference for a more detailed description of
-the mathematics.
-The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be thought
-of as a single `draw' from an underlying (assumed Gaussian) probability
+be negligable.}, may be used.
+We tested a model of the data where the intrinsic scatter {\it does} depend on
+the dependent variable but found that there was no strong dependence, so opted
+to use the analytic version.
+In what follows we give an overview of the importance sampling procedure, but
+encourage the reader to refer back to the original reference for a more
+detailed description of the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+thought of as a single `draw' from an underlying (assumed Gaussian) probability
 distribution, with mean equal to the `true' value\footnote{We use `true' to
@@
 
-Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
-models.
-The shaded regions show the 1 and 2$\sigma$ confidence interval.
-These intervals are representative of the uncertainty in the flicker relations,
-but also of the intrinsic scatter in those relations.
-% Given a measurement of flicker, the predictive probability distribution of
-% \logg and \rhostar will be a Gaussian distribution
-
-We also used an analytic form of the marginal likelihood, which accounts for
-2-d uncertainties but does not allow for the modelling of the intrinsic scatter
-as a function of $x$ or $y$.
+We also performed this inference using an analytic form of the marginal
+likelihood, which accounts for 2-D uncertainties but does not allow the
+intrinsic scatter to be a function of the dependent or independent variables.
 For the relation between flicker and \rhostar This likelihood function can be
@@
 \end{eqnarray}
+\label{eq:likelihood}
 and similarly for \logg.
-The parameter values found using this simpler likelihood function are included
+The parameter values found using this simple likelihood function are included
 in table \ref{tab:results2}.
+
+The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
+cases where the intrinsic scatter is not a function of $x$ or $y$.
+
+Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
+models.
+The shaded regions show the 1 and 2$\sigma$ confidence interval.
+These intervals are representative of the uncertainty in the flicker relations,
+but also of the intrinsic scatter in those relations.
+% Given a measurement of flicker, the predictive probability distribution of
+% \logg and \rhostar will be a Gaussian distribution
 
@@
 \begin{center}
-\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
+\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker_simple.pdf}
 \caption{
@@
 \begin{center}
-\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
+\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker_simple.pdf}
 \caption{

### 2015-10-14 version1.0/HBM.tex
small adjustments to the intro, discussion, abstract and method

--- 2015-08-26 version1.0/HBM.tex
@@
 \begin{equation}
-	\log_{10}(\rho_\star) \sim \mathcal{N}(\mu = \alpha + \beta
-	\log_{10}(F_8), \sigma = \sigma_{\rho}),
+	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
+	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
 \end{equation}
@@
 \begin{equation}
-	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
-	\sigma = \sigma_g).
+	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
+	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
 \end{equation}
 \label{eq:logg}
-The free parameters of the two models are $\alpha$, $\beta$, $\sigma_{\rho}$,
-$\gamma$, $\delta$, and $\sigma_g$.
-These relationships can be described as Gaussian distributions with means
+The free parameters of the two models are $\alpha$, $\beta$, $\gamma$
+$\sigma_{\rho}$, $\delta$, $\epsilon$, $\zeta$ and $\sigma_g$.
+These relations are Gaussian distributions with means
 given by the equation of a straight line, and standard deviations which
@@
 be negligable.}.
-In what follows we give an overview of this procedure, but encourage the reader
-to refer back  to this original reference for a more detailed description of
-the mathematics.
-The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be thought
-of as a single `draw' from an underlying (assumed Gaussian) probability
+In what follows we give an overview of this procedure, but encourage the
+reader to refer back  to this original reference for a more detailed
+description of the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+thought of as a single `draw' from an underlying (assumed Gaussian)
+probability
+In what follows we give an overview of the importance sampling procedure, but
+encourage the reader to refer back to the original reference for a more
+detailed description of the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+thought of as a single `draw' from an underlying (assumed Gaussian) probability
 distribution, with mean equal to the `true' value\footnote{We use `true' to
@@
 
+We also performed this inference using an analytic form of the marginal
+likelihood, which accounts for 2-D uncertainties but does not allow the
+intrinsic scatter to be a function of the dependent or independent variables.
+For the relation between flicker and \rhostar, this likelihood function can be
+written as
+\begin{eqnarray}
+	& p(\rho_*| F_8, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
+						      & \exp \left[-\frac{1}{2}
+		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
+	{\left[\beta \sigma_{F8, n}^2 + (\sigma_{\rho *, n}
+	+ \sigma_{\rho})^2\right]}\right]
+	\\ \nonumber
+	& + \beta^2 \sigma_{F8, n}^2 + (\sigma_{\rho *, n} + \sigma_{\rho})^2,
+\end{eqnarray}
+\label{eq:likelihood}
+and similarly for \logg.
+The parameter values found using this simple likelihood function are included
+in table \ref{tab:results2}.
+% The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
+% cases where the intrinsic scatter is not a function of $x$ or $y$.
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
 models.
-The shaded regions show the 1 and 2$\sigma$ confidence interval.
-These intervals are representative of the uncertainty in the flicker relations,
-but also of the intrinsic scatter in those relations.
+The shaded regions show the 1 and 2$\sigma$ confidence interval which are
+representative of the intrinsic scatter in the relations.
 % Given a measurement of flicker, the predictive probability distribution of
 % \logg and \rhostar will be a Gaussian distribution
-
-We also used an analytic form of the marginal likelihood, which accounts for
-2-d uncertainties but does not allow for the modelling of the intrinsic scatter
-as a function of $x$ or $y$.
-For the relation between flicker and \rhostar This likelihood function can be
-written as
-\begin{eqnarray}
-	p(\rho_*|\alpha, \beta, \sigma_{\rho}) = &
-	\exp \left[-\frac{1}{2} \sum_{n=1}^N\frac{[y_n-(\alpha + \beta x_n)]^2}
-	{\left[\beta \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2\right]}\right]
-	\\ \nonumber
-	& + \beta^2 \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2,
-\end{eqnarray}
-and similarly for \logg.
-The parameter values found using this simpler likelihood function are included
-in table \ref{tab:results2}.
 
@@
     $\beta$          &   \\
+    $\gamma$         &   \\
     $\sigma_{\rho}$  &   \\
-    $\gamma$         &   \\
     $\delta$         &   \\
+    $\epsilon$       &   \\
+    $\zeta$          &   \\
     $\sigma_{g}$     &   \\

### 2015-10-14 version1.0/discussion.tex
small adjustments to the intro, discussion, abstract and method

@@
 fluctuations in the {\it Kepler} light curves of stars (flicker) with both
-stellar density and surface gravity.
+stellar density and surface gravity, whilst including parameters to describe
+the intrinsic scatter in these relationships.
+We included both an additive variance term {\it and} a term that allows for
+some abscissa-dependent variance, presented in table \ref{tab:table}.
+The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both greater than zero,
+suggesting that there is an additional source of scatter in the relations, not
+accounted for by the observational uncertainties alone.
+This is either caused by intrinsic scatter in the physical relationship
+between flicker and density and \logg, produced by some physical process that
+is not accounted for in the model, or by an underestimation of the
+observational uncertainties.
+Interestingly, the abscissa-dependent variances, $\gamma$ and $\zeta$ are both
+less than zero.
+This indicates that the intrinsic scatter either decreases with increasing
+flicker and decreasing stellar density and surface gravity, or that the
+uncertainty underestimation is more severe for the giant stars.
+Regardless of the cause of the additional `jitter' needed to describe the
+relations between these parameters, the conclusions of this study are the same.
+Firstly, naively modelling stochastic relationships deterministically will
+lead to an underprediction of the dependent variable.
+Secondly, we find no evidence for a weakening dependence at higher surface
+gravities and densities, i.e. the relations between flicker, surface gravity
+and stellar density are equally valid for dwarfs and giants alike.
+
 This is a simple `fitting a line to data' exercise, however it continues the
@@
 uncertainties by marginalizing over latent variables using importance sampling.
-Although none of these methods, nor the data used here are new taken alone,
-we use the example of flicker to demonstrate the ideas and the importance of
-probabilistic, hierarchical inference.
+% Although none of these methods, nor the data used here are new taken alone,
+% we use the example of flicker to demonstrate the ideas and the importance of
+% probabilistic, hierarchical inference.
 Relationships between astronomical parameters are almost always

### 2015-10-14 version1.0/introduction.tex
small adjustments to the intro, discussion, abstract and method

--- 2015-06-08 version1.0/introduction.tex
@@
 
-TEST
-
 Accurate stellar characterization plays a vital role for many active research
-fields within astronomy. For example, stellar populations, galactic archaeology, 
-the study of binary stars, asteroseismology and exoplanet studies all rely on 
-inferences of basic stellar parameters to varying degrees. Empirically-derived
-and reliable estimates are of particular value, increasing our confidence in the
-end-product results built upon these inputs.
+fields within astronomy. For example, stellar populations, galactic
+archaeology, the study of binary stars, asteroseismology and exoplanet studies
+all rely on inferences of basic stellar parameters to varying degrees.
+Empirically-derived and reliable estimates are of particular value, increasing
+our confidence in the end-product results built upon these inputs.
 
@@
 can be inferred using one (or more) of several types of observations, such as
-spectroscopy, photometry, interferometry, etc. This inference can be done by
-invoking theoretical models or by building an empirical calibration library.
+spectroscopy, photometry, interferometry, etc. This inference can be performed
+by invoking theoretical models or by building an empirical calibration
+library.
 For example, an observed stellar spectrum could be matched against a library of
@@
 a particular observational input always returns a single-valued output for a
-parameter of interest. In this framework, nature itself has no variance and
-our underlying model is considered to be a perfect description of reality.
+parameter of interest, i.e. nature itself has no variance and the underlying
+model is considered to be a perfect description of reality.
 
-An alternative approach for inferring model parameters is the 
-``probabilistic'' one. Unlike the deterministic case, a single observational 
-input is intepretted to be caused by a range of possible model parameters
-described by a probability distribution. This statement is true even in the case
-of a perfect observation of infinite signal-to-noise, since our underlying 
-model itself comes with uncertainty. In recent years, there has been a shift 
-towards such methods in several areas of astronomy, particularly within the 
-exoplanet community. For example, \citet{wolfgang:2015} considered that
-the mass-radius relationship of exoplanets is probabilistic, since a particular
-sized planet could be due to a range of planet masses, cdue to unmodeled 
-variances in compositions, environment and other complications. These recent 
-demonstrations in exoplanetary science have prompted us to consider the need for 
-treating the parent stars in the same probabilistic framework, with potential 
-applications spanning many fields of astronomy.
+An alternative approach for inferring model parameters is to allow
+relationships between observables to be stochastic.
+Unlike the deterministic case, a single observational input is intepreted to
+be caused by a range of possible model parameters, described by a probability
+distribution.
+This statement is true even in the case of a perfect observation of infinite
+signal-to-noise, since the underlying model itself comes with uncertainty.
+In recent years, there has been a shift towards such methods in several areas
+of astronomy, particularly within the exoplanet community.
+For example, \citet{wolfgang:2015} considered that the mass-radius
+relationship of exoplanets is stochastic, since a particular sized planet
+could be have a range of planet masses due to unmodeled variances in
+compositions, environment and other complications.
+These recent demonstrations in exoplanetary science have prompted us to
+consider the need for treating the parent stars in the same probabilistic
+framework, with potential applications spanning many fields of astronomy.
 
-\textbf{
 The demand for probilistic stellar parameters is not only motivated by the fact
 that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, but also... (ruth will continue this sentence)
-}
+about astrophysical parameters, it also has a practical purpose.
+When using data published in the astronomical literature to, for example, infer
+relationships between observed parameters, that inference can be performed as
+the final stage in a hierarchical treatment
+\citep[see, e.g.][]{forman-mackey:2014}.
+This is useful if one wants to account for multi-dimensional uncertainties by
+marginalizing over the `true' parameter values upon which the noisy
+observations are conditioned.
+This point is discussed further in \textsection\ref{sec:HBM}.
 
@@
 One of the more recent tools developed to characterize stars is known as
-``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
-8-hour timescale (denoted as $F_8$) in a broad visible bandpass time series
-photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
-mission. A more detailed account of the proceedure to calculate flicker is
-described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
-displays a remarkable correlation to the asteroseismically determined parent
-star surface gravities (\logg). Turning this around, the observation implies
-that flicker can be used to blindly infer surface gravities at the level of
-$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
-curves available through the array of exoplanet transit missions flying and
-scheduled to launch.
+``flicker'' \citep{bastien:2013}.
+Flicker is a proxy for the scatter on an 8-hour timescale (denoted as $F_8$)
+in a broad visible bandpass time series photometric light curve, such as that
+from \textit{Kepler} or the upcoming TESS mission. A more detailed account of
+the proceedure to calculate flicker is described in \citet{bastien:2013}. As
+shown in \citet{bastien:2013}, flicker displays a remarkable correlation to
+the asteroseismically determined parent star surface gravities (\logg).
+Turning this around, the observation implies that flicker can be used to
+blindly infer surface gravities at the level of $\sim0.1$\,dex, an attractive
+proposition given the wealth of photometric light curves available through the
+array of exoplanet transit missions flying and scheduled to launch.
 
@@
 indeed reproduce a flicker effect in close agreement with that observed by
-\citet{bastien:2013}, providing a physically-plausible explanation. Since
-surface gravity is highly correlated to mean stellar density (\rhostar) on
-evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
-be used to infer \rhostar, which is more useful for exoplanet transit
-analysis \citep{seager:2003}.
+\citet{bastien:2013}, providing a physically-plausible explanation.
+Since surface gravity is highly correlated with mean stellar density (\rhostar)
+on evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
+be used to infer \rhostar, which is more useful for exoplanet transit analysis
+\citep{seager:2003}.
 
@@
 of the problem which are attractive for our purposes of a simple demonstration
-of probabilistic inference of stellar parameters. Firstly, in log-log space the
-relationship is very simple, appearing to be linear \citep{kipping:2014}. 
+of probabilistic inference of stellar parameters.
+Firstly, in log-log space the relationship is very simple, appearing to be
+linear \citep{kipping:2014}.
 Secondly, there is a sufficiently large number of points in the sample (439
-stars) to constrain a population-based model. Thirdly, there is significant 
-excess scatter around the best-fitting relation implying that a deterministic 
-model is inadequate. This is not surprising given that granulation is a complex 
-and messy process for which one should not expect any parametric model to 
-provide a perfect description. For these reasons, we identify the calibration of 
+stars) to constrain a population-based model.
+Thirdly, there is significant excess scatter around the best-fitting relation
+implying that a deterministic model is inadequate.
+This is not surprising given that granulation is a complex and messy process
+for which one should not expect any parametric model to provide a perfect
+description.
+Fourthly, the physical processes that produce surface granulation, of which
+flicker is an observational tracer, may be more `noisy' for different types of
+stars.
+We will test whether flicker has greater predictive power in certain regions
+of parameter space; i.e. is flicker significantly more informative for
+subgiants than for dwarfs?
+For these reasons, we identify the calibration of
 flicker to \logg\ and \rhostar\ as a well-posed problem to first demonstrate

### 2015-10-14 version1.0/manuscript.tex
small adjustments to the intro, discussion, abstract and method

@@
 % results
-These new relations will provide probability distributions over values of
-\logg\ or \rhostar\ for given measurements of flicker.
+We find evidence for additional scatter in the relationships, that cannot be
+accounted for by the observational uncertainties alone.
+This intrinsic scatter increases with decreasing flicker, i.e. there is more
+additional scatter for subgiants than for dwarfs, suggesting that the flicker
+relationships are equally (if not more) valid for dwarf stars.
 

### 2015-10-26 code/extra.py
fixing bugs

@@ -9,7 +9,7 @@ import sys
 import h5py
 
 def lnprior_extra(pars, mm=False):
-    if -10 < pars[0] < 0 and 0 < pars[1] < 10 and -100 < pars[2] < 100 \
+    if -10 < pars[0] < 10 and -10 < pars[1] < 10 and -100 < pars[2] < 100 \
             and -100 < pars[3] < 100:
         return 0., 0.
     return -np.inf, None
@@ -27,6 +27,8 @@ def lnprob(pars, samples, obs, u, extra, f):
 #     print lnlikeHF(pars, samples, obs, u, extra=extra) + \
 #             lnprior_extra(pars)
 #     raw_input('tetner')
+#     print lnlikeHF(pars, samples, obs, u, extra=extra) + \
+#             lnprior_extra(pars)
     return lnlikeHF(pars, samples, obs, u, extra=extra) + \
             lnprior_extra(pars)
 #     elif f:
@@ -46,8 +48,8 @@ def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
     """
 
     # set initial parameters
-    if fname == "f_extra":
-        rho_pars = [-1.793214679, 5.34215688, 0.02334097, -.0002600777]
+    if extra:
+        rho_pars = [-1.793214679, 5.34215688, 0.02334097, .0002600777]
         logg_pars = [-1.02143776, 5.69156647, .24239756, .049233887]
     else:
         rho_pars = [-1.69293833, 5.1408906, .0065]
@@ -82,6 +84,11 @@ def MCMC(whichx, nsamp, fname, nd, extra, f, bigdata, burnin=500, run=1000):
     flat_lls = np.reshape(lls, (np.shape(lls)[0]*np.shape(lls)[1]))
     samp = np.vstack((sampler.chain[:, :, :].reshape(-1, ndim).T, flat_lls)).T
 
+    sa = samp.T[0]
+    print type(sa)
+    print np.isfinite(sa)
+    print sa
+    print np.shape(sa), np.shape(sa[np.isfinite(sa)])
     # save samples
     f = h5py.File("%s_samples_%s.h5" % (whichx, fname), "w")
     data = f.create_dataset("samples", np.shape(samp))
@@ -101,7 +108,7 @@ def make_plots(whichx, fname):
     with h5py.File("%s_samples_%s.h5" % (whichx, fname)) as f:
         samp = f["samples"][:, :-1]
 
-    if fname == "f_extra":
[…]

### 2015-10-26 code/plots.py

@@ -22,6 +22,29 @@ def fit_straight_line(x, y, yerr):
 def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
                           fractional=False, extra=False):
 
+    # use highest likelihood samples
+    lls = samples[:, -1]
+    m = lls == max(lls)
+
+    if fname == "simple":
+        beta, alpha, tau = [samples[i, m]
+                               for i in range(np.shape(samples)[0]-1)]
+    else:
+        beta, alpha, tau, f = [samples[i, m]
+                           for i in range(np.shape(samples)[0]-1)]
+#     sigma = abs(tau)**.5
+    sigma = tau
+    print "parameters:"
+    bp = np.percentile(samples[0, :], [16, 50, 84])
+    print "beta =", beta, "-", bp[1]-bp[0], "+", bp[2]-bp[1]
+    ap = np.percentile(samples[1, :], [16, 50, 84])
+    print "alpha =", alpha, "-", ap[1]-ap[0], "+", ap[2]-ap[1]
+    tp = np.percentile(samples[1, :], [16, 50, 84])
+    print "tau =", tau, "-", tp[1]-tp[0], "+", tp[2]-tp[1]
+    if fname != "simple":
+        fp = np.percentile(samples[1, :], [16, 50, 84])
+        print "f =", f, "-", tp[1]-tp[0], "+", tp[2]-tp[1]
+
     # draw samples from post
     b_samp = np.random.choice(samples[0, :], ndraws)
     a_samp = np.random.choice(samples[1, :], ndraws)
@@ -34,6 +57,13 @@ def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
     if whichx == "rho":
         plt.ylabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
         col = "#FF33CC"
+        plt.text(1.4, .5, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \
+                \\left(\\alpha + \\beta \log_{10}(F_8), \
+                \\sigma=\sqrt{\\sigma_{\\rho}^2+\\gamma F_8}\\right)$")
+        plt.text(1.95, .24, "$\\alpha = %.3f$" % (alpha-3))
+        plt.text(1.95, .09, "$\\beta = %.3f$" % beta)
[…]

### 2015-10-26 code/plots.py
printing latex table

@@ -34,16 +34,33 @@ def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
                            for i in range(np.shape(samples)[0]-1)]
 #     sigma = abs(tau)**.5
     sigma = tau
+    if whichx == "rho":
+        pars = ["alpha", "beta", "rho", "gamma"]
+        alpha -= 3
+    if whichx == "logg":
+        pars = ["delta", "epsilon", "g", "zeta"]
     print "parameters:"
-    bp = np.percentile(samples[0, :], [16, 50, 84])
-    print "beta =", beta, "-", bp[1]-bp[0], "+", bp[2]-bp[1]
     ap = np.percentile(samples[1, :], [16, 50, 84])
-    print "alpha =", alpha, "-", ap[1]-ap[0], "+", ap[2]-ap[1]
+    print "$\\%s$ &    %s$_{-%s}^{+%s}$ & \\" \
+            % (pars[0], np.round(alpha[0], 3),
+                    np.round(ap[1], 2)-np.round(ap[0], 2),
+                    np.round(ap[2], 2)-np.round(ap[1], 2))
+    bp = np.percentile(samples[0, :], [16, 50, 84])
+    print "$\\%s$ &    %s$_{-%s}^{+%s}$ & \\" \
+            % (pars[1], np.round(beta[0], 3),
+                    np.round(bp[1], 2)-np.round(bp[0], 2),
+                    np.round(bp[2], 2)-np.round(bp[1], 2))
     tp = np.percentile(samples[1, :], [16, 50, 84])
-    print "tau =", tau, "-", tp[1]-tp[0], "+", tp[2]-tp[1]
+    print "$\\sigma_%s$ &    %s$_{-%s}^{+%s}$ & \\" \
+            % (pars[2], np.round(tau[0], 3),
+                    np.round(tp[1], 2)-np.round(tp[0], 2),
+                    np.round(tp[2], 2)-np.round(tp[1], 2))
     if fname != "simple":
         fp = np.percentile(samples[1, :], [16, 50, 84])
-        print "f =", f, "-", tp[1]-tp[0], "+", tp[2]-tp[1]
+        print "$\\%s$ &    %s$_{-%s}^{+%s}$ & \\" \
+            % (pars[3], np.round(f[0], 3),
+                    np.round(fp[1], 2)-np.round(fp[0], 2),
+                    np.round(fp[2], 2)-np.round(fp[1], 2))
 
[…]

### 2015-10-27 version1.0/HBM.tex
updating table
finished the words, just need to update the table and figures

@@
 measurement uncertainties may have an unrepresentative greater weighting
-during the fitting process.
+during the fitting process \citep{hogg:2010b}.
 Secondly, we are interested in producing probability distributions over stellar
@@
 these probability distributions through to subsequent analyses.
-Several recent studies have required posterior Probability Density
+Several recent studies have required posterior Probability Distribution
 Function (PDF) samples, in order to conduct their (hierarchical)
@@
 given by the equation of a straight line, and standard deviations which
-describe the intrinsic scatter about the mean.
+describe the intrinsic scatter about the mean as a function of flicker.
 
@@
 The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
-thought of as a single `draw' from an underlying (assumed Gaussian) probability
-distribution, with mean equal to the `true' value\footnote{We use `true' to
-mean the value that would have been observed, given infinitely high
-signal-to-noise.} and standard deviation equal to the observational
-uncertainty.
+thought of as a single `draw' from an underlying (assumed Gaussian)
+probability distribution, with mean equal to the `true' value---the value that
+would have been observed, given infinitely high signal-to-noise---and standard
+deviation equal to the observational uncertainty.
 We compute the likelihood of the data given the model, marginalized over the
@@
 if they were available, however generating new samples is an acceptable
-approximation provided the posteriors are Gaussian (or log-Normal) and the
-priors used by the previous fitters were uninformative.
-Such an approximation {\it will not be necessary} for those studies using
-\rhostar\ or \logg\ calculated using the newly-calibrated relations
-presented here, since we have published our posterior samples.
+approximation provided the posteriors are Gaussian and the priors used by the
+previous fitters were uninformative.
+Such an approximation will not be necessary for those studies using \rhostar\
+or \logg\ calculated using the newly-calibrated relations presented here,
+since we have published our posterior samples.
 After producing new posterior PDF samples we then add up the individual
@@
 latent variables.
-We used flat priors for all except the $\sigma$ parameters, for which we used
-priors which were flat in log-space.
+% We used flat priors for all except the $\sigma$ parameters, for which we used
+% priors which were flat in log-space.
+We used uniform priors over all our parameters:
+\begin{eqnarray}
+	&	\alpha, \beta, \delta, \epsilon
+	\sim U(-10:10) \\ \nonumber
+ 	&	\sigma_{rho}, \gamma, \sigma_g, \zeta  \sim U(-100:100).
+\end{eqnarray}
+\label{eq:priors}
 
-We also performed this inference using an analytic form of the marginal
-likelihood, which accounts for 2-D uncertainties but does not allow the
-intrinsic scatter to be a function of the dependent or independent variables.
+We also performed this inference using an alternative, analytic form of the
+marginal likelihood, which accounts for 2-D uncertainties but does not allow
+the intrinsic scatter to be a function of the dependent or independent
+variables.
 For the relation between flicker and \rhostar, this likelihood function can be
@@
 The parameter values found using this simple likelihood function are included
-in table \ref{tab:results2}.
+in table \ref{tab:results}.
 % The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
@@
 \begin{table}
-\caption{Highest-likelihood parameter values with their 16th and 84th
+\caption{Highest-likelihood parameter values with 16th and 84th
 	percentile uncertainties.}
-\begin{tabular}{lcc}
+\begin{tabular}{lccc}
 \hline\hline
-    Parameter & highest-likelihood value \\
+Parameter & highest-likelihood value & simple model result \\
     \hline
-    $\alpha$         &   \\
-    $\beta$          &   \\
-    $\gamma$         &   \\
-    $\sigma_{\rho}$  &   \\
-    $\delta$         &   \\
-    $\epsilon$       &   \\
-    $\zeta$          &   \\
-    $\sigma_{g}$     &   \\
+    $\alpha$ &    2.287$_{-0.03}^{+0.14}$ &     2.191$_{-0.05}^{+0.05}$  \\
+    $\beta$ &    -1.762$_{-0.07}^{+0.02}$ &     -1.724$_{-0.03}^{+0.03}$ \\
+    $\sigma_{rho}$ &    0.009$_{-0.03}^{+0.14}$ & 0.085$_{-0.05}^{+0.05}$  \\
+    $\gamma$ &    -0.0$_{-0.03}^{+0.14}$ &  \\
+    $\delta$ &    5.287$_{-0.03}^{+0.14}$ &     5.739$_{-0.03}^{+0.03}$  \\
+    $\epsilon$ &    -1.762$_{-0.07}^{+0.02}$ &  -1.095$_{-0.02}^{+0.01}$ \\
+    $\sigma_g$ &    0.009$_{-0.03}^{+0.14}$ &   0.015$_{-0.03}^{+0.03}$  \\
+    $\zeta$ &    -0.0$_{-0.03}^{+0.14}$ & \\
     \hline
+\label{tab:results}
 \end{tabular}

### 2015-10-27 version1.0/introduction.tex
finished the words, just need to update the table and figures

@@
 the final stage in a hierarchical treatment
-\citep[see, e.g.][]{forman-mackey:2014}.
-This is useful if one wants to account for multi-dimensional uncertainties by
-marginalizing over the `true' parameter values upon which the noisy
-observations are conditioned.
+\citep[see, e.g.][]{foreman-mackey:2014}.
+This is useful if, for example, one wants to account for multi-dimensional
+uncertainties by marginalizing over the `true' parameter values upon which the
+noisy observations are conditioned.
 This point is discussed further in \textsection\ref{sec:HBM}.
@@
 Fourthly, the physical processes that produce surface granulation, of which
-flicker is an observational tracer, may be more `noisy' for different types of
-stars.
+flicker is an observational tracer, may be more or less noisy for different
+types of stars.
 We will test whether flicker has greater predictive power in certain regions

### 2015-10-28 code/plots.py
sorting nan issues

@@ -119,17 +119,25 @@ def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
             ys = model1([b_samp[i], a_samp[i]], xs)
             line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
                     np.random.randn(1)*np.median(f_samp)*xs)**.5
-            plt.plot(xs, line, col, alpha=.05)
-#             lines[i, :] = line
-            lines.append(line)
+#             plt.plot(xs, line, col, alpha=.05)
+            print len(line[np.isfinite(line)])
+            if len(line[np.isfinite(line)])==len(xs):
+                print len(line[np.isfinite(line)])
+                lines.append(line)
 
         plt.xlim(min(xs), max(xs))
+        print np.shape(lines)
         # plot regions
-        quantiles = np.percentile(lines, [2, 16, 84, 98], axis=0)
-        plt.fill_between(xs, quantiles[1], quantiles[2], color=col,
+        quantiles = np.percentile(lines, [16, 84], axis=0)
+#         print np.shape(quantiles)
+#         assert 0
+        print quantiles
+        plt.fill_between(xs, quantiles[0], quantiles[1], color=col,
                          alpha=.4)
-        plt.fill_between(xs, quantiles[0], quantiles[3], color=col,
-                         alpha=.2)
+#         plt.fill_between(xs, quantiles[1], quantiles[2], color=col,
+#                          alpha=.4)
+#         plt.fill_between(xs, quantiles[0], quantiles[3], color=col,
+#                          alpha=.2)
 
         # plot best fit and data
         ym = model1([np.median(b_samp), np.median(a_samp)], xs)

### 2015-11-02 code/plots.py
small change

@@ -84,12 +84,23 @@ def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         plt.ylim(-2, 1)
 
         # plot line draws
+        lines = []
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
             y3 = ys - 3
-            line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
-                    np.random.randn(1)*np.median(f_samp)*xs)**.5 - 3
-            plt.plot(xs, line, col, alpha=.05)
+#             line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
+#                     np.random.randn(1)*np.median(f_samp)*xs)**.5 - 3
+            line = ys + (np.random.randn(1)*np.median(sigma)**2 + \
+                    np.random.randn(1)*np.median(f)*xs)**.5 - 3
+#             plt.plot(xs, line, col, alpha=.05)
+            if len(line[np.isfinite(line)])==len(xs):
+                lines.append(line)
+
+        quantiles = np.percentile(lines, [2, 16, 84, 98], axis=0)
+        plt.fill_between(xs, quantiles[1], quantiles[2], color=col,
+                         alpha=.4)
+        plt.fill_between(xs, quantiles[0], quantiles[3], color=col,
+                         alpha=.4)
 
         # plot best fit and data
         ym = model1([np.median(b_samp), np.median(a_samp)], xs)
@@ -117,22 +128,24 @@ def make_flicker_plot(x, xerr, y, yerr, samples, whichx, fname, ndraws,
         lines = []
         for i in range(ndraws):
             ys = model1([b_samp[i], a_samp[i]], xs)
-            line = ys + (np.random.randn(1)*np.median(s_samp)**2 + \
-                    np.random.randn(1)*np.median(f_samp)*xs)**.5
+            if np.all(tau[0]**2+f[0]*xs) > 0:
+                line = ys + np.random.randn(1)*(tau**2 + f*xs)**.5
 #             plt.plot(xs, line, col, alpha=.05)
-            print len(line[np.isfinite(line)])
+#             print len(line[np.isfinite(line)])
             if len(line[np.isfinite(line)])==len(xs):
-                print len(line[np.isfinite(line)])
+#                 print len(line[np.isfinite(line)])
                 lines.append(line)
 
         plt.xlim(min(xs), max(xs))
-        print np.shape(lines)
+#         print np.shape(lines)
         # plot regions
[…]

### 2015-11-02 version1.0/discussion.tex

@@
 some abscissa-dependent variance, presented in table \ref{tab:table}.
-The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both greater than zero,
-suggesting that there is an additional source of scatter in the relations, not
-accounted for by the observational uncertainties alone.
+The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both non-zero,
+suggesting that there {\it is} an additional source of scatter in the
+relations, not accounted for by the observational uncertainties alone.
 This is either caused by intrinsic scatter in the physical relationship
@@
 uncertainty underestimation is more severe for the giant stars.
-Regardless of the cause of the additional `jitter' needed to describe the
+Regardless of the cause of the additional scatter needed to describe the
 relations between these parameters, the conclusions of this study are the same.
-Firstly, naively modelling stochastic relationships deterministically will
-lead to an underprediction of the dependent variable.
-Secondly, we find no evidence for a weakening dependence at higher surface
+Firstly, naively modelling the relationship between flicker and surface
+gravity or stellar density, or any other stochastic relationship
+deterministically, will lead to an underprediction of the uncertainties in
+dependent variable.
+Secondly, we find no evidence for a increased scatter at higher surface
 gravities and densities, i.e. the relations between flicker, surface gravity
@@
 Gaussian and uncorrelated which is rarely true and certainly not the case when
-the model is a straight line (the slope and intercept are correlated by
-definition).
+the model is a straight line (the slope and intercept are alway correlated).
 A flicker value with uncertainties (or even better: posterior PDF

### 2015-11-02 version1.0/manuscript.tex

--- 2015-08-26 version1.0/manuscript.tex
@@
 The relations between observable stellar parameters are usually assumed to be
-deterministic.
-That is, given an infinitely precise measurement of independant
+deterministic, that is, given an infinitely precise measurement of independant
 variable, `$x$', and some model, the value of dependant variable, `y' can be
@@
 % methods
-The relation between short-timescale brightness fluctuations (flicker) of stars
-and both surface gravity \citep{bastien:2013} and stellar density
+The relation between short-timescale brightness fluctuations (flicker) of
+stars and both surface gravity \citep{bastien:2013} and stellar density
 \citep{kipping:2014} are two such stochastic relations that have, until now,
@@
 We recalibrate these relations in a probabilistic framework, using
-Hierarchical Bayesian Modelling (HBM) and constrain the instrinsic scatter in
+Hierarchical Bayesian Modelling (HBM) to constrain the instrinsic scatter in
 the relations.
-Our treatment also fully accounts for the observational uncertainties on all
-variables.
+% Our treatment also fully accounts for the observational uncertainties on all
+% variables.
 % results
-These new relations will provide probability distributions over values of
-\logg\ or \rhostar\ for given measurements of flicker.
+We find evidence for additional scatter in the relationships, particularly in
+the relation between flicker and surface gravity, that cannot be accounted for
+by the observational uncertainties alone.
+The scatter in surface gravity increases with decreasing flicker, i.e. there
+is greater scatter for subgiants than for dwarfs, suggesting that using flicker
+as a proxy for surface gravity is just as (if not more) valid for dwarf stars,
+despite the fact that the observational uncertainties tend to be larger for
+these stars.
 
@@
 RA would like to thank Dan Foreman-Mackey for his extremely helpful suggestions
-and comments, and Angie Wolfgang and John Johnson for their fantastic advice.
+and comments, Angie Wolfgang for her expert guidance and John Johnson for hit
+fantastic advice and support.
 RA would also like to thank the SAMSI institute and the other members of the
 SAMSI {\it Kepler} exoplanet statistics working groups.
-DMK thanks...
 
@@
 2166
+\bibitem[\protect\citeauthoryear{Hogg et al.}{2010b}]{hogg:2010b}
+Hogg, D.~W., Bovy, J. \& Lang, D., 2010
 \bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}

### 2015-12-01 version1.0/manuscript.tex
changing figures to pdf and combining files

@@
 
-\input{shortcuts.tex}
+\newcommand{\logg}{$\log g$}
+\newcommand{\rhostar}{$\rho_{\star}$}
+\newcommand{\Teff}{$T_{\mathrm{eff}}$}
+\newcommand{\FeH}{[Fe/H]}
+\newcommand{\Kepler}{\textit{Kepler}}
 
@@
 
-\input{introduction.tex}
+%% INTRODUCTION
+%%
+
+Accurate stellar characterization plays a vital role for many active research
+fields within astronomy. For example, stellar populations, galactic
+archaeology, the study of binary stars, asteroseismology and exoplanet studies
+all rely on inferences of basic stellar parameters to varying degrees.
+Empirically-derived and reliable estimates are of particular value, increasing
+our confidence in the end-product results built upon these inputs.
+
+Basic stellar parameters, such as effective temperature and surface gravity,
+can be inferred using one (or more) of several types of observations, such as
+spectroscopy, photometry, interferometry, etc. This inference can be performed
+by invoking theoretical models or by building an empirical calibration
+library.
+For example, an observed stellar spectrum could be matched against a library of
+theoretical spectra generated using stellar atmosphere models, or, against a
+library of observed spectra of ``standard stars'', serving as calibrators.
+%Fundamental stellar parameters, such as mass and radius, which cannot be
+%directly inferred from the available observations can often be inferred by
+%invoking theoretical stellar models, e.g. fitting model isochrones to
+%spectroscopic measurements.
+Regardless of the approach, be it theoretical or empirical, the methods used
+for the inference of stellar parameters are traditionally ``deterministic''.
+In this context, a deterministic model can be loosely described as one where
+a particular observational input always returns a single-valued output for a
+parameter of interest, i.e. nature itself has no variance and the underlying
+model is considered to be a perfect description of reality.
+
+An alternative approach for inferring model parameters is to allow
+relationships between observables to be stochastic.
+Unlike the deterministic case, a single observational input is intepreted to
+be caused by a range of possible model parameters, described by a probability
+distribution.
+This statement is true even in the case of a perfect observation of infinite
+signal-to-noise, since the underlying model itself comes with uncertainty.
+In recent years, there has been a shift towards such methods in several areas
+of astronomy, particularly within the exoplanet community.
+For example, \citet{wolfgang:2015} considered that the mass-radius
+relationship of exoplanets is stochastic, since a particular sized planet
+could be have a range of planet masses due to unmodeled variances in
+compositions, environment and other complications.
+These recent demonstrations in exoplanetary science have prompted us to
+consider the need for treating the parent stars in the same probabilistic
+framework, with potential applications spanning many fields of astronomy.
+
+The demand for probilistic stellar parameters is not only motivated by the fact
+that probability distributions are far more representative of our `beliefs'
+about astrophysical parameters, it also has a practical purpose.
+When using data published in the astronomical literature to, for example, infer
+relationships between observed parameters, that inference can be performed as
+the final stage in a hierarchical treatment
+\citep[see, e.g.][]{foreman-mackey:2014}.
+This is useful if, for example, one wants to account for multi-dimensional
+uncertainties by marginalizing over the `true' parameter values upon which the
+noisy observations are conditioned.
+This point is discussed further in \textsection\ref{sec:HBM}.
+
+%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
+%(HBM) to model the dispersion, or intrinsic scatter in relations between
+%astrophysical properties.
+%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
+%relation for sub-Neptunes and infer the level of astrophysical dispersion
+
+% RA text...
+
+One of the more recent tools developed to characterize stars is known as
+``flicker'' \citep{bastien:2013}.
+Flicker is a proxy for the scatter on an 8-hour timescale (denoted as $F_8$)
+in a broad visible bandpass time series photometric light curve, such as that
+from \textit{Kepler} or the upcoming TESS mission. A more detailed account of
+the proceedure to calculate flicker is described in \citet{bastien:2013}. As
+shown in \citet{bastien:2013}, flicker displays a remarkable correlation to
+the asteroseismically determined parent star surface gravities (\logg).
+Turning this around, the observation implies that flicker can be used to
+blindly infer surface gravities at the level of $\sim0.1$\,dex, an attractive
+proposition given the wealth of photometric light curves available through the
+array of exoplanet transit missions flying and scheduled to launch.
+
+\citet{cranmer:2014} demonstrated that models of stellar surface granulation
+indeed reproduce a flicker effect in close agreement with that observed by
+\citet{bastien:2013}, providing a physically-plausible explanation.
+Since surface gravity is highly correlated with mean stellar density (\rhostar)
+on evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
+be used to infer \rhostar, which is more useful for exoplanet transit analysis
+\citep{seager:2003}.
+
+Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
+of the problem which are attractive for our purposes of a simple demonstration
+of probabilistic inference of stellar parameters.
+Firstly, in log-log space the relationship is very simple, appearing to be
+linear \citep{kipping:2014}.
+Secondly, there is a sufficiently large number of points in the sample (439
+stars) to constrain a population-based model.
+Thirdly, there is significant excess scatter around the best-fitting relation
+implying that a deterministic model is inadequate.
+This is not surprising given that granulation is a complex and messy process
+for which one should not expect any parametric model to provide a perfect
+description.
+Fourthly, the physical processes that produce surface granulation, of which
+flicker is an observational tracer, may be more or less noisy for different
+types of stars.
+We will test whether flicker has greater predictive power in certain regions
+of parameter space; i.e. is flicker significantly more informative for
+subgiants than for dwarfs?
+For these reasons, we identify the calibration of
+flicker to \logg\ and \rhostar\ as a well-posed problem to first demonstrate
+probabilistic inference in the arena of stellar characterization.
 
@@
 
-\input{HBM.tex}
+%% HBM
+%%
+
+\subsection{Calibration Data}
+
+For our calibration data, we used a sample of \Kepler\ stars with
+both asteroseismic and flicker measurements available. \citet{chaplin:2014}
+report asteroseismic \rhostar\ estimates (and the associated uncertainties) for
+518 \Kepler\ stars. The authors report three different sets of results,
+depending on the choice of \Teff\ and \FeH, and in this work we elected to use
+values reported in their Table 6 over Table 5, and Table 5 over Table 4. We
+additionally used the 71 additional planet hosting stars with asteroseismology
+reported in \citet{huber:2013} but not reported in \citet{chaplin:2014}. Values
+for flicker and ``range'' were taken from \citet{kipping:2014}, based upon the
+methods described in \citet{bastien:2013}. Following \citet{kipping:2014} and
+for reasons described there-in, we only include targets in our calibration for
+which:
+
+\begin{itemize}
+\item[{\tiny$\blacksquare$}] Range (defined in \citealt{bastien:2013})
+$<1$\,ppt
+\item[{\tiny$\blacksquare$}] $4500<T_{\mathrm{eff}}<6500$\,K
+\item[{\tiny$\blacksquare$}] $K_P<14$
+\item[{\tiny$\blacksquare$}] $1.2 < \log_{10}$($F_8$\,[ppm])$< 2.2$
+\end{itemize}
+
+We use the same sample for our calibration of \logg, except that we exclude the
+\citet{huber:2013} data, since these authors do not provide estimates of
+\logg\footnote{Whilst we could compute \logg\ ourselves from the reported
+masses and radii, this could only be done under the incorrect assumption of
+zero covariance between $M_{\star}$ and $R_{\star}$.}.
+
+\subsection{Hierarchical Bayesian Model}
+
+We model the stochastic relationship between $F_8$ and \logg\ or \rhostar,
+accounting for the fact that there exists some intrinsic scatter in
+the dependent variable, and including the heteroskedastic uncertainties on both
+the dependent and independent variables.
+There are two excellent reasons for modelling the relation stochastically;
+firstly, if the intrinsic scatter is ignored and the relation between
+variables is assumed to be deterministic, those data points with smaller
+measurement uncertainties may have an unrepresentative greater weighting
+during the fitting process \citep{hogg:2010b}.
+Secondly, we are interested in producing probability distributions over stellar
+densities and surface gravities, as opposed to point estimates, and propagating
+these probability distributions through to subsequent analyses.
+Several recent studies have required posterior Probability Distribution
+Function (PDF) samples, in order to conduct their (hierarchical)
+analyses \citep[e.g.][]{foreman-mackey:2014, rogers:2015, angus:2015}.
+
+Including observational uncertainties on the independent variable within our
+analysis is also important, because ordinary least squares regression methods
+performed on data with two-dimensional uncertainties will result in a slope
+that is biased towards zero
+\citep[e.g.][]{fuller:1987, fox:1997}.
+For a demonstration of the affects of neglecting intrinsic scatter and
+two-dimensional uncertainties, see \citet{kelly:2007}.
+
+We write the two models describing the relationship between $F_8$, \logg\ and
+\rhostar\ as
+\begin{equation}
+	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
+	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
+\end{equation}
+\label{eq:rho}
+and
+\begin{equation}
+	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
+	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
+\end{equation}
+\label{eq:logg}
+The free parameters of the two models are $\alpha$, $\beta$, $\gamma$
+$\sigma_{\rho}$, $\delta$, $\epsilon$, $\zeta$ and $\sigma_g$.
+These relations are Gaussian distributions with means
+given by the equation of a straight line, and standard deviations which
+describe the intrinsic scatter about the mean as a function of flicker.
+
+In order to account for the two-dimensional observational uncertainties in
+this data set, we marginalize over the latent, `true' values of $F_8$,
+\rhostar\ and \logg\ using the importance sampling method of
+\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
+\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
+likelhood, we expect any bias introduced by the importance sampling method to
+be negligable.}.
+In what follows we give an overview of this procedure, but encourage the
+reader to refer back  to this original reference for a more detailed
+description of the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+thought of as a single `draw' from an underlying (assumed Gaussian)
+probability
+In what follows we give an overview of the importance sampling procedure, but
+encourage the reader to refer back to the original reference for a more
+detailed description of the mathematics.
+The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+thought of as a single `draw' from an underlying (assumed Gaussian)
+probability distribution, with mean equal to the `true' value---the value that
+would have been observed, given infinitely high signal-to-noise---and standard
+deviation equal to the observational uncertainty.
+We compute the likelihood of the data given the model, marginalized over the
+true values, using importance sampling.
+We sample from the posterior PDFs of the true values of the variables,
+conditioned on the observed values, and compute the likelihood of each of those
+samples.
+% These posterior PDFs could be produced by the astronomers who provide the
+% catalog.
+% If posterior samples are made available, they can be used in this stage of
+% the inference.
+% In the event that they are not provided however, they can be generated.
+% For example, the $log(g)$ values used here are point estimates of the
+% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
+% {\it Kepler} light curves and stellar models.
+We would use the posterior samples generated in previous model fitting process
+if they were available, however generating new samples is an acceptable
+approximation provided the posteriors are Gaussian and the priors used by the
+previous fitters were uninformative.
+Such an approximation will not be necessary for those studies using \rhostar\
+or \logg\ calculated using the newly-calibrated relations presented here,
+since we have published our posterior samples.
+After producing new posterior PDF samples we then add up the individual
+log-likelihoods of each sample to compute the total marginalized likelihood
+for a star.
+The marginalized likelihood of the whole dataset is then computed as the sum
+of the log-likelihoods of each star.
+The importance sampling method allows us to approximate the integral over the
+latent variables.
+% We used flat priors for all except the $\sigma$ parameters, for which we used
+% priors which were flat in log-space.
+We used uniform priors over all our parameters:
+\begin{eqnarray}
+	&	\alpha, \beta, \delta, \epsilon
+	\sim U(-10:10) \\ \nonumber
+ 	&	\sigma_{rho}, \gamma, \sigma_g, \zeta  \sim U(-100:100).
+\end{eqnarray}
+\label{eq:priors}
+
+We also performed this inference using an alternative, analytic form of the
+marginal likelihood, which accounts for 2-D uncertainties but does not allow
+the intrinsic scatter to be a function of the dependent or independent
+variables.
+For the relation between flicker and \rhostar, this likelihood function can be
+written as
+\begin{eqnarray}
+	& p(\rho_*| F_8, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
+						      & \exp \left[-\frac{1}{2}
+		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
+	{\left[\beta \sigma_{F8, n}^2 + (\sigma_{\rho *, n}
+	+ \sigma_{\rho})^2\right]}\right]
+	\\ \nonumber
+	& + \beta^2 \sigma_{F8, n}^2 + (\sigma_{\rho *, n} + \sigma_{\rho})^2,
+\end{eqnarray}
+\label{eq:likelihood}
+and similarly for \logg.
+The parameter values found using this simple likelihood function are included
+in table \ref{tab:results}.
+% The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
+% cases where the intrinsic scatter is not a function of $x$ or $y$.
+Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
+models.
+The shaded regions show the 1 and 2$\sigma$ confidence interval which are
+representative of the intrinsic scatter in the relations.
+% Given a measurement of flicker, the predictive probability distribution of
+% \logg and \rhostar will be a Gaussian distribution
+
+%%% rhostar plot
+\begin{figure}
+\begin{center}
+% \includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
+\includegraphics[width=8.4cm,angle=0,clip=true]{new_rho.pdf}
+\caption{
+Stellar density vs. flicker.
+This figure shows the model, conditioned on the data.
+The solid black line shows the highest-likelihood sample, and the pink shaded
+regions show the 1 and 2$\sigma$ confidence intervals.}
+\label{fig:rhostar}
+\end{center}
+\end{figure}
+
+%%% logg plot
+\begin{figure}
+\begin{center}
+% \includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
+\includegraphics[width=8.4cm,angle=0,clip=true]{new_logg.pdf}
+\caption{
+$\log(g)$ vs. flicker.
+As in \ref{fig:rhostar} this figure shows the model, conditioned on the data.
+The solid black line shows the highest-likelihood sample, and the blue shaded
+regions show the 1 and 2$\sigma$ confidence intervals.}
+\label{fig:logg}
+\end{center}
+\end{figure}
+
+\begin{table}
+\caption{Highest-likelihood parameter values with 16th and 84th
+	percentile uncertainties.}
+\begin{tabular}{lccc}
+\hline\hline
+Parameter & highest-likelihood value & simple model result \\
+    \hline
+$\alpha$ &    2.358$_{-0.04}^{+0.08}$ &       2.191$_{-0.05}^{+0.05}$  \\
+$\beta$ &    -1.79$_{-0.03}^{+0.04}$ &        -1.724$_{-0.03}^{+0.03}$ \\
+$\sigma_{rho}$ &    0.015$_{-0.04}^{+0.08}$ &   0.085$_{-0.05}^{+0.05}$  \\
+$\gamma$ &    -0.0$_{-0.04}^{+0.08}$ & \\
+$\delta$ &    5.535$_{-0.07}^{+0.07}$ &      5.739$_{-0.03}^{+0.03}$  \\
+$\epsilon$ &    -1.005$_{-0.04}^{+0.03}$ &   -1.095$_{-0.02}^{+0.01}$ \\
+$\sigma_g$ &    0.168$_{-0.07}^{+0.07}$ &    0.015$_{-0.03}^{+0.03}$  \\
+$\zeta$ &    -0.014$_{-0.07}^{+0.07}$ & \\
+    \hline
+\label{tab:results}
+\end{tabular}
+\end{table}
 
@@
 
-\input{discussion.tex}
+%% DISCUSSION
+%%
+
+We have recalibrated the relation between short timescale brightness
+fluctuations in the {\it Kepler} light curves of stars (flicker) with both
+stellar density and surface gravity, whilst including parameters to describe
+the intrinsic scatter in these relationships.
+We included both an additive variance term {\it and} a term that allows for
+some abscissa-dependent variance, presented in table \ref{tab:table}.
+The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both non-zero,
+suggesting that there {\it is} an additional source of scatter in the
+relations, not accounted for by the observational uncertainties alone.
+This is either caused by intrinsic scatter in the physical relationship
+between flicker and density and \logg, produced by some physical process that
+is not accounted for in the model, or by an underestimation of the
+observational uncertainties.
+Interestingly, the abscissa-dependent variances, $\gamma$ and $\zeta$ are both
+less than zero.
+This indicates that the intrinsic scatter either decreases with increasing
+flicker and decreasing stellar density and surface gravity, or that the
+uncertainty underestimation is more severe for the giant stars.
+Regardless of the cause of the additional scatter needed to describe the
+relations between these parameters, the conclusions of this study are the same.
+Firstly, naively modelling the relationship between flicker and surface
+gravity or stellar density, or any other stochastic relationship
+deterministically, will lead to an underprediction of the uncertainties in
+dependent variable.
+Secondly, we find no evidence for a increased scatter at higher surface
+gravities and densities, i.e. the relations between flicker, surface gravity
+and stellar density are equally valid for dwarfs and giants alike.
+
+This is a simple `fitting a line to data' exercise, however it continues the
+discussion of probabilistic modelling that is an active topic within the
+fields of exoplanet and stellar astronomy.
+We use Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
+scatter in the relationship between flicker, surface gravity and density.
+We also include the effects of the non-negligable two-dimensional observational
+uncertainties by marginalizing over latent variables using importance sampling.
+% Although none of these methods, nor the data used here are new taken alone,
+% we use the example of flicker to demonstrate the ideas and the importance of
+% probabilistic, hierarchical inference.
+Relationships between astronomical parameters are almost always
+non-deterministic; an element of stochasticity effects the physical parameters
+of stars so one can never perfectly predict $y$ given an observation of $x$.
+We advocate a probabilistic approach in both the `fitting the model to data'
+step, {\it and} when using an empirically calibrated model to predict
+parameter values.
+The fitting stage benefits because if the relationships between parameters are
+falsely assumed to be deterministic, they will be skewed by data points with
+unrepresentative uncertainties.
+The prediction stage benefits from the stochastic treatment both because a
+probability distribution is in many ways more representative of an observation
+than a point estimate, and because posterior PDF samples can be used in
+subsequent studies (provided the prior used during the fitting process is
+described).
+
+We provide posterior PDF samples and the code used in this project at
+\url{https://github.com/RuthAngus/flicker}.
+Whenever a prediction for the surface gravity or density of a star is required,
+for a given estimate of flicker, we recommend using these posterior samples
+within the calculation of $\rho_{*}$ or $\log(g)$ and its (Monte Carlo)
+uncertainty.
+These posterior samples will naturally fold-in the covariances between
+parameters.
+Simple analytical uncertainty propagation is only valid when uncertainties are
+Gaussian and uncorrelated which is rarely true and certainly not the case when
+the model is a straight line (the slope and intercept are alway correlated).
+A flicker value with uncertainties (or even better: posterior PDF
+samples), input into our model will result in a probability distribution over
+stellar densities or surface gravities which reflects both the uncertainties
+on the flicker measurement, the uncertainties on the model parameters {\it and}
+the intrinsic scatter in the flicker-$\rho_{*}$-$\log(g)$ relations.
 

### 2015-12-02 version1.0/manuscript.tex
now with Dave's comments, about to submit

@@
 \author{
-	{\bf	Ruth~Angus\altaffilmark{1,2},
-		David.~M.~Kipping\altaffilmark{2,3},
+	{\bf	Ruth~Angus\altaffilmark{1},
+		David.~M.~Kipping\altaffilmark{2},
 	}
@@
 
-\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics,
-		Cambridge, MA 02138, USA; email: dkipping@cfa.harvard.edu}
-
-\altaffiltext{3}{Department of Astronomy, Columbia University, 550 W 120th St.,
+\altaffiltext{2}{Department of Astronomy, Columbia University, 550 W 120th St.,
                  New York, NY 10027, US}
@@
 
-The demand for probilistic stellar parameters is not only motivated by the fact
+The demand for probabilistic stellar parameters is not only motivated by the fact
 that probability distributions are far more representative of our `beliefs'
@@
 description.
-Fourthly, the physical processes that produce surface granulation, of which
+Finally, the physical processes that produce surface granulation, of which
 flicker is an observational tracer, may be more or less noisy for different
@@
 and similarly for \logg.
-The parameter values found using this simple likelihood function are included
-in table \ref{tab:results}.
+The parameter values found using this simple likelihood function, with a
+constant dispersion are included in table \ref{tab:results}.
 % The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
@@
 \hline\hline
-Parameter & highest-likelihood value & simple model result \\
+Parameter & Variable dispersion model & Constant dispersion model \\
     \hline
@@
 $\gamma$ &    -0.0$_{-0.04}^{+0.08}$ & \\
+\hline
 $\delta$ &    5.535$_{-0.07}^{+0.07}$ &      5.739$_{-0.03}^{+0.03}$  \\
@@
     \hline
-\label{tab:results}
 \end{tabular}
 \end{table}
+\label{tab:results}
 
@@
 dependent variable.
-Secondly, we find no evidence for a increased scatter at higher surface
+Secondly, we find marginal evidence for decreased scatter at higher surface
 gravities and densities, i.e. the relations between flicker, surface gravity
-and stellar density are equally valid for dwarfs and giants alike.
+and stellar density are valid for dwarfs and giants alike, despite the fact
+that observational uncertainties tend to be larger for dwarfs.
 
@@
 fields of exoplanet and stellar astronomy.
-We use Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
+We used Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
 scatter in the relationship between flicker, surface gravity and density.
-We also include the effects of the non-negligable two-dimensional observational
+We also included the effects of the non-negligable two-dimensional observational
 uncertainties by marginalizing over latent variables using importance sampling.

### 2016-01-20 version1.0/manuscript.tex
starting to rewrite

@@
 accounting for the fact that there exists some intrinsic scatter in
-the dependent variable, and including the heteroskedastic uncertainties on both
-the dependent and independent variables.
+the dependent variable.
+ % and including the heteroskedastic uncertainties on both
+% the dependent and independent variables.
 There are two excellent reasons for modelling the relation stochastically;

### 2016-01-20 version1.0/manuscript.tex
changing HBM

@@
 The relations between observable stellar parameters are usually assumed to be
-deterministic, that is, given an infinitely precise measurement of independant
-variable, `$x$', and some model, the value of dependant variable, `y' can be
-known exactly.
+deterministic.
+That is, given an infinitely precise measurement of independent variable,
+`$x$', and some model, the value of dependent variable, `$y$' can be known
+exactly.
 In practise this assumption is rarely valid and intrinsic stochasticity means
@@
 
-We model the stochastic relationship between $F_8$ and \logg\ or \rhostar,
+We model the stochastic relationship between $F_8$, \logg\ and \rhostar,
 accounting for the fact that there exists some intrinsic scatter in
@@
 during the fitting process \citep{hogg:2010b}.
-Secondly, we are interested in producing probability distributions over stellar
-densities and surface gravities, as opposed to point estimates, and propagating
-these probability distributions through to subsequent analyses.
+Secondly, we are interested in producing probability distributions over
+stellar densities and surface gravities, as opposed to point estimates, and
+propagating these probability distributions through to subsequent analyses.
 Several recent studies have required posterior Probability Distribution
-Function (PDF) samples, in order to conduct their (hierarchical)
-analyses \citep[e.g.][]{foreman-mackey:2014, rogers:2015, angus:2015}.
+Function (PDF) samples, in order to conduct their hierarchical analyses
+\citep[e.g.][]{foreman-mackey:2014, rogers:2015, angus:2015}
+% These studies account for multi-dimensional uncertainties using the sampling
+% method outlined in \citet{hogg:2010}.
+% This can be important because ordinary least squares regression methods
+% performed on data with two-dimensional uncertainties can result in a slope
+% that is biased towards zero \citep[e.g.][]{fuller:1987, fox:1997}.
+% For a demonstration of the affects of neglecting intrinsic scatter and
+% two-dimensional uncertainties, see \citet{kelly:2007}.
 
-Including observational uncertainties on the independent variable within our
-analysis is also important, because ordinary least squares regression methods
-performed on data with two-dimensional uncertainties will result in a slope
-that is biased towards zero
-\citep[e.g.][]{fuller:1987, fox:1997}.
-For a demonstration of the affects of neglecting intrinsic scatter and
-two-dimensional uncertainties, see \citet{kelly:2007}.
+% Multi-dimensional uncertainties have been included in a number of recent
+% studies, including \citet{foreman-mackey:2014, rogers:2015, angus:2015}.
+% In order to account for the two-dimensional observational uncertainties in
+% this data set, we marginalize over the latent, `true' values of $F_8$,
+% \rhostar\ and \logg\ using the importance sampling method of
+% \citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
 
-We write the two models describing the relationship between $F_8$, \logg\ and
-\rhostar\ as
+The two models we use to describe the relationships between $F_8$, \logg\ and
+\rhostar\ are
 \begin{equation}
-	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
-	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
+% 	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
+% 	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
+	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha_\rho +
+    \beta_\rho \log_{10}(F_8), \sigma^2 = \sigma_\rho^2 \right),
 \end{equation}
@@
 \begin{equation}
-	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
-	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
+% 	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
+% 	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
+	\log_{10}(g) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
+    \log_{10}(F_8), \sigma^2 = \sigma_g^2 \right).
 \end{equation}
 \label{eq:logg}
-The free parameters of the two models are $\alpha$, $\beta$, $\gamma$
-$\sigma_{\rho}$, $\delta$, $\epsilon$, $\zeta$ and $\sigma_g$.
-These relations are Gaussian distributions with means
-given by the equation of a straight line, and standard deviations which
-describe the intrinsic scatter about the mean as a function of flicker.
+The free parameters of the two models are $\alpha_\rho$, $\beta_\rho$,
+$\sigma_\rho$, $\alpha_g$, $\beta_g$ and $\sigma_g$.
+These relations are Gaussian distributions with means given by the equation of
+a straight line, and standard deviations which describe the intrinsic scatter
+about the mean.
 
-In order to account for the two-dimensional observational uncertainties in
-this data set, we marginalize over the latent, `true' values of $F_8$,
-\rhostar\ and \logg\ using the importance sampling method of
-\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
-\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
-likelhood, we expect any bias introduced by the importance sampling method to
-be negligable.}.
-In what follows we give an overview of this procedure, but encourage the
-reader to refer back  to this original reference for a more detailed
-description of the mathematics.
-The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
-thought of as a single `draw' from an underlying (assumed Gaussian)
-probability
-In what follows we give an overview of the importance sampling procedure, but
-encourage the reader to refer back to the original reference for a more
-detailed description of the mathematics.
-The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
-thought of as a single `draw' from an underlying (assumed Gaussian)
-probability distribution, with mean equal to the `true' value---the value that
-would have been observed, given infinitely high signal-to-noise---and standard
-deviation equal to the observational uncertainty.
-We compute the likelihood of the data given the model, marginalized over the
-true values, using importance sampling.
-We sample from the posterior PDFs of the true values of the variables,
-conditioned on the observed values, and compute the likelihood of each of those
-samples.
-% These posterior PDFs could be produced by the astronomers who provide the
-% catalog.
-% If posterior samples are made available, they can be used in this stage of
-% the inference.
-% In the event that they are not provided however, they can be generated.
-% For example, the $log(g)$ values used here are point estimates of the
-% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
-% {\it Kepler} light curves and stellar models.
-We would use the posterior samples generated in previous model fitting process
-if they were available, however generating new samples is an acceptable
-approximation provided the posteriors are Gaussian and the priors used by the
-previous fitters were uninformative.
-Such an approximation will not be necessary for those studies using \rhostar\
-or \logg\ calculated using the newly-calibrated relations presented here,
-since we have published our posterior samples.
-After producing new posterior PDF samples we then add up the individual
-log-likelihoods of each sample to compute the total marginalized likelihood
-for a star.
-The marginalized likelihood of the whole dataset is then computed as the sum
-of the log-likelihoods of each star.
-The importance sampling method allows us to approximate the integral over the
-latent variables.
-% We used flat priors for all except the $\sigma$ parameters, for which we used
-% priors which were flat in log-space.
-We used uniform priors over all our parameters:
-\begin{eqnarray}
-	&	\alpha, \beta, \delta, \epsilon
-	\sim U(-10:10) \\ \nonumber
- 	&	\sigma_{rho}, \gamma, \sigma_g, \zeta  \sim U(-100:100).
-\end{eqnarray}
-\label{eq:priors}
+We also tested a model in which the additional scatter depends on flicker
+itself, defined as
+\begin{equation}
+	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha_\rho +
+    \beta_\rho \log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 +
+    \gamma_\rho F_8}\right),
+\end{equation}
+\label{eq:rho2}
+for flicker vs $\rho_\star$ and similarly for \logg.
+This model allowed us to determine whether there the magnitude of additional
+scatter varied as a function of \logg or \rhostar.
+In other words, whether flicker was a better proxy for \logg\ or \rhostar\ for
+either dwarf or giant stars.
+We found that the maximum {\it a-posteriori} values for the $\gamma$
+parameters were consistent with zero: $\gamma_{rho} = 0.006 \pm 0.02, \gamma_g
+= -0.01 \pm\ 0.01$, and interpret this as evidence for a constant intrinsic
+scatter level across evolutionary stages.
 
-We also performed this inference using an alternative, analytic form of the
-marginal likelihood, which accounts for 2-D uncertainties but does not allow
-the intrinsic scatter to be a function of the dependent or independent
-variables.
+% In order to account for the two-dimensional observational uncertainties in
+% this data set, we marginalize over the latent, `true' values of $F_8$,
+% \rhostar\ and \logg\ using the importance sampling method of
+% \citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
+% \citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
+% likelhood, we expect any bias introduced by the importance sampling method to
+% be negligable.}.
+% In what follows we give an overview of this procedure, but encourage the
+% reader to refer back  to this original reference for a more detailed
+% description of the mathematics.
+% The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+% thought of as a single `draw' from an underlying (assumed Gaussian)
+% probability
+% In what follows we give an overview of the importance sampling procedure, but
+% encourage the reader to refer back to the original reference for a more
+% detailed description of the mathematics.
+% The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be
+% thought of as a single `draw' from an underlying (assumed Gaussian)
+% probability distribution, with mean equal to the `true' value---the value that
+% would have been observed, given infinitely high signal-to-noise---and standard
+% deviation equal to the observational uncertainty.
+% We compute the likelihood of the data given the model, marginalized over the
+% true values, using importance sampling.
+% We sample from the posterior PDFs of the true values of the variables,
+% conditioned on the observed values, and compute the likelihood of each of those
+% samples.
+% % These posterior PDFs could be produced by the astronomers who provide the
+% % catalog.
+% % If posterior samples are made available, they can be used in this stage of
+% % the inference.
+% % In the event that they are not provided however, they can be generated.
+% % For example, the $log(g)$ values used here are point estimates of the
+% % posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
+% % {\it Kepler} light curves and stellar models.
+% We would use the posterior samples generated in previous model fitting process
+% if they were available, however generating new samples is an acceptable
+% approximation provided the posteriors are Gaussian and the priors used by the
+% previous fitters were uninformative.
+% Such an approximation will not be necessary for those studies using \rhostar\
+% or \logg\ calculated using the newly-calibrated relations presented here,
+% since we have published our posterior samples.
+% After producing new posterior PDF samples we then add up the individual
+% log-likelihoods of each sample to compute the total marginalized likelihood
+% for a star.
+% The marginalized likelihood of the whole dataset is then computed as the sum
+% of the log-likelihoods of each star.
+% The importance sampling method allows us to approximate the integral over the
+% latent variables.
+% % We used flat priors for all except the $\sigma$ parameters, for which we used
+% % priors which were flat in log-space.
+% as a function of flicker.
+
+We tested a likelihood function which accounts for 2-D uncertainties but does
+not allow the intrinsic scatter to be a function of the dependent or
+independent variables.
 For the relation between flicker and \rhostar, this likelihood function can be
@@
 and similarly for \logg.
-The parameter values found using this simple likelihood function, with a
-constant dispersion are included in table \ref{tab:results}.
-% The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
-% cases where the intrinsic scatter is not a function of $x$ or $y$.
+We found that the posterior PDFs for the model parameters obtained using this
+likelihood function were consistent with those obtained using a model that
+only accounts for the uncertainties on the flicker measurements, so instead
+used a standard Gaussian likelihood function $(-1/2\chi^2)$ in our analysis.
+% Since the uncertainties on flicker were significantly larger than those on
+% \logg or \rhostar, regression was performed with flicker as the dependent
+% variable.
+
+We used uniform priors over all our parameters:
+\begin{eqnarray}
+	&	\alpha, \beta, \delta, \epsilon
+	\sim U(-10:10) \\ \nonumber
+ 	&	\sigma_{rho}, \gamma, \sigma_g, \zeta  \sim U(-100:100).
+\end{eqnarray}
+\label{eq:priors}
+
+% The parameter values found using this simple likelihood function, with a
+% constant dispersion are included in table \ref{tab:results}.
+% % The analytic likelihood function, equation \ref{eq:likelihood} is suitable for
+% % cases where the intrinsic scatter is not a function of $x$ or $y$.
+
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit

### 2016-01-26 code/analytic_with_prior.py
adding analytic_with_prior.py

@@ -1,0 +1,150 @@
+# coding: utf-8
+
+# Testing the analytic lhf with jake's prior
+
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+import emcee
+
+
+# Define prior, log prob and log like
+def lnprior(pars):
+    m, b, s = pars
+    if s < 0:
+        return -np.inf
+    else:
+        return -1.5 * np.log(1 + b ** 2) - np.log(s)
+
+
+def lnprob(pars, x, y, xerr, yerr):
+    return lnprior(pars) + lnlike(pars, x, y, xerr, yerr)
+
+
+def lnlike(pars, x, y, xerr, yerr):
+    m, b, s = pars
+    sig2 = (m*xerr)**2 + (yerr**2 + s**2)
+    ll = -.5 * np.sum((m*x-y+b)**2/sig2 + np.log(sig2))
+    return ll
+
+
+# Load and plot the data
+# load data
+f, ferr, r, rerr = np.genfromtxt("../data/flickers.dat").T
+r -= 3  # convert rho to g/cm^3
+
+# fit a line
+AT = np.vstack((r, np.ones_like(r)))
+ATA = np.dot(AT, AT.T)
+m, c = np.linalg.solve(ATA, np.dot(AT, f))
+print("params = ", m, c)
+
+# plot data with best fit line
+xs = np.linspace(min(r)-.2, max(r)+.2, 100)
+ys = m * xs + c
+plt.clf()
+plt.errorbar(r, f, xerr=rerr, yerr=ferr, fmt="k.", capsize=0)
+plt.plot(xs, ys)
+plt.ylabel("log flicker")
+plt.xlabel("log rho")
+
+pars_init = [m, c, 0.06]
+print(lnlike(pars_init, r, f, rerr, ferr)) # check the lhf works
+
+ndim, nwalkers = 3, 24
+pos = [pars_init + 1e-4 * np.random.randn(ndim) for i in range(nwalkers)]  # initialisation
+sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=(r, f, rerr, ferr))
+sampler.run_mcmc(pos, 2000)  # run MCMC
+samples = sampler.chain[:, 200:, :].reshape((-1, ndim))  # cut off burn in and flatten chains
[…]

### 2016-01-26 code/analytic_with_prior.py
small changes

@@ -62,8 +62,8 @@ plt.clf()
 fig = corner.corner(samples, labels=["m", "c", "s"])
 
 # Make figure for paper
-plotpar = {'axes.labelsize': 18,
-           'text.fontsize': 14,
+plotpar = {'axes.labelsize': 25,
+           'text.fontsize': 18,
            'legend.fontsize': 18,
            'xtick.labelsize': 18,
            'ytick.labelsize': 18,
@@ -90,12 +90,16 @@ plt.xlabel("$\log_{10}(\\rho_{\star}[\mathrm{g~cm}^{-3}])$")
 #plt.text(-1.15, 2.25, "$\log_{10} (\\rho_{\star}) \sim \mathcal{N} \left(\\alpha_\\rho + \\beta_\\rho \log_{10}(F_8), \
 #                \\sigma=\\sigma_{\\rho}\\right)$")
 plt.text(-1.15, 2.25, "$\log_{10} (F_8) \sim \mathcal{N} \left(\\alpha_\\rho + \\beta_\\rho \log_{10}(\\rho_{\star}),                 \\sigma=\\sigma_{\\rho}\\right)$")
-plt.text(-1.45, 1.2, "$\\alpha_\\rho = %.2f \pm %.2f$" % (np.round(alpha, 2), np.round(ap, 2)))
-plt.text(-1.45, 1., "$\\beta_\\rho = %.2f \pm %.2f $" % (np.round(beta, 2), np.round(bp, 2)))
-plt.text(-1.45, .8, "$\\sigma_{\\rho} = %.3f \pm %.3f $" % (np.round(sigma, 2), np.round(sp, 3)))
+plt.text(-1.45, 1.2, "$\\alpha_\\rho = %.3f \pm %.3f$" % (np.round(alpha, 3), np.round(ap, 3)))
+plt.text(-1.45, 1., "$\\beta_\\rho = %.3f \pm %.3f $" % (np.round(beta, 3), np.round(bp, 3)))
+plt.text(-1.45, .8, "$\\sigma_{\\rho} = %.3f \pm %.3f $" % (np.round(sigma, 3), np.round(sp, 3)))
 plt.xlim(min(r)-.2, max(r)+.2)
 plt.ylim(.4, 2.5)
 plt.savefig("../version1.0/flicker_vs_rho.pdf")
+print("rho")
+print(alpha, ap, am)
+print(beta, bp, bm)
[…]

### 2016-01-26 version1.0/manuscript.tex
changes to figures, tables and equations

@@
 % 	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
-	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(F_8), \sigma^2 = \sigma_\rho^2 \right),
+	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
+    \beta_\rho \log_{10}(\rho_\star), \sigma^2 = \sigma_\rho^2 \right),
 \end{equation}
@@
 % 	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
-	\log_{10}(g) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
-    \log_{10}(F_8), \sigma^2 = \sigma_g^2 \right).
+	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
+    \log_{10}(g), \sigma^2 = \sigma_g^2 \right).
 \end{equation}
@@
 \begin{equation}
-	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 +
+	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
+    \beta_\rho \log_{10}(\rho_\star), \sigma = \sqrt{\sigma_{\rho}^2 +
     \gamma_\rho F_8}\right),
@@
 This model allowed us to determine whether there the magnitude of additional
-scatter varied as a function of \logg or \rhostar.
+scatter varied as a function of flicker.
 In other words, whether flicker was a better proxy for \logg\ or \rhostar\ for
@@
 
-We tested a likelihood function which accounts for 2-D uncertainties but does
+We used a likelihood function which accounts for 2-D uncertainties but does
 not allow the intrinsic scatter to be a function of the dependent or
@@
 \begin{eqnarray}
-	& p(\rho_*| F_8, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
+	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
 						      & \exp \left[-\frac{1}{2}
@@
 likelihood function were consistent with those obtained using a model that
-only accounts for the uncertainties on the flicker measurements, so instead
-used a standard Gaussian likelihood function $(-1/2\chi^2)$ in our analysis.
+only accounts for the uncertainties on the flicker measurements.
+Accounting for uncertainties on $y$ {\it and} $x$ is therefore not essential
+in this case but still good practise and will, in general, produce more
+accurate model parameters and uncertainties.
+
+% so instead
+% used a standard Gaussian likelihood function $(-1/2\chi^2)$ in our analysis.
 % Since the uncertainties on flicker were significantly larger than those on
@@
 
-We used uniform priors over all our parameters:
-\begin{eqnarray}
-	&	\alpha, \beta, \delta, \epsilon
-	\sim U(-10:10) \\ \nonumber
- 	&	\sigma_{rho}, \gamma, \sigma_g, \zeta  \sim U(-100:100).
-\end{eqnarray}
+We used the uninformative prior for the parameters of a straight line with
+unknown variance, outlined in \citet{vanderplas},
+\begin{equation}
+P(\alpha, \beta, \sigma) \propto \frac{1}{\sigma} \left( 1 + \beta^2
+\right)^{-3/2}.
+\end{equation}
 \label{eq:priors}
@@
 \begin{center}
+\includegraphics[width=8.4cm,angle=0,clip=true]{flicker_vs_rho.pdf}
 % \includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
-\includegraphics[width=8.4cm,angle=0,clip=true]{new_rho.pdf}
+% \includegraphics[width=8.4cm,angle=0,clip=true]{new_rho.pdf}
 \caption{
@@
 This figure shows the model, conditioned on the data.
-The solid black line shows the highest-likelihood sample, and the pink shaded
-regions show the 1 and 2$\sigma$ confidence intervals.}
+The solid black line shows the model with the parameter values quoted in the
+text.
+The solid pink lines show the 1$\sigma$ region where the extra scatter is not
+included and the pink shaded regions show the 1 and 2$\sigma$ regions with the
+additional scatter.}
 \label{fig:rhostar}
@@
 \begin{center}
+\includegraphics[width=8.4cm,angle=0,clip=true]{flicker_vs_logg.pdf}
 % \includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
-\includegraphics[width=8.4cm,angle=0,clip=true]{new_logg.pdf}
+% \includegraphics[width=8.4cm,angle=0,clip=true]{new_logg.pdf}
 \caption{
@@
 As in \ref{fig:rhostar} this figure shows the model, conditioned on the data.
-The solid black line shows the highest-likelihood sample, and the blue shaded
-regions show the 1 and 2$\sigma$ confidence intervals.}
+The solid black line shows the model with the parameter values quoted in the
+text.
+The solid blue lines show the 1$\sigma$ region where the extra scatter is not
+included and the blue shaded regions show the 1 and 2$\sigma$ regions with the
+additional scatter.}
 \label{fig:logg}
@@
 \begin{table}
-\caption{Highest-likelihood parameter values with 16th and 84th
+\begin{center}
+\caption{Median parameter values with 16th and 84th
 	percentile uncertainties.}
-\begin{tabular}{lccc}
+\begin{tabular}{lcc}
 \hline\hline
-Parameter & Variable dispersion model & Constant dispersion model \\
+Parameter & Median value \\
     \hline
-$\alpha$ &    2.358$_{-0.04}^{+0.08}$ &       2.191$_{-0.05}^{+0.05}$  \\
-$\beta$ &    -1.79$_{-0.03}^{+0.04}$ &        -1.724$_{-0.03}^{+0.03}$ \\
-$\sigma_{rho}$ &    0.015$_{-0.04}^{+0.08}$ &   0.085$_{-0.05}^{+0.05}$  \\
-$\gamma$ &    -0.0$_{-0.04}^{+0.08}$ & \\
+$\alpha_\rho$   &    1.315$\pm 0.008$   \\
+$\beta_\rho$    &    -0.526$\pm 0.009$   \\
+$\sigma_\rho$   &    0.065$\pm 0.003$   \\
+% 1.31476104613 0.00759733599155 0.00767814819025
+% -0.525867117083 0.00913323668806 0.00926420817321
+% 0.0650160215492 0.00300880167448 0.00288553857806
 \hline
-$\delta$ &    5.535$_{-0.07}^{+0.07}$ &      5.739$_{-0.03}^{+0.03}$  \\
-$\epsilon$ &    -1.005$_{-0.04}^{+0.03}$ &   -1.095$_{-0.02}^{+0.01}$ \\
-$\sigma_g$ &    0.168$_{-0.07}^{+0.07}$ &    0.015$_{-0.03}^{+0.03}$  \\
-$\zeta$ &    -0.014$_{-0.07}^{+0.07}$ & \\
+$\alpha_g$      &    4.90$\pm 0.05$   \\
+$\beta_g$       &   -0.83$\pm 0.01$   \\
+$\sigma_g$      &    0.056$\pm 0.003$   \\
+% 4.90451472909 0.0514398889837 0.0509386027456
+% -0.8252952869 0.0130730719222 0.0131097233196
+% 0.0565263079576 0.00279179051697 0.00271277097572
     \hline
 \end{tabular}
+\end{center}
 \end{table}
@@
 stellar density and surface gravity, whilst including parameters to describe
-the intrinsic scatter in these relationships.
-We included both an additive variance term {\it and} a term that allows for
-some abscissa-dependent variance, presented in table \ref{tab:table}.
+the intrinsic scatter in these relationships, presented in table
+\ref{tab:results}.
 The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both non-zero,
@@
 observational uncertainties.
-Interestingly, the abscissa-dependent variances, $\gamma$ and $\zeta$ are both
-less than zero.
-This indicates that the intrinsic scatter either decreases with increasing
-flicker and decreasing stellar density and surface gravity, or that the
-uncertainty underestimation is more severe for the giant stars.
-Regardless of the cause of the additional scatter needed to describe the
-relations between these parameters, the conclusions of this study are the same.
-Firstly, naively modelling the relationship between flicker and surface
-gravity or stellar density, or any other stochastic relationship
-deterministically, will lead to an underprediction of the uncertainties in
-dependent variable.
-Secondly, we find marginal evidence for decreased scatter at higher surface
-gravities and densities, i.e. the relations between flicker, surface gravity
-and stellar density are valid for dwarfs and giants alike, despite the fact
-that observational uncertainties tend to be larger for dwarfs.
+We also tested a model with both an additive variance term {\it and} a term
+that included flicker-dependent variance.
+We found that the need for additional flicker-dependent variance was not
+supported by the data, indicating that the intrinsic scatter in the relations
+between flicker, \logg\ and \rhostar\ does not depend on evolutionary state.
 
@@
 We used Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
-scatter in the relationship between flicker, surface gravity and density.
-We also included the effects of the non-negligable two-dimensional observational
-uncertainties by marginalizing over latent variables using importance sampling.
+scatter in the relationship between flicker, surface gravity and density and
+included the effects of the non-negligable two-dimensional observational
+uncertainties.
 % Although none of these methods, nor the data used here are new taken alone,
@@
 for a given estimate of flicker, we recommend using these posterior samples
-within the calculation of $\rho_{*}$ or $\log(g)$ and its (Monte Carlo)
+within the calculation of \rhostar or \logg\ and its (Monte Carlo)
 uncertainty.
-These posterior samples will naturally fold-in the covariances between
+These posterior samples will naturally fold in the covariances between
 parameters.
@@
 on the flicker measurement, the uncertainties on the model parameters {\it and}
-the intrinsic scatter in the flicker-$\rho_{*}$-$\log(g)$ relations.
+the intrinsic scatter in the flicker-\rhostar-\logg\ relations.
 
@@
 \bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
+\bibitem[VanderPlas(2014)]{vanderplas}
+VanderPlas, J.\ 2014, arXiv:1411.5018
 Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted

### 2016-01-27 version1.0/manuscript.tex
small changes from referee and equation corrections

@@
 relationships between observables to be stochastic.
-Unlike the deterministic case, a single observational input is intepreted to
-be caused by a range of possible model parameters, described by a probability
-distribution.
-This statement is true even in the case of a perfect observation of infinite
-signal-to-noise, since the underlying model itself comes with uncertainty.
+% Unlike the deterministic case, a single observational input is interpreted to
+% be caused by a range of possible model parameters, described by a probability
+% distribution.
+% This statement is true even in the case of a perfect observation of infinite
+% signal-to-noise, since the underlying model itself comes with uncertainty.
 In recent years, there has been a shift towards such methods in several areas
@@
 
-The demand for probabilistic stellar parameters is not only motivated by the fact
-that probability distributions are far more representative of our `beliefs'
-about astrophysical parameters, it also has a practical purpose.
-When using data published in the astronomical literature to, for example, infer
-relationships between observed parameters, that inference can be performed as
-the final stage in a hierarchical treatment
-\citep[see, e.g.][]{foreman-mackey:2014}.
-This is useful if, for example, one wants to account for multi-dimensional
-uncertainties by marginalizing over the `true' parameter values upon which the
-noisy observations are conditioned.
-This point is discussed further in \textsection\ref{sec:HBM}.
+The demand for probabilistic stellar parameters is not only motivated by the
+fact that probability distributions are far more representative of our
+`beliefs' about astrophysical parameters, it also has a practical purpose.
+When using data published in the astronomical literature to, for example,
+infer relationships between parameters that are themselves the product of an
+inference process (for example, exoplanet transit depth and period), inference
+can be performed as the final stage in a hierarchical treatment \citep[see,
+e.g.][]{foreman-mackey:2014}.
+% This is useful if, for example, one wants to account for multi-dimensional
+% uncertainties by marginalizing over the `true' parameter values upon which the
+% noisy observations are conditioned.
+Studies such as these are benefited by posterior PDF samples, rather than
+point estimates of inferred properties.
+% This point is discussed further in \textsection\ref{sec:HBM}.
 
@@
 Turning this around, the observation implies that flicker can be used to
-blindly infer surface gravities at the level of $\sim0.1$\,dex, an attractive
-proposition given the wealth of photometric light curves available through the
-array of exoplanet transit missions flying and scheduled to launch.
+empirically infer surface gravities at the level of $\sim0.1$\,dex, an
+attractive proposition given the wealth of photometric light curves available
+through the array of exoplanet transit missions flying and scheduled to
+launch.
 
@@
 \item[{\tiny$\blacksquare$}] Range (defined in \citealt{bastien:2013})
-$<1$\,ppt
+$<1000$\,ppm
 \item[{\tiny$\blacksquare$}] $4500<T_{\mathrm{eff}}<6500$\,K
@@
 	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(\rho_\star), \sigma^2 = \sigma_\rho^2 \right),
+    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_\rho^2 \right),
 \end{equation}
@@
 	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
-    \log_{10}(g), \sigma^2 = \sigma_g^2 \right).
+    \log_{10}(g), \sigma^2 + \sigma_g^2 \right).
 \end{equation}
@@
 	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(\rho_\star), \sigma = \sqrt{\sigma_{\rho}^2 +
-    \gamma_\rho F_8}\right),
+    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_{\rho}^2 +
+    \gamma_\rho F_8 \right),
 \end{equation}
@@
 \begin{eqnarray}
-	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
+	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) \propto  \\ \nonumber
 						      & \exp \left[-\frac{1}{2}
 		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
-	{\left[\beta \sigma_{F8, n}^2 + (\sigma_{\rho *, n}
-	+ \sigma_{\rho})^2\right]}\right]
+	{\left[\beta \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2
+	+ \sigma_{\rho}^2\right]}\right]
 	\\ \nonumber
-	& + \beta^2 \sigma_{F8, n}^2 + (\sigma_{\rho *, n} + \sigma_{\rho})^2,
+	& + \beta^2 \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 + \sigma_{\rho}^2,
 \end{eqnarray}
@@
 
-We used the uninformative prior for the parameters of a straight line with
-unknown variance, outlined in \citet{vanderplas},
+We used the uninformative prior for the parameters of a straight line for
+data with unknown uncertainties, outlined in \citet{vanderplas},
 \begin{equation}
@@
 \begin{center}
-\caption{Median parameter values with 16th and 84th
-	percentile uncertainties.}
+\caption{Median parameter values with 1$\sigma$ uncertainties.}
 \begin{tabular}{lcc}
@@
 falsely assumed to be deterministic, they will be skewed by data points with
-unrepresentative uncertainties.
+uncertainties that only represent measurement error and no additional scatter.
 The prediction stage benefits from the stochastic treatment both because a
@@
 {seager:2003}Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
-\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
 \bibitem[VanderPlas(2014)]{vanderplas}
 VanderPlas, J.\ 2014, arXiv:1411.5018
+\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
 Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted

### 2016-02-02 version1.0/manuscript.tex
sent to kipping
equation changes, a few more words. ready to resubmit

@@
 for flicker and ``range'' were taken from \citet{kipping:2014}, based upon the
-methods described in \citet{bastien:2013}. Following \citet{kipping:2014} and
+methods described in \citet{bastien:2013}.
+% Following \citet{kipping:2014} and
+% for reasons described there-in, we only include targets in our calibration for
+% which:
+In order to use the same data set as \citet{kipping:2014} and
 for reasons described there-in, we only include targets in our calibration for
@@
 \begin{eqnarray}
-	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) \propto  \\ \nonumber
+	& p(F_8|\rho_\star, \alpha_\rho, \beta_\rho, \sigma_{\rho}) \propto  \\ \nonumber
 						      & \exp \left[-\frac{1}{2}
-		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
-	{\left[\beta \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2
+		\sum_{n=1}^N \frac{[F_{8n}-(\alpha_\rho + \beta_\rho \rho_{*n})]^2}
+	{\left[\beta_\rho \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2
 	+ \sigma_{\rho}^2\right]}\right]
 	\\ \nonumber
-	& + \beta^2 \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 + \sigma_{\rho}^2,
+	& + \beta_\rho^2 \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 + \sigma_{\rho}^2,
 \end{eqnarray}
@@
 \begin{equation}
-P(\alpha, \beta, \sigma) \propto \frac{1}{\sigma} \left( 1 + \beta^2
+p(\alpha, \beta, \sigma) \propto \frac{1}{\sigma} \left( 1 + \beta^2
 \right)^{-3/2}.
@@
 
+We also tested flat priors and found that the results were insensitive to this
+choice.
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
@@
 This figure shows the model, conditioned on the data.
-The solid black line shows the model with the parameter values quoted in the
-text.
+The solid black line shows the model with the best-fitting parameter values
+quoted in the text.
 The solid pink lines show the 1$\sigma$ region where the extra scatter is not
@@
 As in \ref{fig:rhostar} this figure shows the model, conditioned on the data.
-The solid black line shows the model with the parameter values quoted in the
-text.
+The solid black line shows the model with the best-fitting parameter values
+quoted in the text.
 The solid blue lines show the 1$\sigma$ region where the extra scatter is not

### 2016-02-05 version1.0/manuscript.tex
emcee
adding posterior sample url
changing the abstract

@@
 
-% aims
 The relations between observable stellar parameters are usually assumed to be
@@
 `$y$'s.
-% methods
 The relation between short-timescale brightness fluctuations (flicker) of
@@
 the relations.
-% Our treatment also fully accounts for the observational uncertainties on all
-% variables.
-% results
-We find evidence for additional scatter in the relationships, particularly in
-the relation between flicker and surface gravity, that cannot be accounted for
-by the observational uncertainties alone.
-The scatter in surface gravity increases with decreasing flicker, i.e. there
-is greater scatter for subgiants than for dwarfs, suggesting that using flicker
-as a proxy for surface gravity is just as (if not more) valid for dwarf stars,
-despite the fact that the observational uncertainties tend to be larger for
-these stars.
+We find evidence for additional scatter in the relationships, that cannot be
+accounted for by the observational uncertainties alone.
+The scatter in surface gravity and stellar density does not depend on flicker,
+suggesting that using flicker as a proxy for $\log g$ and $\rho_\star$ is
+equally valid for dwarf and giant stars, despite the fact that the
+observational uncertainties tend to be larger for dwarfs.
 
@@
 about the mean.
+We used the MCMC package, {\tt emcee} \citep{foreman-mackey:2013} to explore the
+posterior PDFs of our model parameters.
 
@@
 We provide posterior PDF samples and the code used in this project at
-\url{https://github.com/RuthAngus/flicker}.
+\url{https://zenodo.org/deposit/95681/}.
+% https://github.com/RuthAngus/flicker
 Whenever a prediction for the surface gravity or density of a star is required,
@@
 ApJ, 795, 64
+\bibitem[Foreman-Mackey et al.(2013)]{foreman-mackey:2013}
+Foreman-Mackey, D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2013, \pasp, 125, 306
 \bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}

### 2016-02-25 code/triangle_plot.py
adding triangle plot

@@ -1,0 +1,21 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py
+import corner
+
+plotpar = {'axes.labelsize': 15,
+           'text.fontsize': 18,
+           'legend.fontsize': 18,
+           'xtick.labelsize': 15,
+           'ytick.labelsize': 15,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+with h5py.File("rho_samples.h5") as f:
+    samples = f["samples"][...]
+nwalkers, nsteps, ndim = np.shape(samples)
+flat = samples.reshape((nwalkers*nsteps, ndim))
+
+labels = ["$\\alpha$", "$\\beta$", "$\sigma$"]
+fig = corner.corner(flat, labels=labels)
+plt.savefig("rho_triangle.pdf")

### 2016-02-25 version1.0/manuscript.tex
starting to add referee comments

@@
 only accounts for the uncertainties on the flicker measurements.
-Accounting for uncertainties on $y$ {\it and} $x$ is therefore not essential
-in this case but still good practise and will, in general, produce more
-accurate model parameters and uncertainties.
+The median values of the model parameters differed by around $0.05 \sigma$
+for the $\alpha$ and $\beta$ parameters, by $0.3 \sigma$ for $\sigma_\rho$
+and by 0.8 $\sigma$ for $\sigma_g$.
+Since they are so dependent on the observational uncertainties, the parameters
+that describe the intrinsic scatter in the relations are more sensitive to
+whether the uncertainties in the $x$-direction are included.
+Accounting for uncertainties on $y$ {\it and} $x$ is not essential in this
+case but is still good practise and will, in general, produce more accurate
+model parameters and uncertainties.
 
-% so instead
-% used a standard Gaussian likelihood function $(-1/2\chi^2)$ in our analysis.
 % Since the uncertainties on flicker were significantly larger than those on
@@
 
-We also tested flat priors and found that the results were insensitive to this
-choice.
+We also tested uniform, flat priors as defined below:
+\begin{eqnarray}
+ 	&	\alpha, \beta \sim U(-10:10) \\ \nonumber
+  	&	\log (\sigma_{rho}), \log(\sigma_g) \sim U(-10:10).
+ \end{eqnarray}
+ \label{eq:priors}
+We found that the results were relatively insensitive to the choice of prior,
+with median parameter values differing by only around $0.05\sigma$.
+MCMC chains were run until the Gelman \& Rubin convergence criterion,
+$\hat{R}$ reached a value of less than 1.002 and the number of autocorrelation
+times was greater than 35.
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
@@
 representative of the intrinsic scatter in the relations.
-% Given a measurement of flicker, the predictive probability distribution of
-% \logg and \rhostar will be a Gaussian distribution
+The marginal posterior PDFs of the model parameters for \rhostar are shown in
+figure \ref{fig:triangle}.
+The marginal posterior PDFs for \logg are similarly Gaussian and, as with
+$\sigma_\rho$, $\sigma_g$ is clearly greater than zero.
 
@@
 \label{fig:logg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=8.4cm,angle=0,clip=true]{rho_triangle.pdf}
+\caption{Marginal posterior PDFs of the model parameters for \rhostar.
+This figure was generated using the {\tt corner} python package
+\citep{corner}.}
+\label{fig:triangle}
 \end{center}
@@
 RA would like to thank Dan Foreman-Mackey for his extremely helpful suggestions
-and comments, Angie Wolfgang for her expert guidance and John Johnson for hit
+and comments, Angie Wolfgang for her expert guidance and John Johnson for his
 fantastic advice and support.
-RA would also like to thank the SAMSI institute and the other members of the
-SAMSI {\it Kepler} exoplanet statistics working groups.
+RA would also like to thank the anonymous referee for their excellent
+suggestions, SAMSI institute and the other members of the SAMSI {\it Kepler}
+exoplanet statistics working groups.
 
@@
 Foreman-Mackey, D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2013, \pasp, 125, 306
+\bibitem[Foreman-Mackey et al. (2014)]{corner}
+Foreman-Mackey, D., Price-Whelan, A., Ryan, G. et al.\ 2014,
+10.5281/zenodo.10598, http://dx.doi.org/10.5281/zenodo.10598
 \bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}

### 2016-02-25 code/analytic_with_prior.py

@@ -1,74 +1,20 @@
 # coding: utf-8
-
-# Testing the analytic lhf with jake's prior
-
 import numpy as np
 import matplotlib.pyplot as plt
 import corner
-import emcee
+import h5py
 
-
-# Define prior, log prob and log like
-def lnprior(pars):
-    m, b, s = pars
-    if s < 0:
-        return -np.inf
-    else:
-        return -1.5 * np.log(1 + b ** 2) - np.log(s)
-
-
-def lnprob(pars, x, y, xerr, yerr):
-    return lnprior(pars) + lnlike(pars, x, y, xerr, yerr)
-
-
-def lnlike(pars, x, y, xerr, yerr):
-    m, b, s = pars
-    sig2 = (m*xerr)**2 + (yerr**2 + s**2)
-    ll = -.5 * np.sum((m*x-y+b)**2/sig2 + np.log(sig2))
-    return ll
-
-
-# Load and plot the data
-# load data
-f, ferr, r, rerr = np.genfromtxt("../data/flickers.dat").T
-r -= 3  # convert rho to g/cm^3
-
-# fit a line
-AT = np.vstack((r, np.ones_like(r)))
-ATA = np.dot(AT, AT.T)
-m, c = np.linalg.solve(ATA, np.dot(AT, f))
-print("params = ", m, c)
-
-# plot data with best fit line
-xs = np.linspace(min(r)-.2, max(r)+.2, 100)
-ys = m * xs + c
-plt.clf()
-plt.errorbar(r, f, xerr=rerr, yerr=ferr, fmt="k.", capsize=0)
-plt.plot(xs, ys)
-plt.ylabel("log flicker")
-plt.xlabel("log rho")
-
-pars_init = [m, c, 0.06]
-print(lnlike(pars_init, r, f, rerr, ferr)) # check the lhf works
-
-ndim, nwalkers = 3, 24
-pos = [pars_init + 1e-4 * np.random.randn(ndim) for i in range(nwalkers)]  # initialisation
-sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=(r, f, rerr, ferr))
-sampler.run_mcmc(pos, 2000)  # run MCMC
[…]

### 2016-02-25 code/triangle_plot.py

@@ -3,11 +3,11 @@ import matplotlib.pyplot as plt
 import h5py
 import corner
 
-plotpar = {'axes.labelsize': 15,
-           'text.fontsize': 18,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 15,
-           'ytick.labelsize': 15,
+plotpar = {'axes.labelsize': 19,
+           'text.fontsize': 19,
+           'legend.fontsize': 19,
+           'xtick.labelsize': 17,
+           'ytick.labelsize': 17,
            'text.usetex': True}
 plt.rcParams.update(plotpar)
 
@@ -16,6 +16,6 @@ with h5py.File("rho_samples.h5") as f:
 nwalkers, nsteps, ndim = np.shape(samples)
 flat = samples.reshape((nwalkers*nsteps, ndim))
 
-labels = ["$\\alpha$", "$\\beta$", "$\sigma$"]
+labels = ["$\\alpha_\\rho$", "$\\beta_\\rho$", "$\sigma_\\rho$"]
 fig = corner.corner(flat, labels=labels)
-plt.savefig("rho_triangle.pdf")
+plt.savefig("../version1.0/rho_triangle.pdf")

### 2016-04-01 version1.0/manuscript.tex
final changes

@@
 \begin{equation}
-% 	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
-% 	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
-	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_\rho^2 \right),
+	\log_{10}(F_8) \sim \mathcal{N} \left(\alpha_\rho +
+    \beta_\rho \log_{10}(\rho_\star), \sigma_\rho^2 \right),
 \end{equation}
@@
 \begin{equation}
-% 	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
-% 	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
-	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
-    \log_{10}(g), \sigma^2 + \sigma_g^2 \right).
+	\log_{10}(F_8) \sim \mathcal{N} \left(\alpha_g + \beta_g
+    \log_{10}(g), \sigma_g^2 \right).
 \end{equation}
@@
 These relations are Gaussian distributions with means given by the equation of
-a straight line, and standard deviations which describe the intrinsic scatter
+a straight line and standard deviations which describe the intrinsic scatter
 about the mean.
@@
 	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
-    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_{\rho}^2 +
-    \gamma_\rho F_8 \right),
+    \beta_\rho \log_{10}(\rho_\star), \sigma_{\rho}^2 + \gamma_\rho
+    \log_{10}(F_8)
+    \right),
 \end{equation}
@@
 \begin{eqnarray}
-	& p(F_8|\rho_\star, \alpha_\rho, \beta_\rho, \sigma_{\rho}) \propto  \\ \nonumber
-						      & \exp \left[-\frac{1}{2}
-		\sum_{n=1}^N \frac{[F_{8n}-(\alpha_\rho + \beta_\rho \rho_{*n})]^2}
-	{\left[\beta_\rho \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2
-	+ \sigma_{\rho}^2\right]}\right]
+	& \ln\left[p(F_8|\rho_\star, \alpha_\rho, \beta_\rho, \sigma_{\rho})\right]
+        \propto  \\ \nonumber
+    & -\frac{1}{2}\sum_{n=1}^N \left[\frac{[F_{8n}-(\alpha_\rho
+    + \beta_\rho \rho_{*n})]^2}
+	{\left[\beta_\rho \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 +
+    \sigma_{\rho}^2\right]} + \ln(\sigma_{F8, n}^2) + \ln(\sigma_{\rho*, n}^2)
+    + \ln(\sigma_\rho^2) \right]
 	\\ \nonumber
-	& + \beta_\rho^2 \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 + \sigma_{\rho}^2,
 \end{eqnarray}
@@
 only accounts for the uncertainties on the flicker measurements.
-The median values of the model parameters differed by around $0.05 \sigma$
-for the $\alpha$ and $\beta$ parameters, by $0.3 \sigma$ for $\sigma_\rho$
-and by 0.8 $\sigma$ for $\sigma_g$.
+{\bf The median values of the model parameters differed by around $0.05
+\sigma$ for the $\alpha$ and $\beta$ parameters, by $0.3 \sigma$ for
+$\sigma_\rho$ and by $0.8 \sigma$ for $\sigma_g$.
 Since they are so dependent on the observational uncertainties, the parameters
 that describe the intrinsic scatter in the relations are more sensitive to
-whether the uncertainties in the $x$-direction are included.
+whether the uncertainties in the $x$-direction are included.}
 Accounting for uncertainties on $y$ {\it and} $x$ is not essential in this
@@
 We found that the results were relatively insensitive to the choice of prior,
-with median parameter values differing by only around $0.05\sigma$.
+{\bf with median parameter values differing by only around $0.05\sigma$.
 MCMC chains were run until the Gelman \& Rubin convergence criterion,
 $\hat{R}$ reached a value of less than 1.002 and the number of autocorrelation
-times was greater than 35.
+times was greater than 35.}
 Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
@@
 representative of the intrinsic scatter in the relations.
-The marginal posterior PDFs of the model parameters for \rhostar are shown in
-figure \ref{fig:triangle}.
-The marginal posterior PDFs for \logg are similarly Gaussian and, as with
+{\bf The marginal posterior PDFs of the model parameters for \rhostar are
+shown in figure \ref{fig:triangle}.
+The marginal posterior PDFs for \logg\ are similarly Gaussian and, as with
 $\sigma_\rho$, $\sigma_g$ is clearly greater than zero.
+We checked the consistency between the two relations by calculating flicker
+values for the Sun, finding $F_8 = 1.24 \pm 0.07$ and $F_8 = 1.21 \pm 0.1$
+from Solar density and surface gravity measurements, respectively.
+All the code used for this project and several ipython notebooks explaining
+our analysis are available at {\url https://github.com/RuthAngus/flicker}.}
 
@@
     \hline
-$\alpha_\rho$   &    1.315$\pm 0.008$   \\
-$\beta_\rho$    &    -0.526$\pm 0.009$   \\
-$\sigma_\rho$   &    0.065$\pm 0.003$   \\
-% 1.31476104613 0.00759733599155 0.00767814819025
-% -0.525867117083 0.00913323668806 0.00926420817321
-% 0.0650160215492 0.00300880167448 0.00288553857806
+$\alpha_\rho$   &    1.31$\pm 0.01$   \\
+$\beta_\rho$    &    -0.53$\pm 0.01$   \\
+$\sigma_\rho$   &    0.060$\pm 0.003$   \\
 \hline
-$\alpha_g$      &    4.90$\pm 0.05$   \\
+$\alpha_g$      &    4.91$\pm 0.05$   \\
 $\beta_g$       &   -0.83$\pm 0.01$   \\
-$\sigma_g$      &    0.056$\pm 0.003$   \\
-% 4.90451472909 0.0514398889837 0.0509386027456
-% -0.8252952869 0.0130730719222 0.0131097233196
-% 0.0565263079576 0.00279179051697 0.00271277097572
+$\sigma_g$      &    0.060$\pm 0.003$   \\
     \hline
@@
 \ref{tab:results}.
-The additive terms, $\sigma_\rho$ and $ \sigma_g$ are both non-zero,
-suggesting that there {\it is} an additional source of scatter in the
-relations, not accounted for by the observational uncertainties alone.
+The terms $\sigma_\rho$ and $ \sigma_g$ are both non-zero, suggesting that
+there {\it is} an additional source of scatter in the relations, not accounted
+for by the observational uncertainties alone.
 This is either caused by intrinsic scatter in the physical relationship
@@
 observational uncertainties.
-We also tested a model with both an additive variance term {\it and} a term
+We also tested a model with both an additional variance term {\it and} a term
 that included flicker-dependent variance.
@@
 
-We provide posterior PDF samples and the code used in this project at
-\url{https://zenodo.org/deposit/95681/}.
-% https://github.com/RuthAngus/flicker
+We provide posterior PDF samples at \url{https://zenodo.org/deposit/105051/}.
+% https://zenodo.org/deposit/95681/}.
 Whenever a prediction for the surface gravity or density of a star is required,
@@
 fantastic advice and support.
-RA would also like to thank the anonymous referee for their excellent
-suggestions, SAMSI institute and the other members of the SAMSI {\it Kepler}
-exoplanet statistics working groups.
+RA would also like to thank the anonymous referee for their patience and
+excellent suggestions which dramatically improved this letter, the SAMSI
+institute and the other members of the SAMSI {\it Kepler} exoplanet statistics
+working groups.
 
@@
 Fuller, W.~A., 1987, Wiley Series in Probability and Mathematical Statistics
-\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
-Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
-2166
+% \bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
+% Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
+% 2166
 \bibitem[\protect\citeauthoryear{Hogg et al.}{2010b}]{hogg:2010b}
